---
title: "Hub in between two other modules"
author: "Daniel Lill"
---

# Load libraries and define some useful paths

```{r Libraries, results='hide', echo=FALSE, message=FALSE}
rm(list = ls())
library(plotly)
library(tidyverse)
library(MRAr)
library(scales)
library(xtable)
path <- paste0(getwd(), "/")
pathout <- paste0(path, "Output/")
```

```{r Names for Fit storage of different initial conditions}
fit_name <- "HubInBetween"
```


```{r Call render without overwriting any previous documents, echo=FALSE, eval=FALSE, message=FALSE}
mytime <- format(Sys.time(), "%Y%m%d-%H%M")

rmarkdown::render("HubInBetween.Rmd", output_format = "html_document", output_file = paste0(pathout, fit_name, mytime, ".html"))
beepr::beep("~/Promotion/Software/rimshot.mp3")
```
# Cascade

## Setup Model 
This is the setup of the ODE model.

### Prediction function
odemodel writes a C-function, which is executed in x and xs.
x is a regular prediction function while xs only finds and returns the steady states.
If the odemodel has been calculated already, it can be loaded with "load".
```{r Prediction function Cascade, results='asis'}
modelpath <- paste0(path, "Models/HubInBetween/")
modelname   <- "HubInBetween"
mymodel     <- as.eqnlist(read.csv(paste0(modelpath, modelname, ".csv")))

# myodemodel <- odemodel(mymodel, modelname = paste0(modelname), deriv = T)
# save(myodemodel, file = paste0(modelpath, "odemodel",modelname, ".RData"))
load(paste0(modelpath, "odemodel",modelname, ".RData"))

# Prediction function
x <- Xs(odemodel = myodemodel)
# Prediction function for steady states
xs <- Xs_steady(myodemodel) # When computing steady states, set deriv = F
# print(mymodel)
# mymodel %>% as.eqnvec() %>% print(pander = T)
# myodemodel[[1]] %>% attr("eq")
```

### Modules, their observables and the observation function
Observation function which depends also on the optimization parameters. With their help, the complexes can smoothly be switched on in the observation.
```{r Observation function with Cx2z2 in x-module}
modules <- modules0 <- c("u", "x", "y", "z", "w")
complexes <- c("Cu2x1", "Cx2y1", "Cx2z1", "Cz2w2")

to_single_module_pars_opt <- paste0("a_to_", outer(modules, complexes, paste, sep = "_")) %>% matrix(ncol = 4)

to_single_module <- prodSymb(to_single_module_pars_opt, t(t(complexes)))
# modules_grid <- outer(modules,modules, paste0)
# modules_grid[upper.tri(modules_grid)] <- t(modules_grid)[upper.tri(modules_grid)]
# diag(modules_grid) <- "0"
# to_both_modules <- paste0("a_to_", outer(modules_grid, complexes, paste, sep = "_")) %>% matrix(ncol = 4)

free_species <- c("u2","x2","y2","z2","w2")


obs <- obs0 <- paste(free_species, to_single_module, sep = " + ")
  
  
g <-  g0 <- Y(as.eqnvec(structure(obs0, names = modules0)), mymodel, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```

### Inner parameters
Load all parameters and format them for further use.
```{r Parameters Cascade}

pars_raw <- read.csv(paste0(modelpath, "pars.csv"), header = FALSE)  # dynamical parameters
pars_raw <- structure(pars_raw$V2, names = as.character(pars_raw$V1))
ic_raw <- read.csv(paste0(modelpath, "IC.csv"), header = F)          # initial conditions
ic_raw <- structure(ic_raw$V2, names = as.character(ic_raw$V1))
# sort ics like they appear in x, so it can be used in stode()
vars_x <- attr(x, "parameters")[attr(x, "parameters") %in% names(ic_raw)]
ic_raw <- ic_raw[vars_x]

pars_inner_opt_0 <- setdiff(attr(g, "parameters"),names(c(pars_raw,ic_raw)))
pars_inner_opt_0 <- structure(rep(0, length(pars_inner_opt_0)), 
                              names = pars_inner_opt_0)

pars_inner_0 <- pars_inner <- c(ic_raw, pars_raw, pars_inner_opt_0)


# check if all parameters are assigned a value. Both lines should evaluate to the same number.
# x_pars <- unique(union((g) %>% attr("parameters"),(xs) %>% attr("parameters")))
# names(pars_inner) %in% x_pars %>% sum()
# x_pars %in% names(pars_inner) %>% sum()
```


### General log transformation and setting some pars to zero
It's better to do a logtransformation for all Fitting purposes and work with the log-Parameters.
The way to implement is this, is
log_pars %>% p_log %>% p_other %>% p_other_2 %>% ...
```{r}
# Logtrafo of all pars
logtrafo <- structure(paste0("exp( log", names(pars_inner_0), ")"), 
                             names = names(pars_inner_0))


# initial values
logtrafo[c("x2", "Cx2y1", "y2", "Cx2z1", "z2", "u2", "Cu2x1", "w2", "Cz2w2")] <- 0


p_log <- P(trafo = logtrafo,
           compile = TRUE,
           modelname = "p_log")
           

pars_0 <- pars <- structure(log(pars_inner_0+.Machine$double.eps), names = paste0("log", names(pars_inner_0))) 

pars_opt_0 <- pars_opt <- structure(log(pars_inner_opt_0+.Machine$double.eps), names = paste0("log", names(pars_inner_opt_0)))

# pars_opt_0 <- pars_opt <- pars_opt_0[c("loga_1", "loga_2")]
# pars_0
# p_log(pars_0, deriv = F)
```


```{r}
# Which pars shall be perturbed?
pars_perturbed_0 <- pars_perturbed <- c("logu1" = log(0.9), "logx1" = log(0.9), "logy1" = log(0.9), "logz1" = log(0.9), "logw1" = log(0.9))
# pars_perturbed_0 <- pars_perturbed <- c("logx1" = log(0.8), "logy1" = log(0.8), "logz1" = log(0.8))



# pars_perturbed
p_pert <- p_pert_fun(pars_perturbed = pars_perturbed, pars = pars)

```


```{r}

find_small_before_and_big_after_optimization <- function(r_0, r_opt) {
  small_before <- abs(r_0) < 1e-2
  big_after <- abs(r_opt) > 1e-2
  return(small_before & big_after)
}


```

```{r}
pars <- c(100,0,0,       100,0,0,      100,0,0,      100,0,0,      100,0, 
          100,1,100,1,     0.01,0.4,0.05, #km1 - k3cat
          100,5,            0.05,0.4,0.008, #km5 - k6cat
          100,1, # km8 v8
          0.01,0.5,0.01,    100,1.5, #k91 - v10
          0.01,0.4,0.05,    100,1.5, #k111 - v12
          rep(0, length(pars_opt_0))) %>%
  set_names(names(pars_inner_0)) # as in cascade

pars_0 <- pars <- log(pars+.Machine$double.eps) %>% set_names(names(pars_0))# large complex concentrations

perturbation_prediction_0 <- perturbation_prediction <-  (xs*p_log*p_pert)(times = c(0,Inf), pars = pars, deriv = F)
# perturbation_prediction %>% lapply(round,1)

# mytime <- 20000
# (x*p_log*p_pert)(times = seq(0,20000, length.out = 50), pars = pars) %>%
# #   # getDerivs() %>%
# #   # extract2(1) %>%
# #   # apply_expression({x[,c(str_detect(colnames(x), "(Cx2y1\\.)|time"))]}) %>% prdlist() %>% set_names("blas") %>%
#   plotPrediction()
```


```{r}
R <- R_fun(perturbation_prediction_0)

local_response_matrix_eq10 <- function(R, made_easy = F) {
  
  if((!made_easy) & (dim(R)[1]==dim(R)[2])) {
    R_svd <- svd(R)
    
    condition_number <- R_svd$d[1]/R_svd$d[length(R_svd$d)]
    if(abs(condition_number) > 1e12) warning("R is ill conditioned. Condition number is ", condition_number)
    
    R_inv <- R_svd$v %*% diag(1/R_svd$d) %*% t(R_svd$u)
    
    R_inv_diag <- diag(R_inv)
    
    r <-  - R_inv / R_inv_diag
    dimnames(r) <- list(rownames(R), rownames(R))
  } else {
    
    
    
  }
  
  return(r)
}


```

```{r}
R <- matrix(c(1,1,1,0,1,1,0,0,1), nrow = 3)
R <- outer(1:3,1:3, paste0) %>% as.numeric() %>% matrix(ncol=3)
# r <- local_response_matrix_eq10(R) %>% round(1)
# r <- R


R <- R_fun(perturbation_prediction = perturbation_prediction_0)
not_perturbed <- c(1)
# r_made_easy <- function(R, not_perturbed)
R[ , not_perturbed] <- rep(0, nrow(R))

matlist <- lapply(1:nrow(R), function(i) t(R)[-i,-i])
A <- Matrix::bdiag(matlist) %>% as.matrix()
# A
vecR <- t(R)[!diag(T, nrow = nrow(R), ncol = ncol(R))]
vecr <- MASS::ginv(A)%*%vecR

r <- diag(rep(-1, nrow(R)))
r[!diag(T, nrow = nrow(R), ncol = ncol(R))] <- vecr
r <- t(r)
r %>% round(3)
local_response_matrix_eq10(R_fun(perturbation_prediction = perturbation_prediction_0)) %>% round(3)

```


# 1st measurement: Detect that there is sequestration
```{r}
r_0 <-   R_fun(pars_opt = pars_opt_0,
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_0%>% round(2)

find_symmetric_elements_both_nonzero(r_0)
```

# Iteration 1: Complexes are measured, add them to communicating species
It is assumed to be known, which species are active in the complexes.
Therefore, the complexes can be added only to the regulating modules.
Subsequently, make the first iteration of optimization

```{r}
pars_opt_subset <- paste0("log",to_single_module_pars_opt[c(1,7,12,19)])
pars_opt_subset

r_kept_1 <- r_kept_fun(pars_opt = pars_opt[pars_opt_subset],
                       perturbation_prediction = perturbation_prediction, 
                       obs_fun = g,
                       p_fun = (p_log * p_pert), 
                       pars = pars,
                       alpha = -log(0+.Machine$double.eps))
r_kept_1

r_0%>% round(2)

r_1 <-   R_fun(pars_opt = structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)

myfits_1 <- mstrust(obj_alpha, center =  structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset), studyname = "Fits", cores = 3, fits = 6, sd = 1, mypars = pars, perturbation_prediction = perturbation_prediction_0, r_kept = r_kept_1)
myfits_1 %>% as.parframe()
best_fit_1 <- myfits_1 %>% as.parframe() %>% as.parvec()
# exp(best_fit)
r_opt_1 <- r_alpha_fun(pars_opt = unclass(best_fit_1), pars = pars, perturbation_prediction = perturbation_prediction) %>% round(2)
r_opt_1
# pars
```

# Iteration 2: some connections are increased by the optimization procedure
Determine, which complexes have to be added where for further disentangling
```{r}

find_small_before_and_big_after_optimization(r_0,r_opt_1)
r_kept_2 <- r_kept_1 | find_small_before_and_big_after_optimization(r_0,r_opt_1)

r_kept_2
# matrix(1:20, ncol = 4)
# to_single_module_pars_opt
```


```{r}
pars_opt_subset <- paste0("log",to_single_module_pars_opt[c(1,7,12,19, # as before
                                                            13,18,     # to y: cxz and czw
                                                            9)])       # to z: cxy

r_0%>% round(2)

r_1 <-   R_fun(pars_opt = structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)

myfits_2 <- mstrust(obj_alpha, center =  structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset), studyname = "Fits", cores = 3, fits = 12, sd = 1, mypars = pars, perturbation_prediction = perturbation_prediction_0, r_kept = r_kept_2)
myfits_2 %>% as.parframe()
best_fit_2 <- myfits_2 %>% as.parframe() %>% extract() %>% as.parvec()
# exp(best_fit)
r_opt_2 <- r_alpha_fun(pars_opt = unclass(best_fit_2), pars = pars, perturbation_prediction = perturbation_prediction) %>% round(2)
r_opt_2
# pars
```

# Iteration 3: some connections are increased by the optimization procedure
Determine, which complexes have to be added where for further disentangling
```{r}
find_small_before_and_big_after_optimization(r_0,r_opt_2)
r_kept_3 <- r_kept_2 | find_small_before_and_big_after_optimization(r_0,r_opt_2)
r_kept_3

# matrix(1:20, ncol = 4)
# to_single_module_pars_opt
```


```{r}
pars_opt_subset <- paste0("log",to_single_module_pars_opt[c(1,7,12,19, # as before: to u: cux, to x: cxy and cxz, to z: czw
                                                            13,18,     # to y: cxz and czw
                                                            9,         # to z: cxy
                                                            10         # to w: cxy
                                                            )])       

r_0%>% round(2)

r_1 <-   R_fun(pars_opt = structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)

myfits_3 <- mstrust(obj_alpha, center =  structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset), studyname = "Fits", cores = 3, fits = 12, sd = 1, mypars = pars, perturbation_prediction = perturbation_prediction_0, r_kept = r_kept_3)
myfits_3 %>% as.parframe()
best_fit_3 <- myfits_3 %>% as.parframe() %>% as.parvec()
# exp(best_fit)
r_opt_3 <- r_alpha_fun(pars_opt = unclass(best_fit_3), pars = pars, perturbation_prediction = perturbation_prediction) %>% round(2)
r_opt_3
# pars

```

# Iteration 4: The algorithm terminates
No additional elements are added to be kept in the optimization procedure.
No additional free parameters are introduced by the algorithm.

```{r}
find_small_before_and_big_after_optimization(r_0,r_opt_3)
r_kept_4 <- r_kept_3 | find_small_before_and_big_after_optimization(r_0,r_opt_3)
r_kept_4

cat("Are the matrices identical?\n")
identical(r_kept_3, r_kept_4)

matrix(1:20, ncol = 4)
to_single_module_pars_opt
```

Just for curiosity, add cxy to w and czw to x
```{r}
pars_opt_subset <- paste0("log",to_single_module_pars_opt[c(1,7,12,19, # as before: to u: cux, to x: cxy and cxz, to z: czw
                                                            13,18,     # to y: cxz and czw
                                                            9,         # to z: cxy
                                                            10,         # to w: cxy
                                                            15,17)])    # to w:cxz and to to x: czw (just for fun)   

r_0%>% round(2)

r_1 <-   R_fun(pars_opt = structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)

myfits_4 <- mstrust(obj_alpha, center =  structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset), studyname = "Fits", cores = 3, fits = 12, sd = 1, mypars = pars, perturbation_prediction = perturbation_prediction_0, r_kept = r_kept_4)
myfits_4 %>% as.parframe()
best_fit_4 <- myfits_4 %>% as.parframe() %>% as.parvec()
# exp(best_fit)
r_opt_4 <- r_alpha_fun(pars_opt = unclass(best_fit_4), pars = pars, perturbation_prediction = perturbation_prediction) %>% round(2)
r_opt_4
# pars

```
