---
title: "Hub motif within a cascade"
author: "Daniel Lill"
---

# Load libraries and define some useful paths

```{r Libraries, results='hide', echo=FALSE, message=FALSE}
rm(list = ls())
library(plotly)
library(tidyverse)
library(MRAr)
library(scales)
library(xtable)
```

# Setup Model 
This is the setup of the ODE model.

Our (internal) states x are described by the odemodel which is turned into a prediction via the prediction function x():

dx/dt = f(x,p_dyn)
x = x(t, p_dyn, x_ini)

p_dyn are dynamical parameters like rate constants etc, x_ini are initial values for all states.
The prediction function xs() runs the stimulation until it reaches a steady state.


Then, these states are linked to observables y via the observation function g(). In our case the observables will be the communicating species:

y = g(x, p_obs)

p_obs are observation parameters like scaling factors.


Since all parameters are positive, it is beneficial to use log-parameters, therefore all parameters are transformed to the log-scale. These we call "outer parameters", p_outer. To feed them into our original model, we transform them back to the linear scale ("p_inner") by a parameter transformation function p(). This way, all our calculations can be made in log-space:

p_inner = p(p_outer) = p(log(p_obs), log(p_dyn), log(x_ini))


These three types of functions can be concatenated by the "*"-operator.
g(x(t, p(p_outer))) = g*x*p(t, p_outer)


## Prediction function
odemodel() writes the function f as a C-function, which is called by x and xs.
```{r Prediction function Cascade, results='asis'}
reactions <- NULL %>% 
  addReaction(from = "u1", to = "u2", rate = "((v1 *  u1)/(km1 + u1))", description = "act_u") %>% 
  addReaction(from = "u2", to = "u1", rate = "((v2 * u2 )/(km2 + u2))", description = "deact_u") %>% 
  addReaction(from = "u2 + x1", to = "Cu2x1", rate = "k91*u2*x1 - k92*Cu2x1", description = "u2x1 to comp") %>% 
  addReaction(from = "Cu2x1", to = "u2 + x2", rate = "k9cat * Cu2x1", description = "x2 formation") %>% 
  addReaction(from = "x2", to = "x1", rate = "v10 * x2 /(km10 + x2)", description = "x deactivation") %>% 
  addReaction(from = "x2 + y1", to = "Cx2y1", rate = "(k31 * y1 *x2)  -  (k32 *Cx2y1)", description = "x2_y_to_comp") %>% 
  addReaction(from = "Cx2y1", to = "x2 + y2", rate = "(k3cat * Cx2y1) * (1+ kcrossztoy  * z2 )/(1 + z2)", description = "y2_formation") %>% 
  addReaction(from = "y2", to = "y1", rate = "((v5 * y2)/(km5 + y2)) ", description = "deact_y") %>% 
  addReaction(from = "x2 + z1", to = "Cx2z1", rate = "(k61 * z1 *x2) -  (k62 *Cx2z1)", description = "x2_z_to_comp") %>% 
  addReaction(from = "Cx2z1", to = "x2 + z2", rate = "(k6cat * Cx2z1)", description = "z2_formation") %>% 
  addReaction(from = "z2", to = "z1", rate = "((v8 * z2)/(km8 + z2))", description = "deact_z") %>% 
  addReaction(from = "z2 + w2", to = "Cz2w2", rate = "k111*z2*w2 - k112*Cz2w2", description = "z2w2 to comp") %>% 
  addReaction(from = "Cz2w2", to = "z2 + w1", rate = "k11cat * Cz2w2", description = "z1 formation") %>% 
  addReaction(from = "w1", to = "w2", rate = "v12 * w1 /(km12 + w1)", description = "z activation") %>% 
  addReaction(from = "w2 + s1", to = "Cw2s1", rate = "k131* w2 * s1 - k132 * Cw2s1", description = "w2 s1 to comp") %>% 
  addReaction(from = "Cw2s1", to = "w2 + s2", rate = "k13cat * Cw2s1", description = "w2 formation") %>% 
  addReaction(from = "s2", to = "s1", rate = "v14 * s2 / (km14 + s2)", description = "w2 deactivation") %>% {.}
myodemodel <- odemodel(reactions, modelname = paste0("hub_within_cascade"), fixed = c("x1", "k1"))

# Prediction function
x <- Xs(odemodel = myodemodel)
# Prediction function for steady states
xs <- Xs_steady(myodemodel) # When computing steady states, set deriv = F
```

## Observation function
Observation function which returns our communicating species. 
The free parameters a_ij are introduced here to perform the linear combination of states which we wish to observe.
We chose to call them alpha_pars, since they will be used to optimize the objective function later on.
```{r Observation function with Cx2z2 in x-module}
modules <- c("u", "x", "y", "z", "w", "s")
complexes <- c("Cu2x1", "Cx2y1", "Cx2z1", "Cz2w2", "Cw2s1")
to_single_module_alpha_pars <- paste0("a_to_", outer(modules, complexes, paste, sep = "_")) %>% matrix(ncol = 5)
to_single_module <- prodSymb(to_single_module_alpha_pars, t(t(complexes)))
free_species <- c("u2","x2","y2","z2","w2", "s2")
obs <- paste(free_species, to_single_module, sep = " + ")
g <- Y(as.eqnvec(structure(obs, names = modules)), reactions, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```

## Inner parameters
Load all parameters and format them for further use.
```{r Parameters Cascade}
pars_inner <- c(u1 = 100, u2 = 0, Cu2x1 = 0, x1 = 100, x2 = 0, Cx2y1 = 0, y1 = 100, 
y2 = 0, Cx2z1 = 0, z1 = 100, z2 = 0, Cz2w2 = 0, w1 = 100, w2 = 0, 
Cw2s1 = 0, s1 = 100, s2 = 0, km1 = 100, v1 = 1, km2 = 100, v2 = 1, 
k31 = 0.01, k32 = 0.4, k3cat = 0.05, km5 = 100, v5 = 5, k61 = 0.05, 
k62 = 0.4, k6cat = 0.008, km8 = 100, v8 = 1, k91 = 0.01, k92 = 0.5, 
k9cat = 0.01, km10 = 100, v10 = 1.5, k111 = 0.01, k112 = 0.4, 
k11cat = 0.05, km12 = 100, v12 = 1.5, k131 = 0.01, k132 = 0.4, 
k13cat = 0.05, km14 = 100, v14 = 1.5, kcrossztoy = 1, a_to_u_Cu2x1 = 0, 
a_to_u_Cx2y1 = 0, a_to_u_Cx2z1 = 0, a_to_u_Cz2w2 = 0, a_to_u_Cw2s1 = 0, 
a_to_x_Cu2x1 = 0, a_to_x_Cx2y1 = 0, a_to_x_Cx2z1 = 0, a_to_x_Cz2w2 = 0, 
a_to_x_Cw2s1 = 0, a_to_y_Cu2x1 = 0, a_to_y_Cx2y1 = 0, a_to_y_Cx2z1 = 0, 
a_to_y_Cz2w2 = 0, a_to_y_Cw2s1 = 0, a_to_z_Cu2x1 = 0, a_to_z_Cx2y1 = 0, 
a_to_z_Cx2z1 = 0, a_to_z_Cz2w2 = 0, a_to_z_Cw2s1 = 0, a_to_w_Cu2x1 = 0, 
a_to_w_Cx2y1 = 0, a_to_w_Cx2z1 = 0, a_to_w_Cz2w2 = 0, a_to_w_Cw2s1 = 0, 
a_to_s_Cu2x1 = 0, a_to_s_Cx2y1 = 0, a_to_s_Cx2z1 = 0, a_to_s_Cz2w2 = 0, 
a_to_s_Cw2s1 = 0)
```


## Parameter transformation function
```{r}
# Logtrafo of all pars
logtrafo <- structure(paste0("exp( log", names(pars_inner), ")"), 
                             names = names(pars_inner))
# initial values
logtrafo[c("x2", "Cx2y1", "y2", "Cx2z1", "z2", "u2", "Cu2x1", "w2", "Cz2w2", "s2", "Cw2s1")] <- 0

p_log <- P(trafo = logtrafo,
           compile = TRUE,
           modelname = "p_log")
   
pars <- structure(log(pars_inner+.Machine$double.eps), names = paste0("log", names(pars_inner))) 
alpha_pars <- pars[str_subset(names(pars), "loga")]
```


# Export 

## Tables
Print parameters
```{r}
pars_inner %>% 
  .[!str_detect(names(.), "^a")] %>% 
  print2word_namedvec("pars_hub_within_cascade", names_in_mathenv = T)
```

Print ODEs
```{r}
reactions %>% 
  as.eqnvec() %>% 
  print2word_eqnvec("odes_hub_within_cascade")
```

## Matrices
```{r}
which_pars_perturbed <- c("logu1", "logx1", "logy1", "logz1", "logw1", "logs1")
which_alpha_pars <- names(alpha_pars) %>% str_subset("to_(.)_C\\1")
SI_matrices <- algo(pars, which_pars_perturbed, which_alpha_pars, alpha_pars, 0, g, xs, p_log)

SI_matrices[-4] %>% unlist(F) %>% print2word_matrixlist2("matrices_hub_within_cascade")
```



## Clean up Workdir
```{r}
unlink(paste0("*.", c("c", "o", "so", "tex", "lnk")))
unlink("Fits/", T)
unlink(".pars_controlled/", T)
```
