---
title: "Hub motif within a cascade"
author: "Daniel Lill"
---

# Load libraries and define some useful paths

```{r Libraries, results='hide', echo=FALSE, message=FALSE}
rm(list = ls())
library(plotly)
library(tidyverse)
library(MRAr)
library(scales)
library(xtable)
```

# Setup Model 
This is the setup of the ODE model.

Our (internal) states x are described by the odemodel which is turned into a prediction via the prediction function x():

dx/dt = f(x,p_dyn)
x = x(t, p_dyn, x_ini)

p_dyn are dynamical parameters like rate constants etc, x_ini are initial values for all states.
The prediction function xs() runs the stimulation until it reaches a steady state.


Then, these states are linked to observables y via the observation function g(). In our case the observables will be the communicating species:

y = g(x, p_obs)

p_obs are observation parameters like scaling factors.


Since all parameters are positive, it is beneficial to use log-parameters, therefore all parameters are transformed to the log-scale. These we call "outer parameters", p_outer. To feed them into our original model, we transform them back to the linear scale ("p_inner") by a parameter transformation function p(). This way, all our calculations can be made in log-space:

p_inner = p(p_outer) = p(log(p_obs), log(p_dyn), log(x_ini))


These three types of functions can be concatenated by the "*"-operator.
g(x(t, p(p_outer))) = g*x*p(t, p_outer)


## Prediction function
odemodel() writes the function f as a C-function, which is called by x and xs.
```{r Prediction function Cascade, results='asis'}
modelpath <- paste0("Models/HubInCascade/")
modelname   <- "HubInCascade"
mymodel     <- as.eqnlist(read.csv(paste0(modelpath, modelname, ".csv")))

# myodemodel <- odemodel(mymodel, modelname = paste0(modelname), deriv = T)
# save(myodemodel, file = paste0(modelpath, "odemodel",modelname, ".RData"))
load(paste0(modelpath, "odemodel",modelname, ".RData"))

# Prediction function
x <- Xs(odemodel = myodemodel)
# Prediction function for steady states
xs <- Xs_steady(myodemodel) # When computing steady states, set deriv = F
# print(mymodel)
# mymodel %>% as.eqnvec() %>% print(pander = T)

# write.csv(myodemodel[[1]] %>% attr("eq") %>% data.frame(), "~/Promotion/Writing/Papers/2017 02 MRA optimization procedure Paper/pictures/SI/Hub in Cascade model", quote = F)
```

## Observation function
Observation function which returns our communicating species. 
The free parameters a_ij are introduced here to perform the linear combination of states which we wish to observe.
We chose to call them pars_opt, since they will be used to optimize the objective function later on.
```{r Observation function with Cx2z2 in x-module}
modules <- modules0 <- c("u", "x", "y", "z", "w", "s")
complexes <- c("Cu2x1", "Cx2y1", "Cx2z1", "Cz2w2", "Cw2s1")

to_single_module_pars_opt <- paste0("a_to_", outer(modules, complexes, paste, sep = "_")) %>% matrix(ncol = 5)

to_single_module <- prodSymb(to_single_module_pars_opt, t(t(complexes)))
# modules_grid <- outer(modules,modules, paste0)
# modules_grid[upper.tri(modules_grid)] <- t(modules_grid)[upper.tri(modules_grid)]
# diag(modules_grid) <- "0"
# to_both_modules <- paste0("a_to_", outer(modules_grid, complexes, paste, sep = "_")) %>% matrix(ncol = 4)

free_species <- c("u2","x2","y2","z2","w2", "s2")


obs <- obs0 <- paste(free_species, to_single_module, sep = " + ")
  
  
g <-  g0 <- Y(as.eqnvec(structure(obs0, names = modules0)), mymodel, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```

## Inner parameters
Load all parameters and format them for further use.
```{r Parameters Cascade}

pars_raw <- read.csv(paste0(modelpath, "pars.csv"), header = FALSE)  # dynamical parameters
pars_raw <- structure(pars_raw$V2, names = as.character(pars_raw$V1))
ic_raw <- read.csv(paste0(modelpath, "IC.csv"), header = F)          # initial conditions
ic_raw <- structure(ic_raw$V2, names = as.character(ic_raw$V1))
# sort ics like they appear in x, so it can be used in stode()
vars_x <- attr(x, "parameters")[attr(x, "parameters") %in% names(ic_raw)]
ic_raw <- ic_raw[vars_x]

pars_inner_opt_0 <- setdiff(attr(g, "parameters"),names(c(pars_raw,ic_raw)))
pars_inner_opt_0 <- structure(rep(0, length(pars_inner_opt_0)), 
                              names = pars_inner_opt_0)

pars_inner_0 <- pars_inner <- c(ic_raw, pars_raw, pars_inner_opt_0)


# check if all parameters are assigned a value. Both lines should evaluate to the same number.
# x_pars <- unique(union((g) %>% attr("parameters"),(xs) %>% attr("parameters")))
# names(pars_inner) %in% x_pars %>% sum()
# x_pars %in% names(pars_inner) %>% sum()
```


## Parameter transformation function
```{r}
# Logtrafo of all pars
logtrafo <- structure(paste0("exp( log", names(pars_inner_0), ")"), 
                             names = names(pars_inner_0))


# initial values
logtrafo[c("x2", "Cx2y1", "y2", "Cx2z1", "z2", "u2", "Cu2x1", "w2", "Cz2w2", "s2", "Cw2s1")] <- 0


p_log <- P(trafo = logtrafo,
           compile = TRUE,
           modelname = "p_log")
           

pars_0 <- pars <- structure(log(pars_inner_0+.Machine$double.eps), names = paste0("log", names(pars_inner_0))) 

pars_opt_0 <- pars_opt <- structure(log(pars_inner_opt_0+.Machine$double.eps), names = paste0("log", names(pars_inner_opt_0)))

# pars_opt_0 <- pars_opt <- pars_opt_0[c("loga_1", "loga_2")]
# pars_0
# p_log(pars_0, deriv = F)
```

# Simulate perturbation data

For the perturbation experiments we use another parameter transformation to scale some parameters.
p_pert_fun automatically generates a parameter transformation function, which returns different sets of parameters for different conditions and a control condition.
Eg. if we have pars_perturbed = c("logx1" = log(0.9)), p_pert_fun(pars_perturbed) will return a parameter transformation with the control condition with unchanged parameters and one condition in which logx' = logx + log(0.9)

"log(0.9)" corresponds to a ~10% reduction of the original parameter value.
```{r}
# Which pars shall be perturbed?
pars_perturbed_0 <- pars_perturbed <- c("logu1" = log(0.9), "logx1" = log(0.9), "logy1" = log(0.9), "logz1" = log(0.9), "logw1" = log(0.9), "logs1" = log(0.9))
# pars_perturbed_0 <- pars_perturbed <- c("logx1" = log(0.8), "logy1" = log(0.8), "logz1" = log(0.8))



# pars_perturbed
p_pert <- p_pert_fun(pars_perturbed = pars_perturbed, pars = pars)

```


```{r}
find_symmetric_elements_both_nonzero <- function(mat) {
  myupper <- mat[upper.tri(mat)]
  mylower <- t(mat)[upper.tri(mat)]

  out <- matrix(FALSE, nrow = nrow(mat), ncol = ncol(mat))
  out[upper.tri(out)] <- (abs(myupper) > 0.01 & abs(mylower) > 0.01)
  out <- out + t(out)
  return(out)
}

r_kept_fun <- function(pars_opt = pars_opt_0,
                      perturbation_prediction = perturbation_prediction_0,
                      obs_fun = g0,
                      p_fun = (p_log*p_pert),
                      pars = pars_0,
                      alpha = -log(0+.Machine$double.eps)) {

  r_0 <- R_fun(pars_opt = pars_opt,
               perturbation_prediction = perturbation_prediction,
               obs_fun = obs_fun,
               p_fun = p_fun,
               pars = pars) %>% local_response_matrix_eq10()

  r_1 <- R_fun(pars_opt = pars_opt + alpha, #yields asymptotically the same results as only complexes without free species.
               perturbation_prediction = perturbation_prediction,
               obs_fun = obs_fun,
               p_fun = p_fun,
               pars = pars) %>% local_response_matrix_eq10()

  # Which elements are kept in the optimization process?
  
  # sign switch
  signs <- (sign(r_0) - sign(r_1)) %>% as.logical() %>% matrix(nrow = nrow(r_0))
  
  # small elements
  # smalls <- abs(r_0) < 1e-3
  
  # corresponding symmetric element is non-zero
  syms <- find_symmetric_elements_both_nonzero(r_0)
  
  keep <- (signs & syms) #| smalls

  return(keep)
}

find_small_before_and_big_after_optimization <- function(r_0, r_opt) {
  small_before <- abs(r_0) < 1e-2
  big_after <- abs(r_opt) > 1e-2
  return(small_before & big_after)
}


```


## Final parameter values
```{r}
pars <- c(100,0,0,       100,0,0,      100,0,0,      100,0,0,      100,0,0,    100,0,
          100,1,100,1,     0.01,0.4,0.05, #km1 - k3cat
          100,5,            0.05,0.4,0.008, #km5 - k6cat
          100,1, # km8 v8
          0.01,0.5,0.01,    100,1.5, #k91 - v10
          0.01,0.4,0.05,    100,1.5, #k111 - v12
          0.01,0.4,0.05,    100,1.5, #k131 - v14
          1, # kcrossztoy 
          rep(0, length(pars_opt_0))) %>%
  set_names(names(pars_inner_0)) # as in cascade
pars %>% data.frame()

# write.csv(pars %>% data.frame(), "~/Promotion/Writing/Papers/2017 02 MRA optimization procedure Paper/pictures/SI/Hub in Cascade parameters", quote = F)
# pars <- c(100,0,0,       100,0,0,      100,0,0,      100,0,0,      100,0,0,    100,0, 
#           100,1,100,1,     0.02,0.3,0.05, #km1 - k3cat   x to y
#           100,5,            0.01,0.4,0.008, #km5 - k6cat  x to z 
#           100,1, # km8 v8
#           0.005,0.5,0.01,    100,1.5, #k91 - v10   u to x
#           0.005,0.4,0.05,    100,1.5, #k111 - v12
#           0.005,0.4,0.05,    100,1.5, #k131 - v14
#           0, # kcrossztoy 
#           rep(0, length(pars_opt_0))) %>%
#   set_names(names(pars_inner_0)) # as in cascade

pars_0 <- pars <- log(pars+.Machine$double.eps) %>% set_names(names(pars_0))# large complex concentrations

```

## Generate the perturbation data
```{r}
perturbation_prediction_0 <- perturbation_prediction <-  (xs*p_log*p_pert)(times = c(0,Inf), pars = pars, deriv = F)
perturbation_prediction %>% lapply(round,1)

(x*p_log*p_pert)(times = seq(0,20000, length.out = 50), pars = pars) %>%
  plotPrediction()
```




## 1st measurement: Detect that there is sequestration
```{r}

# perturbation_prediction <- lapply(perturbation_prediction_0, function(i)  exp(log(i) + rnorm(length(i), sd = 0.005)))
# perturbation_prediction %>% lapply(round,0)# %>% lapply(log)

r_0 <-   R_fun(pars_opt = pars_opt_0,
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_0%>% round(1)

find_symmetric_elements_both_nonzero(r_0)
```

## Iteration 1: Complexes are measured, add them to communicating species
It is assumed to be known, which species are active in the complexes.
Therefore, the complexes can be added only to the regulating modules.
Subsequently, make the first iteration of optimization
```{r}
pars_opt_subset <- paste0("log",to_single_module_pars_opt[c(1,8,14,22,29)])
# pars_opt_subset <- paste0("log",to_single_module_pars_opt[c(1,14)])
pars_opt_subset

r_kept_1 <- r_kept_fun(pars_opt = pars_opt[pars_opt_subset],
                       perturbation_prediction = perturbation_prediction, 
                       obs_fun = g,
                       p_fun = (p_log * p_pert), 
                       pars = pars,
                       alpha = -log(0+.Machine$double.eps))
r_kept_1


# r_kept_1 <- matrix(F, nrow = 6, ncol = 6)
# r_kept_1[2,4] <- T
# r_kept_1[1,2] <- T

r_0%>% round(2)

r_1 <-   R_fun(pars_opt = structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)

myfits_1 <- mstrust(obj_alpha, center =  structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset), studyname = "Fits", cores = 3, fits = 6, sd = 1, mypars = pars, perturbation_prediction = perturbation_prediction_0, r_kept = r_kept_1)
myfits_1 %>% as.parframe()
best_fit_1 <- myfits_1 %>% as.parframe() %>% as.parvec()
# exp(best_fit)
r_opt_1 <- r_alpha_fun(pars_opt = unclass(best_fit_1), pars = pars, perturbation_prediction = perturbation_prediction) 
r_opt_1 %>% round(1)
# pars
```


## Iteration 2: some connections are increased by the optimization procedure
Determine, which complexes have to be added where for further disentangling
```{r}

find_small_before_and_big_after_optimization(r_0,r_opt_1)
r_kept_2 <- r_kept_1 | find_small_before_and_big_after_optimization(r_0,r_opt_1)

r_kept_2
matrix(1:30, ncol = 5)
to_single_module_pars_opt
```


```{r}
pars_opt_subset <- paste0("log",to_single_module_pars_opt[c(1,8,14,22,29 # as before
                                                            ,15 # to y: cxz
                                                            ,21     #to y: czw
                                                            ,10      # to z: cxy
                                                            # 3         # to y: cux
                                                            )])

r_0%>% round(2)

r_1 <-   R_fun(pars_opt = structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)

myfits_2 <- mstrust(obj_alpha, center =  structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset), studyname = "Fits", cores = 3, fits = 12, sd = 1, mypars = pars, perturbation_prediction = perturbation_prediction_0, r_kept = r_kept_2)
myfits_2 %>% as.parframe()
best_fit_2 <- myfits_2 %>% as.parframe() %>% extract() %>% as.parvec()
# exp(best_fit)
r_opt_2 <- r_alpha_fun(pars_opt = unclass(best_fit_2), pars = pars, perturbation_prediction = perturbation_prediction) %>% round(2)
r_opt_2
# pars
```

## Iteration 3: some connections are increased by the optimization procedure
Determine, which complexes have to be added where for further disentangling
```{r}
find_small_before_and_big_after_optimization(r_0,r_opt_2)
r_kept_3 <- r_kept_2 | find_small_before_and_big_after_optimization(r_0,r_opt_2)
r_kept_3

matrix(1:20, ncol = 4)
to_single_module_pars_opt
```


```{r}
pars_opt_subset <- paste0("log",to_single_module_pars_opt[c(1,8,14,22,29 # as before
                                                            # ,15 # to y: cxz
                                                            ,21     #to y: czw
                                                            ,10      # to z: cxy
                                                            # 3         # to y: cux
                                                            ,9         # to w cxy
                                                            )])       

r_0%>% round(2)

r_1 <-   R_fun(pars_opt = structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)

myfits_3 <- mstrust(obj_alpha, center =  structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset), studyname = "Fits", cores = 3, fits = 12, sd = 1, mypars = pars, perturbation_prediction = perturbation_prediction_0, r_kept = r_kept_3)
myfits_3 %>% as.parframe()
best_fit_3 <- myfits_3 %>% as.parframe() %>% as.parvec()
# exp(best_fit)
r_opt_3 <- r_alpha_fun(pars_opt = unclass(best_fit_3), pars = pars, perturbation_prediction = perturbation_prediction) %>% round(2)
r_opt_3
# pars

```

## Iteration 4: The algorithm terminates
No additional elements are added to be kept in the optimization procedure.
No additional free parameters are introduced by the algorithm.

```{r}
find_small_before_and_big_after_optimization(r_0,r_opt_3)
r_kept_4 <- r_kept_3 | find_small_before_and_big_after_optimization(r_0,r_opt_3)
r_kept_4

cat("Are the matrices identical?\n")
identical(r_kept_3, r_kept_4)

matrix(1:20, ncol = 4)
to_single_module_pars_opt
```

Just for curiosity, add cxy to w and czw to x
```{r}
pars_opt_subset <- paste0("log",to_single_module_pars_opt[c(1,7,12,19, # as before: to u: cux, to x: cxy and cxz, to z: czw
                                                            13,18,     # to y: cxz and czw
                                                            9,         # to z: cxy
                                                            10,         # to w: cxy
                                                            15,17)])    # to w:cxz and to to x: czw (just for fun)   

r_0%>% round(2)

r_1 <-   R_fun(pars_opt = structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)

myfits_4 <- mstrust(obj_alpha, center =  structure(rep(0, length(pars_opt_subset)), names = pars_opt_subset), studyname = "Fits", cores = 3, fits = 12, sd = 1, mypars = pars, perturbation_prediction = perturbation_prediction_0, r_kept = r_kept_4)
myfits_4 %>% as.parframe()
best_fit_4 <- myfits_4 %>% as.parframe() %>% as.parvec()
# exp(best_fit)
r_opt_4 <- r_alpha_fun(pars_opt = unclass(best_fit_4), pars = pars, perturbation_prediction = perturbation_prediction) %>% round(2)
r_opt_4
# pars

```
This doesn't help anymore.