---
title: "Orthogonality Condition"
author: "Daniel Lill"
output: 
  pdf_document:
    keep_tex: true
---

# Load libraries and define some useful paths

```{r Libraries, results='hide', echo=FALSE, message=FALSE}
library(MRAr)
library(scales)
library(xtable)
path <- paste0(getwd(), "/")
pathout <- paste0(path, "Output/")
```

```{r Names for Fit storage of different initial Conditions}
fit_name <- "Orthogonality_condition"
```

```{r Call render without overwriting any previous documents, echo=FALSE, eval=FALSE, message=FALSE, include=FALSE}

rmarkdown::render("Orthogonality condition.Rmd", output_file = paste0(pathout, format(Sys.time(), "%Y%m%d-%H%M")
, " - " , fit_name, ".pdf"))
beepr::beep("~/Promotion/Software/rimshot.mp3")
```

# Prabakaran-Model

### Prediction function.
odemodel writes a C-function, which is executed in x and xs.
x is a regular prediction function while xs only finds and returns the steady states.
If the odemodel has been calculated already, it can be loaded with "load".
```{r Prediction function}
modelpath <- paste0(path, "Models/Prabakaran/")
modelname   <- "Prabakaran"
mymodel     <- as.eqnlist(read.csv(paste0(modelpath, modelname, ".csv")))
# mymodel


modelname   <- "Prabakaranwithxy"
# myodemodel <- odemodel(mymodel, modelname = modelname, deriv = T)
# save(myodemodel, file = paste0(modelpath, "odemodel",modelname, ".RData"))
load(paste0(modelpath, "odemodel",modelname, ".RData"))


# Prediction function
x <- Xs(odemodel = myodemodel)
# Prediction function for steady states
xs <- Xs_steady(myodemodel)
print(mymodel, pander = T)
```

### Modules, their observables and the observation function. {#modulesandobsfuns}
Observation function which depends also on the optimization parameters. With their help, the complexes can smoothly be switched on in the observation.

This is the naive modularization used in the Prabakaran Paper, desregarding, which species are actually the communicating ones. This is fair enough, because usually the "active" forms are consedered to be the communicating species in a non-parametric network reconstruction.
```{r Naive observation function}
modules <- modules0 <- c("m2P","e2P")

obs <- obs0 <- c("m2+a_1*Cm2e1",
                 "e2+Ce2p1")
g <- g0 <- Y(as.eqnvec(structure(obs0, names = modules0)), mymodel, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```



### Parameters.
Load all parameters and format them for further use.
```{r Parameters}

pars_raw <- read.csv(paste0(modelpath, "pars.csv"), header = FALSE)  # dynamical parameters
pars_raw <- structure(pars_raw$V2, names = as.character(pars_raw$V1))
ic_raw <- read.csv(paste0(modelpath, "IC.csv"), header = F)          # initial conditions
ic_raw <- structure(ic_raw$V2, names = as.character(ic_raw$V1))
# sort ics like they appear in x, so it can be used in stode()
vars_x <- attr(x, "parameters")[attr(x, "parameters") %in% names(ic_raw)]
ic_raw <- ic_raw[vars_x]

pars_opt_0 <- setdiff((g) %>% attr("parameters"),names(c(pars_raw,ic_raw)))
pars_opt_0 <- structure(rep(0, length(pars_opt_0)), names = pars_opt_0)

pars_0 <- pars <- c(ic_raw, pars_raw, pars_opt_0)


# check if all parameters are assigned a value. Both lines should evaluate to the same number.
x_pars <- unique(union((g) %>% attr("parameters"),(xs) %>% attr("parameters")))
names(pars) %in% x_pars %>% sum()
x_pars %in% names(pars) %>% sum()

```


### Parameter trafos for convenient perturbation data.
Set up the perturbations
```{r which pars are perturbed}
# Which pars shall be perturbed?
pars_perturbed_0 <- pars_perturbed <- c(m1 = 1.1, e1 = 1.1)

p_pert <- p_pert_fun(pars_perturbed = pars_perturbed_0,
                     pars = pars_0)
```

### Perturbation data generation and elements to keep.
Just to show the code...
```{r Generate perturbation data}
perturbation_prediction_0 <-  (xs*p_pert)(times = c(0,Inf), pars = pars_0, deriv = F)
r_kept_0 <- r_kept_fun(obs_fun = g0, p_fun = p_pert)

```


# System reduction
This part I am doing in Mathematica.
I use "Analytics with complexes in both modules_OR_2-1.nb"

Ugly hardcoded shit!
```{r}
dfxde <- "(k2*(e2P*k31 - etot*k31 - k32 - k4 + k31*m2P + sqrt(4*(e2P - etot)*k31^2*m2P + (k32 + k4 + k31*(-e2P + etot + m2P))^2)))/(2*sqrt(4*(e2P - etot)*k31^2*m2P + (k32 + k4 + k31*(-e2P + etot + m2P))^2))"
mypars <- getSymbols(dfxde)
mypars <- c(pars, etot= unname(pars["e1"]))[mypars]
mypars <- mypars[!is.na(mypars)]

myfun <- funC0(dfxde, parameters = names(mypars))

```

```{r}
pars <- pars_0
pars["a_1"] <- 1
myss <- (g*xs)(c(0,Inf), pars)[[1]]
complexes <- (xs)(c(0,Inf), pars)[[1]]
myfun(m2P = unname(myss[,"m2P"]), 
      e2P = unname(myss[,"e2P"]), 
      k2 = unname(pars["k2"]), 
      k31 =unname(pars["k31"]), 
      k32 = unname(pars["k32"]), 
      k4 = unname(pars["k4"]), 
      etot = unname(pars["e1"]))
```


# Dose Response function.
Scan through a parameter (In this case k4) and get the derivative, the steady state concentrations and the local response matrix as a data.frame.

```{r}

mydoseresponse <- lapply(seq(-2,2, by = .01), function(i) {
  pars <- pars_0
  pars["a_1"] <- 1
  pars["k31"] <- pars["k31"]*10^i
  pars["k4"] <- pars["k4"]*10^(-i)
  
  myss <- (g*xs)(c(0,Inf), pars, deriv = F)[[1]]
  complexes <- (xs)(c(0,Inf), pars, deriv = F)[[1]]
  d <- myfun(m2P = unname(myss[,"m2P"]), 
        e2P = unname(myss[,"e2P"]), 
        k2 = unname(pars["k2"]), 
        k31 =unname(pars["k31"]), 
        k32 = unname(pars["k32"]), 
        k4 = unname(pars["k4"]), 
        etot = unname(pars["e1"]))
  
  
  perturbation_prediction <-  (xs*p_pert)(times = c(0,Inf), pars = pars, deriv = F)
  r <- R_fun(pars_opt = c(a_1 = 1), 
             perturbation_prediction = perturbation_prediction, 
             obs_fun = g, p_fun = p_pert, pars = pars) %>% local_response_matrix_eq10()
  dim(r) <- NULL
  names(r) <-  paste0("r" ,outer(as.character(1:2),as.character(1:2), paste0))
  
  return(data.frame(m2P = myss[,"m2P"], 
                    e2P = myss[,"e2P"], 
                    complexes, 
                    deriv = d, 
                    par_dose = 10^i, 
                    t(r)))
}) %>% do.call(rbind,.)

# mydoseresponse
```

# Plotting of the dose responses

## Parameter values
```{r}
pars_0[c("k31", "k32", "k4")]
```


## Plot against the ratio of cme/m2
As expected, r12 is smaller the smaller the ratio.
But on the other hand, the derivative dfxdetot is not.
```{r}
library(tidyverse)
theme_set(theme_dMod())
mydoseresponse %>% 
  mutate(log10_ratio_complex_m2 = log10(Cm2e1/m2)) %>% 
  filter(log10_ratio_complex_m2 < 0, log10_ratio_complex_m2 >-2) %>% 
  mutate(x = m2P, y = e2P, dfxdetot = deriv) %>% 
  gather(key = key,value =value , dfxdetot, r21,r12,Cm2e1) %>%  
  ggplot(aes(x = log10_ratio_complex_m2, y = log10(value))) +
  geom_line() +
  facet_wrap(~key, scales = "free")

```

## Plot against fold changes of the parameter scanned through
```{r}
mydoseresponse %>% 
  mutate(log10_ratio_complex_m2 = log10(Cm2e1/m2P)) %>% 
  # mutate(deriv = log10(deriv)) %>%
  mutate(x = m2P, y = e2P, dfxdetot = deriv) %>% 
  gather(key = key,value =value , dfxdetot, r21,r12, Cm2e1, log10_ratio_complex_m2) %>% 
  ggplot(aes(x = log10(par_dose), y = value)) +
  geom_line() +
  facet_wrap(~key, scales = "free")
```

```{r}
mydoseresponse %>% 
  mutate(log10_ratio_complex_m2 = log10(Cm2e1/m2P)) %>% 
  # mutate(deriv = log10(deriv)) %>%
  # gather(key = key,value =value , deriv, r11,r21,r12,r22) %>% 
  ggplot(aes(x = deriv, y = r12)) +
  geom_path() 
  # facet_wrap(~key, scales = "free")
```

```{r}
steady <- xs(c(0,Inf), pars_0)

jacobian_names <- outer(paste0(names(steady[[1]][1,-1]), "."), names(steady[[1]][1,-1]), paste0) %>% as.character()

jacobian_equations <- myodemodel$func %>% 
  attr("eq") %>% 
  jacobianSymb(variables = myodemodel$func %>% attr("var"))

JI <- as.numeric(jacobian_equations)
JI[is.na(JI)] <- 1
JI <- JI %>% matrix(ncol = sqrt(length(JI))) %>% set_colnames(myodemodel$func %>% attr("var")) %>% set_rownames(myodemodel$func %>% attr("var"))
"bla"
JI
JIR <- (JI & t(JI)) %>% as.numeric() %>% matrix(ncol = sqrt(length(JI))) %>% set_colnames(myodemodel$func %>% attr("var"))
JIR
```

```{r}
library("ggnet")
net <- network::network.adjacency(JIR)
ggnet2(net)

net = rgraph(10, mode = "graph", tprob = 0.5)
net = network(net, directed = FALSE)
ggnet2(net)

```

