---
title: "Prabakaran-Model"
author: "Daniel Lill"
output: 
  pdf_document:
    keep_tex: true
---

# Load libraries and define some useful paths

```{r Libraries, results='hide', echo=FALSE, message=FALSE}
library(MRAr)
library(scales)
library(xtable)
path <- paste0(getwd(), "/")
pathout <- paste0(path, "Output/")
```

```{r Names for Fit storage of different initial Conditions}
fit_name <- "Prabakaran"
```

```{r Call render without overwriting any previous documents, echo=FALSE, eval=FALSE, message=FALSE, include=FALSE}

rmarkdown::render("Prabakaran.Rmd", output_file = paste0(pathout, format(Sys.time(), "%Y%m%d-%H%M")
, " - " , fit_name, ".pdf"))
beepr::beep("~/Promotion/Software/rimshot.mp3")
```

# Prabakaran-Model

### Prediction function.
odemodel writes a C-function, which is executed in x and xs.
x is a regular prediction function while xs only finds and returns the steady states.
If the odemodel has been calculated already, it can be loaded with "load".
```{r Prediction function}
modelpath <- paste0(path, "Models/Prabakaran/")
modelname   <- "Prabakaran"
mymodel     <- as.eqnlist(read.csv(paste0(modelpath, modelname, ".csv")))
# mymodel

# myodemodel <- odemodel(mymodel, modelname = modelname, deriv = T)
# save(myodemodel, file = paste0(modelpath, "odemodel",modelname, ".RData"))
load(paste0(modelpath, "odemodel",modelname, ".RData"))


# Prediction function
x <- Xs(odemodel = myodemodel)
# Prediction function for steady states
xs <- Xs_steady(myodemodel)
print(mymodel)
```

### Modules, their observables and the observation function. {#modulesandobsfuns}
Observation function which depends also on the optimization parameters. With their help, the complexes can smoothly be switched on in the observation.

This is the naive modularization used in the Prabakaran Paper, desregarding, which species are actually the communicating ones. This is fair enough, because usually the "active" forms are consedered to be the communicating species in a non-parametric network reconstruction.
```{r Naive observation function}
modules <- modules0 <- c("m2","e2tot")

obs <- obs0 <- c("m2+a_1*Cm2e1",
                 "e2+Ce2p1")
g <- g0 <- Y(as.eqnvec(structure(obs0, names = modules0)), mymodel, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```

Better modularization (also, set some defaults.)
```{r good observation function}
modules1 <- c("Cm2e1","e1")

obs1 <- c(  "a_1*m2+Cm2e1",
                   "e1")
g1 <- Y(as.eqnvec(structure(obs1, names = modules1)), mymodel, compile = TRUE, modelname = "obsfn1", attach.input = FALSE)
```


```{r Observation function with better modularization in m-module, but not in e-module.}
modules2 <- c("Cm2e1","e2")

obs2 <- c(  "a_1*m2+Cm2e1",
                   "e2+Ce2p1")
g2 <- Y(as.eqnvec(structure(obs2, names = modules2)), mymodel, compile = TRUE, modelname = "obsfn2", attach.input = FALSE)
```


### Parameters.
Load all parameters and format them for further use.
```{r Parameters}

pars_raw <- read.csv(paste0(modelpath, "pars.csv"), header = FALSE)  # dynamical parameters
pars_raw <- structure(pars_raw$V2, names = as.character(pars_raw$V1))
ic_raw <- read.csv(paste0(modelpath, "IC.csv"), header = F)          # initial conditions
ic_raw <- structure(ic_raw$V2, names = as.character(ic_raw$V1))
# sort ics like they appear in x, so it can be used in stode()
vars_x <- attr(x, "parameters")[attr(x, "parameters") %in% names(ic_raw)]
ic_raw <- ic_raw[vars_x]

pars_opt_0 <- setdiff((g) %>% attr("parameters"),names(c(pars_raw,ic_raw)))
pars_opt_0 <- structure(rep(0, length(pars_opt_0)), names = pars_opt_0)

pars_0 <- pars <- c(ic_raw, pars_raw, pars_opt_0)


# check if all parameters are assigned a value. Both lines should evaluate to the same number.
x_pars <- unique(union((g) %>% attr("parameters"),(xs) %>% attr("parameters")))
names(pars) %in% x_pars %>% sum()
x_pars %in% names(pars) %>% sum()

```


## Example code
Now the module can be simulated to generate perturbation data.

### Parameter trafos for convenient perturbation data.
Set up the perturbations
```{r which pars are perturbed}
# Which pars shall be perturbed?
pars_perturbed_0 <- pars_perturbed <- c(m1 = 1.1, e1 = 1.1)

p_pert <- p_pert_fun(pars_perturbed = pars_perturbed_0,
                     pars = pars_0)
```

### Perturbation data generation and elements to keep.
Just to show the code...
```{r Generate perturbation data}
perturbation_prediction_0 <-  (xs*p_pert)(times = c(0,Inf), pars = pars_0, deriv = F)
r_kept_0 <- r_kept_fun(obs_fun = g0, p_fun = p_pert)

```

```{r}
rbenchmark::benchmark({r_alpha_fun(pars_opt = pars_opt_0+runif(length(pars_opt_0),0,10))}, replications = 100)[,-1]
```



### Fitting.
Just to show the code...
```{r Fit, eval=FALSE}
bla <- sapply(seq(-3,3, by = 0.5), function(k1) {
  pars <- pars_0
  pars["k1"] <- pars["k1"]*exp(k1)
perturbation_prediction <-  (xs*p_pert)(times = c(0,Inf), pars = pars, deriv = F)

myfits <- mstrust(obj_alpha, 
                  center =  pars_opt_0+0.5, 
                  studyname = "Fits", 
                  cores = 3, 
                  fits = 3, 
                  sd = 0.3, 
                  obs_fun = g0,
                  p_fun = p_pert,
                  perturbation_prediction = perturbation_prediction)
# myfits %>% as.parframe() %>% plotValues()
c(myfits %>% as.parframe() %>% as.parvec(), pars["k1"]/(pars["k2"]+pars["k1"]))
})
bla
# pars["k1"] <- 
# perturbation_prediction_0 <-  (xs*p_pert)(times = c(0,Inf), pars = pars_0, deriv = F)
# 
# myfits <- mstrust(obj_alpha, 
#                   center =  pars_opt_0+0.5, 
#                   studyname = "Fits", 
#                   cores = 3, 
#                   fits = 3, 
#                   sd = 0.3, 
#                   obs_fun = g0,
#                   p_fun = p_pert)
# # myfits %>% as.parframe() %>% plotValues()
# myfits %>% as.parframe() %>% as.parvec()
# best_fit <- best_fit[best_fit$a_1>0, ] %>% as.parvec()
```

## Explore the effect of the alpha when choosing elements to keep
```{r Effect of alpha on r_kept, results='asis', echo=FALSE}
r_alpha <-function(pars_opt = pars_opt_0,
                      perturbation_prediction = perturbation_prediction_0,
                      obs_fun = g0,
                      p_fun = p_pert,
                      pars = pars_0) {
  # set pars_opt to zero
  sapply(pars_opt, function(i) {
    names(i) <- names(pars_opt_0)
    R_fun(pars_opt = i,
               perturbation_prediction = perturbation_prediction,
               obs_fun = obs_fun,
               p_fun = p_fun,
               pars = pars) %>% local_response_matrix_eq10() %>% unclass
    })
}

mydf <- cbind(log10alpha = seq(-3,3,length.out = 50), r_alpha(pars_opt = 10^seq(-3,3,length.out = 50))[-c(1,4),] %>% t %>% structure(. , dimnames = list(NULL,c( "r21", "r12")))) %>% as.data.frame(stringsAsFactors =  F)
mydf <- tidyr::gather_(mydf, key = "r_element", value = "Value", gather_cols = c("r21", "r12"))

myplot <- ggplot(mydf, mapping = aes(x = log10alpha, y = Value, color = r_element))+geom_line() + theme_dMod() + scale_color_dMod() + scale_x_continuous(breaks = scales::pretty_breaks()) 
# ggsave(filename = "Output/Prabakaran.r-values.pdf", plot = myplot)
```

## Fit and analyze different conditions and perturbations.

### Find good Parameter values to compare behaviour for different conditions.

Run this chunk of code with the "plot"-line uncommented to see predictions.
Once happy, comment out the plotting and proceed.
```{r Find good pars, eval=FALSE}
ic_list <- list(e1 = c(1000,1100),
                p1 = c(100,110)
                )

pars <- pars_0
pars[names(reduce_ic_list(ic_list)[[2]])] <- reduce_ic_list(ic_list)[[2]]
ic_list <- reduce_ic_list(ic_list)[[1]]
p_ic <- p_ic_fun(ic_list)
plot((g0*x*p_ic)(times = seq(0,100, length.out = 50), pars))
# print(pars)
```


### Create lists for different settings.
```{r Summary of pars and ic_list}
pars_perturbed_list <- list()
ic_list_list <- list()
pars_list <- list()
obs_fun_list <- list()
setting_description <- c("")
```

### Settings
A setting is all the circumstances which need to be taken into consideration when simulating.
Especially, they are:
  1. Which Parameters are perturbed? All perturbed parameters are increased 1.1-fold as their perturbation.
    This I might adapt in the future to add more control.
  2. Which modularization is to be chosen? This is governed by the observation function, which can perform any kind of combination of states. If you are interested in the modularization, go back to section [modulesandobsfuns] to read about the observation function that I used here.
  3. "Parameter conditions" To conveniently check the settings in different parameter conditions, one can specify parameters that deviate from the original pars in an *ic_list*, which is used for automated condition generation with P().
```{r Setting 1}
pars_perturbed_list <- list(c(m1 = 1.1, e1 = 1.1),
                                  c("m1" = 1.1, "p1" = 1.1),
                                  c("m1" = 1.1, "k4" = 1.1),
                                  c("m1" = 1.1, "k5on" = 1.1),
                                  c("m1" = 1.1, "k5off" = 1.1),
                                  c("m1" = 1.1, "k6" = 1.1) )
                                       
obs_fun_list <- lapply(seq_along(pars_perturbed_list), function(i) g0)

ic_list <- list(e1 = c(1000))

pars <- pars_0
# pars[names(reduce_ic_list(ic_list)[[2]])] <- reduce_ic_list(ic_list)[[2]]
# ic_list <- reduce_ic_list(ic_list)[[1]]

p_ic <- p_ic_fun(ic_list)
plot((x*p_ic)(times = seq(0,100, length.out = 50), pars))

pars_list <- lapply(seq_along(pars_perturbed_list), function(i) pars)
ic_list_list <- lapply(seq_along(pars_perturbed_list), function(i) ic_list)
```


## Fit the multiple conditions and save the fits to a file..
Run this only once before knitting and knit the document with the loaded dataset.
```{r Fit multiple conditions, eval=FALSE}
machine <- "localhost"
fit_job1<- runbg({
  myfits <- lapply(seq_along(pars_list), function(i) {
    # cat("Setting ", i, " of ", length(pars_list), " \n")
    run_simulations(ic_list = ic_list_list[[i]],
                    nfits = 12,
                    cores = 3,
                    pars = pars_list[[i]],
                    pars_perturbed = pars_perturbed_list[[i]],
                    obs_fun = obs_fun_list[[i]],
                    alpha = 1)
    })
  names(myfits) <- paste0("Setting", seq_along(pars_list))
  myfits
}, machine = machine, filename = "Prabakaran_job_intermediates")
```

```{r, eval=FALSE}
while(!fit_job1$check()) {
  Sys.sleep(5)
}
beepr::beep(2)
multiple_fits <- fit_job1$get()
multiple_fits <- multiple_fits[[machine]]
fit_job1$purge()
save(multiple_fits, file = paste0(pathout,fit_name, "Prabakaran", ".RData"))
```


## Analyze the fits.
```{r Analyze Fits to knitr-Document, results='asis', out.width='45%'}
# Load the multiple fits
load(paste0(pathout,fit_name, "Prabakaran", ".RData"))

lapply(seq_along(pars_list), function(setting) {
  
  cat("\\newpage
# Fit setting: ", setting, ": ", setting_description[setting], "
\n
### Standard pars of the following fits \n")
  print(xtable(as.matrix(pars_list[[setting]], nrow = 1), digits = 3), comment = F, floating = F)
    
  
  
  # multiple_fits <- lapply(seq_along(multiple_fits), function(j) {
  #   if(is.null(multiple_fits[[j]]$condition)) multiple_fits[[j]]$condition <- names(multiple_fits)[j]
  #   return(multiple_fits[[j]])
  # })
  output <- lapply(multiple_fits[[setting]], function(cond) {
    
    cat("\\newpage
      # Condition: ",(cond$condition %>% str_split("_") %>% extract2(1) %>% paste(collapse = ", ") %>% str_replace_all("=", " = ")), "\n
      # Perturbed Parameters: ", names(cond$pars_perturbed), " = ", cond$pars_perturbed, "\n")
    
    # Plot the result of the mstrust()
    print(plotValues(cond$all_fits) )
    print(plot(cond$prediction))
    
    cat(" \n Local response matrices before and after optimization.\n")
    cat(" Left is the local response matrix of free species, right is the matrix with complexes. \n")
    print(xtable(cbind(cond$r_0, rep(0, nrow(cond$r_0)), cond$r_alpha), digits = 5), comment = F, floating = F)
    cat(" ")
    print(xtable(cond$r_kept, digits = 5), comment = F, floating = F)
    cat(" ")
    print(xtable(cond$r_best_fit, digits = 5), comment = F, floating = F)
    
    cat("\n Best fit with reasonable parameter: \n")
    print(cond$best_fit %>% unclass %>%  matrix(.,nrow = 1, dimnames = list(NULL, names(.))) %>% xtable(digits = 3), comment = F, floating = F)
    
    cat("\n Steady state: \n")
    print(xtable(as.matrix(cond$steady_states[[1]])), comment = F, floating = F)
    
    cat("\n All fits: \n")
    print(xtable(cond$all_fits, digits = 4), comment = F, floating = F)
    return()
  })
})
```