---
title: "Ras"
author: "Daniel Lill"
output:
  pdf_document:
    keep_tex: yes
---

# Load libraries and define some useful paths

```{r Libraries, results='hide', echo=FALSE, message=FALSE}
# rm(list = ls())
library(MRAr)
library(scales)
library(xtable)
library(magrittr)
path <- paste0(getwd(), "/")
pathout <- paste0(path, "Output/")
```

```{r Names for Fit storage of different initial conditions}
fit_name <- "Ras"
```


```{r Call render without overwriting any previous documents, echo=FALSE, eval=FALSE, message=FALSE}
mytime <- format(Sys.time(), "%Y%m%d-%H%M")

rmarkdown::render("Ras.Rmd", output_file = paste0(pathout, fit_name, mytime, ".pdf"))
beepr::beep("~/Promotion/Software/rimshot.mp3")
```
# Ras

## Setup Model 
This is the setup of the ODE model.

### Prediction function
odemodel writes a C-function, which is executed in x and xs.
x is a regular prediction function while xs only finds and returns the steady states.
If the odemodel has been calculated already, it can be loaded with "load".
```{r Prediction function Ras, results='asis'}
modelpath <- paste0(path, "Models/Ras/")
modelname   <- "Ras"
mymodel     <- as.eqnlist(read.csv(paste0(modelpath, modelname, ".csv")))

# myodemodel <- odemodel(mymodel, modelname = paste0(modelname), deriv = T)
# save(myodemodel, file = paste0(modelpath, "odemodel",modelname, ".RData"))
load(paste0(modelpath, "odemodel",modelname, ".RData"))

# Prediction function
x <- Xs(odemodel = myodemodel)
# Prediction function for steady states
xs <- Xs_steady(myodemodel) # When computing steady states, set deriv = F
# print(mymodel)
# mymodel %>% as.eqnvec() %>% print(pander = T)
# myodemodel[[1]] %>% attr("eq")
```






### Modules, their observables and the observation function
Observation function which depends also on the optimization parameters. With their help, the complexes can smoothly be switched on in the observation.


```{r Observation function with Cx2z2 in x-module}
# modules <- modules0 <- c("x2","y2", "z2")
# 
# obs <- obs0 <- c("x2+a_1*Cx2y1",
#           "y2+a_2*Cy2z1",
#           "z2+a_3*Cx2z2")
# g <- g0 <- Y(as.eqnvec(structure(obs0, names = modules0)), mymodel, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```


```{r Most general observation function}
modules <- modules0 <- c("Ras","RalGEF", "RalGTP", "Raf", "MEK")

complexes <- c("RalGEF_Ras", 
               "aRalGEF_Ras", 
               "aRalGEF_Ras_RalGDP", 
               "aRalGEF_RalGDP", 
               "Raf_Ras", 
               "aRaf_Ras",  
               "aRaf_Ras_MEK", 
               "aRaf_MEK")
pars_opt_names <- paste0("a", c(1:4, 1:4), rep(c("l", "r"), each = 4))
adding_complexes <- paste0(" + ", pars_opt_names, " * ", complexes)

obs <- obs0 <- c(paste0(c("RasGTP", adding_complexes[c(2:2,6:6)]), collapse = ""), 
                 paste0(c("aRalGEF", adding_complexes[2:4]), collapse = ""), 
                 "RalGTP", 
                 paste0(c("aRaf", adding_complexes[6:8]), collapse = ""), 
                 "ppMEK")
# obs <- obs0 <- paste(obs, 
#                      c(" + a2l * RalGEF_Ras + a2r * Raf_Ras", 
#                        " + a2l * RalGEF_Ras",
#                        "",
#                        " + a2r * Raf_Ras",
#                        ""
#                        ))

g <- g0 <- Y(as.eqnvec(structure(obs0, names = modules0)), mymodel, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```


### Inner parameters
Load all parameters and format them for further use.
```{r Parameters Ras}

pars_raw <- read.csv(paste0(modelpath, "pars.csv"), header = FALSE)  # dynamical parameters
pars_raw <- structure(pars_raw$V2, names = as.character(pars_raw$V1))
ic_raw <- read.csv(paste0(modelpath, "IC.csv"), header = F)          # initial conditions
ic_raw <- structure(ic_raw$V2, names = as.character(ic_raw$V1))
# sort ics like they appear in x, so it can be used in stode()
vars_x <- attr(x, "parameters")[attr(x, "parameters") %in% names(ic_raw)]
ic_raw <- ic_raw[vars_x]

pars_inner_opt_0 <- setdiff(attr(g, "parameters"),names(c(pars_raw,ic_raw)))
pars_inner_opt_0 <- structure(rep(0, length(pars_inner_opt_0)), 
                              names = pars_inner_opt_0)

pars_inner_0 <- pars_inner <- c(ic_raw, pars_raw, pars_inner_opt_0)


# check if all parameters are assigned a value. Both lines should evaluate to the same number.
x_pars <- unique(union((g) %>% attr("parameters"),(xs) %>% attr("parameters")))
# names(pars_inner) %in% x_pars %>% sum()
# x_pars %in% names(pars_inner) %>% sum()
```


### General log transformation and setting some pars to zero
It's better to do a logtransformation for all Fitting purposes and work with the log-Parameters.
The way to implement is this, is
log_pars %>% p_log %>% p_other %>% p_other_2 %>% ...
```{r}
# Logtrafo of all pars
logtrafo <- structure(paste0("exp( log", names(pars_inner_0), ")"), 
                             names = names(pars_inner_0))

zero_at_start <- c("RasGTP", "aRalGEF", "RalGEF_Ras", "aRalGEF_Ras", "RalGTP", "aRaf", "Raf_Ras", "aRaf_Ras", "ppMEK", "aRalGEF_Ras_RalGDP", "aRaf_Ras_MEK", "aRalGEF_RalGDP", "aRaf_MEK"  )
logtrafo[zero_at_start] <- "0"

p_log <- P(trafo = logtrafo,
           # compile = TRUE,
           modelname = "p_log")
           


pars_inner
# pars_0
# p_log(pars_0, deriv = F)
```
# Simulate the system to explore the behavior and get good parameters.

```{r Find good pars}
pars <- pars_inner_0
pars["k7f"] <- 1
pars["k7r"] <- 600
pars["k8"] <- 100 

pars["k9f"] <- 1e-4 # activation of MEK via aRaf_Ras
pars["k9r"] <- 1
pars["k9cat"] <- 1

pars["k16f"] <- 1e-4 # activation of MEK via aRaf
pars["k16r"] <- 1
pars["k16cat"] <- 1

pars["k13"] <- 5
pars["k14"] <- 200 # deactivation of aRaf


pars["k10m"] <- 1e2
pars["v10"] <- 5e0
Raf_path <- names(ic_raw)[str_detect(names(ic_raw), "Raf")|str_detect(names(ic_raw), "MEK")]


pars["k3f"] <- 1
pars["k3r"] <- 400
pars["k4"] <- 10

pars["k5f"] <- 1e-3 # activation of RalGDP by complex
pars["k5r"] <- 1
pars["k5cat"] <- 1

pars["k15f"] <- 1e-3 # activation of RalGDP by aRalGEF
pars["k15r"] <- 1
pars["k15cat"] <- 1

pars["k12"] <- 200 # deactivation of aRalGEF


pars["v6"] <- 1e3 # deactivation aRalGDP
pars["km6"] <- 10000

RalGEF_path <- names(ic_raw)[str_detect(names(ic_raw), "Ral")]

pred <- (x)(10^(seq(0,4, length.out = 1000)), pars = pars, deriv = FALSE) %>% set_names("c1")
predplot <- plotPrediction(pred)#, ! name %in% Raf_path)
predplot + scale_x_log10()
# plotly::ggplotly(predplot)

```

Take parameter values from the previous chunk to continue
```{r}
pars_inner <- pars
pars_inner <- pars_inner[!names(pars_inner)%in%zero_at_start]


pars_0 <- pars <- log(pars_inner) %>% set_names(paste0("log", names(pars_inner))) 

pars_opt_0 <- pars_opt <- structure(log(pars_inner_opt_0+.Machine$double.eps), names = paste0("log", names(pars_inner_opt_0)))
pars[names(pars_opt_0)] <- -750
# pars_opt_0 <- NULL
```



## Show the code
### Perturb modules
Now the module can be simulated to generate perturbation data.

### Parameter trafos for convenient perturbation data
Just to show the code...
Set up the perturbations
```{r}
# Which pars shall be perturbed?
pars_perturbed_0 <- pars_perturbed <- c("logRasGDP" = log(0.9), 
                                        "logRalGEF" = log(0.9), 
                                        "logRalGDP" = log(0.9),
                                        "logRaf" = log(0.9),
                                        "logMEK" = log(0.9)
                                        )
# pars_perturbed_0 <- pars_perturbed <- c("logx1" = log(0.8), "logy1" = log(0.8), "logz1" = log(0.8))



# pars_perturbed
p_pert <- p_pert_fun(pars_perturbed = pars_perturbed, pars = pars)


perturbation_prediction_0 <- perturbation_prediction <-  (xs*p_log*p_pert)(times = c(0,Inf), pars = pars, deriv = F)

# mytime <- 2000
# (x*p_log*p_pert)(times = seq(0,2000, length.out = 50), pars = pars) %>% 
#   getDerivs() %>% 
#   extract2(1) %>%
#   apply_expression({x[,c(str_detect(colnames(x), "(Cx2y1\\.)|time"))]}) %>% prdlist() %>% set_names("blas") %>% 
#   plotPrediction()



# perturbation_prediction %>% lapply(round,1)
R_fun() %>% 
  local_response_matrix_eq10()  %>% 
  round(2) %>% 
  # plot_r_graph(modules = modules0, inhibitors = attr(p_pert, "co")[-1]) %>% 
  apply_expression(x)


```

```{r}
plot_r_graph <- function(r, modules = modules_0, inhibitors = inhibitors_0) {
  vertices <- data.frame(id = paste0("m",1:length(modules)), module = modules, condition = c(inhibitors))
  
  
  make_edges <- function(r) {
    mygrid <- expand.grid(paste0("m",1:nrow(r)), 
                          paste0("m",1:ncol(r)), 
                          stringsAsFactors = FALSE) %>% 
      structure(names = c("to", "from")) %>% 
      extract(c(2,1))
    weights <- abs(r) %>% matrix(ncol = 1)
    type <- sign(r) %>% as.character %>% factor(levels = c("1", "-1"), labels = c("activates", "inhibits"))
    
    return(data.frame(mygrid, value = weights, type = type))
  }
  
  
  edges <- make_edges(r)
  r_graph <- igraph::graph.data.frame(d = edges, vertices = vertices, directed = T) %>% 
    igraph::simplify(remove.loops = TRUE, remove.multiple = FALSE)
  
  igraph::V(r_graph)$label <- igraph::V(r_graph)$module
  # V(r_graph)$size <- 0
  igraph::V(r_graph)$shape <- "none"
  
  type <- igraph::E(r_graph)$type %>% as.factor
  igraph::E(r_graph)$color <- c("black", "red")[type]
  igraph::E(r_graph)$curved <- 0.1
  igraph::E(r_graph)$width <- igraph::E(r_graph)$value/max(igraph::E(r_graph)$value)*5
  igraph::E(r_graph)$arrow.size <- .3
  
  r_layout <- igraph::layout.circle(r_graph)
  
  plot(r_graph, layout = r_layout)
}

```


```{r compare r0 r1 ropt}

r_kept_0 <- r_kept_fun(pars_opt = pars_opt,
                       perturbation_prediction = perturbation_prediction,
                       obs_fun = g,
                       p_fun = (p_log * p_pert),
                       pars = pars,
                       alpha = -log(0+.Machine$double.eps))
r_kept_0

# r_kept_0 <- c(F,F,T,T,F,F,F,T,F) %>% matrix(ncol = 3)

r_0 <-   R_fun(pars_opt = pars_opt,
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_0%>% round(2)

r_1 <-   R_fun(pars_opt = rep(0, length(pars_opt)) %>% set_names(names(pars_opt)),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)

# r_kept <- matrix(F, nrow = 5, ncol = 5)
# r_kept[c(6,9)] <- T

myfits <- mstrust(obj_alpha, center =  rep(0, length(pars_opt)) %>% set_names(names(pars_opt)), studyname = "Fits", cores = 3, fits = 6, sd = 1, mypars = pars, perturbation_prediction = perturbation_prediction_0, r_kept = r_kept_0)
# myfits %>% as.parframe() 
best_fit <- myfits %>% as.parframe() %>% as.parvec()
myfits %>% as.parframe()
r_alpha_fun(pars_opt = unclass(best_fit), pars = pars, perturbation_prediction = perturbation_prediction) %>% round(2)
```
```{r}
sum(r_kept_0)
```



# Dose response plots
```{r, eval=FALSE}
par_opt_names <- pars_opt_0 %>% names %>% sort

par_opt_settings <- list(par_opt_names)

dr_list <- lapply(par_opt_settings, function(i) {
  r_opt_dose_response(which_par = "logRasGDP", 
                      dosages = pars["logRasGDP"]+seq(-2,0,by = 2),
                      pars_opt = pars_opt_0[i],
                      alpha_scan_range = sort(unique(seq(-9,1,by = .1))),
                      pars = pars
  )
})
saveRDS(dr_list, file = "Output/dr_list Ras each complex has its own par_opt.rds")

# dr_list <- lapply(seq_along(par_opt_settings), function(i) {
# 
#   scan_grid(which_par = "logk111",
#             dosages = sort(c(unique(seq(-16,0,by = 0.5)))),
#             pars = pars,
#             alpha_scan_range = c(-9,seq(-3,2,by = 0.1)),
#             pars_opt = pars_opt_0[par_opt_settings[[i]]],
#             grid_points_raw = c(-9,-3,0,2))
# 
# })

```


# General plotting functions
Plot as dos-response with alpha as facets
```{r Plot generation, eval=FALSE}
# dr_list <- readRDS("Output/dr_list Ras.rds")
myplot <- c(dr_list) %>% 
  combine_dr_list() %>%
  filter(! (r_element %in% paste0("r", c(11,22,33,44,55)))) %>%
  # filter(near(dose, pars["logRasGDP"]+seq(-2,0,by = 1)[1])) %>%
  # filter(par_opt_setting == 1) %>% 
  # filter(alpha %in% 0) %>% 
  # filter(matrix != "r_kept") %>%
  apply_expression({
    x$matrix[x$matrix == "r_alpha"] <- "r(a_i = 1)"
    x$matrix[x$matrix == "r_opt"] <- "r(a = a_opt)"
    x$matrix <- x$matrix %>% factor(levels = c("r(a_i = 1)", "r(a = a_opt)", "r_kept"))
    x
  }) %>% 
  mutate(Element = r_element) %>% 
  ggplot(mapping = aes(x = alpha, y = value, color = Element, linetype = Element, frame = dose)) +
  geom_line() +
  theme_dMod() +
  # scale_color_dMod() +
  geom_abline(slope = 0,intercept = 0, linetype = 3) +
  facet_grid(matrix~., scales = "free") #+
  # xlab("log(gu)")
# myplot
plotly::ggplotly(myplot)
# ggsave(plot = myplot,
#        filename = paste0("~/Promotion/Writing/Papers/2017 02 MRA optimization procedure Paper/pictures/plots/", "gu large sequestration",".png"),
#        width = 10,
#        height = 10,
#        units = "cm",
#        device = "png")

```

# Plot the matrix as facet_wrap with frame = alpha to see the effect of certain sign flips.

```{r, eval=FALSE}
# dr_list <- readRDS("Output/dr_list Ras.rds")
myplot <-
  c(dr_list) %>% 
  combine_dr_list() %>%
  # filter(! (r_element %in% paste0("r", c(11,22,33,44,55)))) %>% 
  apply_expression({
    myelements <- paste0("r", outer(1:5,1:5, paste0))
    element_names  <- outer(modules, modules, paste, sep = " by " ) 
    for ( i in 1:25 ) {
      x$r_element[x$r_element==myelements[i]] <- element_names[i]
    }
    element_names <- element_names %>% matrix(ncol = 5) %>% t %>% as.character() # transpose here, because in facet_wrap, "dir" doesn't work
    x$r_element <- x$r_element %>% factor(levels = element_names)
    x
  }) %>% 
  filter(near(dose, pars["logRasGDP"]+seq(-2,0,by = 2)[1])) %>%
  # filter(par_opt_setting == 1) %>% 
  # filter(alpha %in% 0) %>% 
  filter(matrix %in% c("r_opt", "r_kept")) %>%
  apply_expression({
    x$matrix[x$matrix == "r_alpha"] <- "r(a_i = 1)"
    x$matrix[x$matrix == "r_opt"] <- "r(a = a_opt)"
    x$matrix <- x$matrix %>% factor(levels = c("r(a_i = 1)", "r(a = a_opt)", "r_kept"))
    x
  }) %>% 
  mutate(Element = r_element) %>% 
  mutate(value = ifelse(matrix == "r_kept", value * 4, value)) %>% 
  apply_expression({
    x$value[abs(x$value)>10] <- NA
    x
  }) %>% 
  mutate(zero = 0) %>% 
  select(c(1,5,6,10,11)) %>% 
  unique() %>% 
  filter(alpha < 1) %>% 
  ggplot(mapping = aes(x = zero, y = value, color = Element, shape = matrix, frame = alpha)) +
  geom_point() +
  theme_dMod() +
  # scale_color_dMod() +
  geom_abline(slope = 0,intercept = 0, linetype = 3) +
  facet_wrap(~Element, scales = "fixed") #+
  # xlab("log(gu)")
# myplot
plotly::ggplotly(myplot)
# ggsave(plot = myplot,
#        filename = paste0("~/Promotion/Writing/Papers/2017 02 MRA optimization procedure Paper/pictures/plots/", "blablabalbalablablablabal",".png"),
#        width = 10,
#        height = 10,
#        units = "cm",
#        device = "png")

```

It seems as if mainly r12 and r14 affect the outcome.
```{r compare r0 r1 ropt with given rkept}

# r_kept_0 <- r_kept_fun(pars_opt = pars_opt,
#                        perturbation_prediction = perturbation_prediction,
#                        obs_fun = g,
#                        p_fun = (p_log * p_pert),
#                        pars = pars,
#                        alpha = -log(0+.Machine$double.eps))
r_kept_0 <- matrix(F, nrow = 5, ncol = 5)
r_kept_0[1,c(2,4)] <- T
r_kept_0[4,2] <- r_kept_0[2,4] <- T
r_kept_0

# r_kept_0 <- c(F,F,T,T,F,F,F,T,F) %>% matrix(ncol = 3)

r_0 <-   R_fun(pars_opt = pars_opt,
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_0%>% round(2)

r_1 <-   R_fun(pars_opt = rep(0, length(pars_opt)) %>% set_names(names(pars_opt)),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)

# r_kept <- matrix(F, nrow = 5, ncol = 5)
# r_kept[c(6,9)] <- T

myfits <- mstrust(obj_alpha, center =  rep(0, length(pars_opt)) %>% set_names(names(pars_opt)), studyname = "Fits", cores = 3, fits = 50, sd = 1, mypars = pars, perturbation_prediction = perturbation_prediction_0, r_kept = r_kept_0)
# myfits %>% as.parframe() 
best_fit <- myfits %>% as.parframe() %>% as.parvec()
myfits %>% as.parframe()
r_alpha_fun(pars_opt = unclass(best_fit), pars = pars, perturbation_prediction = perturbation_prediction) %>% round(2)
```

```{r compare r0 r1 ropt with given rkept}

# r_kept_0 <- r_kept_fun(pars_opt = pars_opt,
#                        perturbation_prediction = perturbation_prediction,
#                        obs_fun = g,
#                        p_fun = (p_log * p_pert),
#                        pars = pars,
#                        alpha = -log(0+.Machine$double.eps))
r_kept_0 <- matrix(F, nrow = 5, ncol = 5)
# r_kept_0[1,c(2,4)] <- T
r_kept_0[4,2] <- r_kept_0[2,4] <- T
r_kept_0

# r_kept_0 <- c(F,F,T,T,F,F,F,T,F) %>% matrix(ncol = 3)

r_0 <-   R_fun(pars_opt = pars_opt,
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_0%>% round(2)

r_1 <-   R_fun(pars_opt = rep(0, length(pars_opt)) %>% set_names(names(pars_opt)),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)

# r_kept <- matrix(F, nrow = 5, ncol = 5)
# r_kept[c(6,9)] <- T

myfits <- mstrust(obj_alpha, center =  rep(0, length(pars_opt)) %>% set_names(names(pars_opt)), studyname = "Fits", cores = 3, fits = 6, sd = 1, mypars = pars, perturbation_prediction = perturbation_prediction_0, r_kept = r_kept_0)
# myfits %>% as.parframe() 
best_fit <- myfits %>% as.parframe() %>% as.parvec()
myfits %>% as.parframe()
r_alpha_fun(pars_opt = unclass(best_fit), pars = pars, perturbation_prediction = perturbation_prediction) %>% round(2)
```


```{r compare r0 r1 ropt with given rkept}

# r_kept_0 <- r_kept_fun(pars_opt = pars_opt,
#                        perturbation_prediction = perturbation_prediction,
#                        obs_fun = g,
#                        p_fun = (p_log * p_pert),
#                        pars = pars,
#                        alpha = -log(0+.Machine$double.eps))
r_kept_0 <- matrix(F, nrow = 5, ncol = 5)
r_kept_0[1,c(2,4)] <- T
# r_kept_0[4,2] <- r_kept_0[2,4] <- T
r_kept_0

# r_kept_0 <- c(F,F,T,T,F,F,F,T,F) %>% matrix(ncol = 3)

r_0 <-   R_fun(pars_opt = pars_opt,
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_0%>% round(2)

r_1 <-   R_fun(pars_opt = rep(0, length(pars_opt)) %>% set_names(names(pars_opt)),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)

# r_kept <- matrix(F, nrow = 5, ncol = 5)
# r_kept[c(6,9)] <- T

myfits <- mstrust(obj_alpha, center =  rep(0, length(pars_opt)) %>% set_names(names(pars_opt)), studyname = "Fits", cores = 3, fits = 6, sd = 1, mypars = pars, perturbation_prediction = perturbation_prediction_0, r_kept = r_kept_0)
# myfits %>% as.parframe() 
best_fit <- myfits %>% as.parframe() %>% as.parvec()
myfits %>% as.parframe()
r_alpha_fun(pars_opt = unclass(best_fit), pars = pars, perturbation_prediction = perturbation_prediction) %>% round(2)
```
















