---
title: "Prabakaran-Model"
author: "Daniel Lill"
output: 
  pdf_document:
    keep_tex: true
---


In this document, I test R_fun_2 on the Prabakaran-model.
R_fun_2 doesn't take the derivative of the linear combination, instead it takes a linear combination of the derivatives. This is much easier measurable with co-ips.



# Load libraries and define some useful paths

```{r Libraries, results='hide', echo=FALSE, message=FALSE}
library(MRAr)
library(scales)
library(xtable)
path <- paste0(getwd(), "/")
pathout <- paste0(path, "Output/")
```

```{r Names for Fit storage of different initial Conditions}
fit_name <- "Prabakaran"
```

```{r Call render without overwriting any previous documents, echo=FALSE, eval=FALSE, message=FALSE, include=FALSE}

rmarkdown::render("Prabakaran.Rmd", output_file = paste0(pathout, format(Sys.time(), "%Y%m%d-%H%M")
, " - " , fit_name, ".pdf"))
beepr::beep("~/Promotion/Software/rimshot.mp3")
```

# Prabakaran-Model

### Prediction function.
odemodel writes a C-function, which is executed in x and xs.
x is a regular prediction function while xs only finds and returns the steady states.
If the odemodel has been calculated already, it can be loaded with "load".
```{r Prediction function}
modelpath <- paste0(path, "Models/Prabakaran/")
modelname   <- "Prabakaran"
mymodel     <- as.eqnlist(read.csv(paste0(modelpath, modelname, ".csv")))
# mymodel

# myodemodel <- odemodel(mymodel, modelname = modelname, deriv = T)
# save(myodemodel, file = paste0(modelpath, "odemodel",modelname, ".RData"))
load(paste0(modelpath, "odemodel",modelname, ".RData"))


# Prediction function
x <- Xs(odemodel = myodemodel)
# Prediction function for steady states
xs <- Xs_steady(myodemodel)
print(mymodel)
```

### Modules, their observables and the observation function. {#modulesandobsfuns}
Observation function which depends also on the optimization parameters. With their help, the complexes can smoothly be switched on in the observation.

This is the naive modularization used in the Prabakaran Paper, desregarding, which species are actually the communicating ones. This is fair enough, because usually the "active" forms are consedered to be the communicating species in a non-parametric network reconstruction.
```{r Naive observation function}
modules <- modules0 <- c("m2","e2tot")

obs <- obs0 <- c("m2+a_1*Cm2e1",
                 "e2+Ce2p1")
g <- g0 <- Y(as.eqnvec(structure(obs0, names = modules0)), mymodel, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```



### Parameters.
Load all parameters and format them for further use.
```{r Parameters}

pars_raw <- read.csv(paste0(modelpath, "pars.csv"), header = FALSE)  # dynamical parameters
pars_raw <- structure(pars_raw$V2, names = as.character(pars_raw$V1))
ic_raw <- read.csv(paste0(modelpath, "IC.csv"), header = F)          # initial conditions
ic_raw <- structure(ic_raw$V2, names = as.character(ic_raw$V1))
# sort ics like they appear in x, so it can be used in stode()
vars_x <- attr(x, "parameters")[attr(x, "parameters") %in% names(ic_raw)]
ic_raw <- ic_raw[vars_x]

pars_opt_0 <- setdiff((g) %>% attr("parameters"),names(c(pars_raw,ic_raw)))
pars_opt_0 <- structure(rep(0, length(pars_opt_0)), names = pars_opt_0)

pars_0 <- pars <- c(ic_raw, pars_raw, pars_opt_0)


# check if all parameters are assigned a value. Both lines should evaluate to the same number.
x_pars <- unique(union((g) %>% attr("parameters"),(xs) %>% attr("parameters")))
names(pars) %in% x_pars %>% sum()
x_pars %in% names(pars) %>% sum()

```


## Example code
Now the module can be simulated to generate perturbation data.

### Parameter trafos for convenient perturbation data.
Set up the perturbations
```{r which pars are perturbed}
# Which pars shall be perturbed?
pars_perturbed_0 <- pars_perturbed <- c(m1 = 1.1, e1 = 1.1)

p_pert <- p_pert_fun(pars_perturbed = pars_perturbed_0,
                     pars = pars_0)
```

### Perturbation data generation and elements to keep.
Just to show the code...
```{r Generate perturbation data}
perturbation_prediction_0 <-  (xs*p_pert)(times = c(0,Inf), pars = pars_0, deriv = F)
r_kept_0 <- r_kept_fun(obs_fun = g0)

```



# Test R_fun_2
Plot r(R_fun_2) against alpha and see if it qualitatively behaves the same as r(R_fun)
```{r}
perturbation_prediction <-  (xs*p_pert)(times = c(0,Inf), pars = pars, deriv = F)
alpha_scan_range <- seq(-2,2,by = 0.1)

r_df <- sapply(alpha_scan_range, function(par_opt) {
  pars_opt <- structure(10^(par_opt), names = c("a_1"))
  myr <- R_fun(pars_opt = pars_opt, p_fun = (p_pert)) %>% local_response_matrix_eq10()
})%>%
  t %>%
  structure(. , dimnames = list(NULL, r_names_0)) %>%
  cbind(log10alpha = alpha_scan_range, .) %>%
  as.data.frame(stringsAsFactors =  F) %>%
    cbind(R= "R") %>% 
  tidyr::gather_(key = "r_element", value = "r_alpha", gather_cols = r_names_0)

r_2_df <- sapply(alpha_scan_range, function(par_opt) {
  pars_opt <- structure(10^(par_opt), names = c("a_1"))
  myr <- R_fun_2(pars_opt = pars_opt) %>% local_response_matrix_eq10()
})%>%
  t %>%
  structure(. , dimnames = list(NULL, r_names_0)) %>%
  cbind(log10alpha = alpha_scan_range, .) %>%
  as.data.frame(stringsAsFactors =  F) %>%
  cbind(R= "R2") %>% 
  tidyr::gather_(key = "r_element", value = "r_alpha", gather_cols = r_names_0)

r_df <- rbind(r_df, r_2_df)

myplot <- r_df %>%
  # filter(! (r_element %in% paste0("r", c(11,22)))) %>%
  # filter( near(alpha,-9) | near(alpha,0) | near(alpha, 9)) %>%
  # filter(dose < 4) %>%
  # gather(key = "matrix", value  = "value", r_alpha, r_kept, r_opt) %>%
  ggplot(mapping = aes(x = log10alpha, y = r_alpha, color = r_element)) +
  geom_line() +
  # xkcd::theme_xkcd() +
  scale_color_dMod() +
  # scale_x_continuous(breaks = -9:12) +
  scale_y_continuous(limits = c(-3,3))+
  geom_hline(yintercept = 0, linetype = 3) +
  geom_vline(xintercept = 0, linetype = 3) +
  facet_wrap(~R, scales = "free")

plotly::ggplotly(myplot)
```
