---
title: "Cascade - x2z2 mass action feedback"
author: "Daniel Lill"
output:
  pdf_document:
    keep_tex: yes
---

# Load libraries and define some useful paths

```{r setup, results='hide', echo=FALSE, message=FALSE}
rm(list = ls())
library(MRAr)
library(scales)
library(xtable)
library(parallel)
path <- paste0(getwd(), "/")
pathout <- paste0(path, "Output/")
```

# Cascade

## Setup Model 
This is the setup of the ODE model.

### Prediction function
odemodel writes a C-function, which is executed in x and xs.
x is a regular prediction function while xs only finds and returns the steady states.
If the odemodel has been calculated already, it can be loaded with "load".
```{r Prediction function Cascade, results='asis'}
modelname   <- "Cascade"
mymodel     <- as.eqnlist(read.csv(paste0( modelname, ".csv")))

# myodemodel <- odemodel(mymodel, modelname = paste0(modelname), deriv = T)
# save(myodemodel, file = paste0( "odemodel",modelname, ".RData"))
load(paste0( "odemodel",modelname, ".RData"))

# Prediction function
x <- Xs(odemodel = myodemodel)
# Prediction function for steady states
xs <- Xs_steady(myodemodel) # When computing steady states, set deriv = F
loadDLL(x)
# print(mymodel)
# mymodel %>% as.eqnvec() %>% print(pander = T)
# myodemodel[[1]] %>% attr("eq")
```

### Modules, their observables and the observation function
Observation function which depends also on the optimization parameters. With their help, the complexes can smoothly be switched on in the observation.

```{r Most general observation function}
modules <- modules0 <- c("X","Y", "Z")

mypars_opt <- paste0("a_", outer(paste0("to", c("x","y","z"), "_C"), c("xy","yz","zx"), paste0)) %>% matrix(ncol=3)
mypars_opt[c(3,4,8)] <- "0"


mycomplexes <- matrix(c("Cx2y1", "Cy2z1", "Cx2z2"), ncol = 1)

samepars_opt <- paste0("a_", outer(rep("C",3), c("xy","yz","zx"), paste0)) %>% matrix(ncol=3)
samepars_opt[c(3,4,8)] <- "0"

obs <- obs0 <- paste(c("x2", "y2", "z2"), "+", prodSymb(mypars_opt, mycomplexes), "+", prodSymb(samepars_opt, mycomplexes))


g <- g0 <- Y(as.eqnvec(structure(obs0, names = modules0)), mymodel, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)

```


### Inner parameters
Load all parameters and format them for further use.
```{r Parameters Cascade}
pars_raw <- read.csv(paste0( "pars.csv"), header = FALSE)  # dynamical parameters
pars_raw <- structure(pars_raw$V2, names = as.character(pars_raw$V1))
ic_raw <- read.csv(paste0( "IC.csv"), header = F)          # initial conditions
ic_raw <- structure(ic_raw$V2, names = as.character(ic_raw$V1))
# sort ics like they appear in x, so it can be used in stode()
vars_x <- attr(x, "parameters")[attr(x, "parameters") %in% names(ic_raw)]
ic_raw <- ic_raw[vars_x]

pars_inner_opt_0 <- setdiff(attr(g, "parameters"),names(c(pars_raw,ic_raw)))
pars_inner_opt_0 <- structure(rep(0, length(pars_inner_opt_0)), 
                              names = pars_inner_opt_0)

pars_inner_0 <- pars_inner <- c(ic_raw, pars_raw, pars_inner_opt_0)
```


### General log transformation and setting some pars to zero
It's better to do a logtransformation for all Fitting purposes and work with the log-Parameters.
The way to implement is this, is
log_pars %>% p_log %>% p_other %>% p_other_2 %>% ...
```{r}
trafo <- getParameters(x,g) %>% 
  setNames(.,.) %>% 
  insert("x~0", x = getSymbols(mytrafo[[i]]) %>% str_subset("deg|syn")) %>% # set synthesis and degradations to zero
  insert("x~0", x = getSymbols(mytrafo[[i]]) %>% str_subset("^C*[xyz](2|1.2)")) %>% # set inits to zero
  insert("x~0", x = c("k131", "k132", "k14")) %>% 
  insert("x~0", x = c("k111", "k112", "k12")) %>% 
  insert("x~exp(x)", x = getSymbols(mytrafo[[i]])) %>%
  # {.} %T>% print %>% 
  {.} 

p_log <- P(trafo = trafo,
           modelname = "p_log",
           compile = T)

pars <- getParameters(p_log) %>% are_names_of(0) %>% sort_by_name()
pars_opt <- names(pars_inner_opt_0) %>% are_names_of(.Machine$double.eps) %>% log
```


```{r}
pars <- c(
  a_Cxy    	=	.Machine$double.eps	, # 1
  a_Cyz    	=	.Machine$double.eps	, # 2
  a_Czx    	=	.Machine$double.eps	, # 3
  a_tox_Cxy	=	.Machine$double.eps	, # 4
  a_tox_Czx	=	.Machine$double.eps	, # 5
  a_toy_Cxy	=	.Machine$double.eps	, # 6
  a_toy_Cyz	=	.Machine$double.eps	, # 7
  a_toz_Cyz	=	.Machine$double.eps	, # 8
  a_toz_Czx	=	.Machine$double.eps	, # 9
  
  gainlf   	=	1	, # 10
  gainuf   	=	1	, # 11
  k9       	=	100	, # 29
  k10      	=	100	, # 13
  
  k1       	=	100	, # 12
  v1       	=	1.5	, # 30
  k2       	=	100	, # 20
  v2       	=	1	, # 31
  
  k31      	=	0.002	, # 21
  k32      	=	0.4	, # 22
  k4       	=	0.05	, # 23
  k5       	=	100	, # 24
  v5       	=	5	, # 32
  
  k61      	=	0.005	, # 25
  k62      	=	0.4	, # 26
  k7       	=	0.05	, # 27
  k8       	=	100	, # 28
  v8       	=	1	, # 33
  
  
  x1       	=	100	, # 34
  y1       	=	100	, # 35
  z1       	=	100	  # 36 
) %>% 
  log


```




# Apply different perturbations
```{r}
pp_x <- c("x1", "k1", "v1", "k2", "v2", "k111", "k112", "k12") 
pp_y <- c("y1", "k5", "v5", "k31", "k32", "k4") 
pp_z <- c("z1", "k8", "v8", "k61", "k62", "k7", "k111", "k112", "k12")

mygrid <- expand.grid(pp_x, pp_y, pp_z, stringsAsFactors = F) 
pp_list <- map(1:nrow(mygrid), function(i) {
  mynames <- mygrid[i,] %>% unlist() %>% unname() %>% unique()
  if(length(mynames)<3) return(NULL)
  mynames %>% are_names_of(log(0.9))
}) %>% 
  .[!map_lgl(., is.null)]
# alpha <- 0

myrunbg_job <- runbg({
  cores <- detectFreeCores()
  pars0 <- pars
  
  gainlfs <- seq(-10,10,5) %>% c(seq(-2,2,0.5)) %>% unique() %>% sort
  map(gainlfs, function(gainlf) {
    pars <- pars0
    pars["gainlf"] <- c(gainlf = gainlf)
    
    alphas <- seq(-2,3,0.2)
    map(alphas, function(alpha) {
      
      mclapply(pp_list, function(pp) {
        p_pert <- p_pert_fun(pp, pars = pars, modelname = NULL)
        
        myframe <- dMod.frame("1", g, x, p_log, NULL, NULL, p_pert = list(p_pert), xs = list(xs), pars = list(pars))
        
        g <- myframe$g[[1]]
        p_pert <- myframe$p_pert[[1]]
        p <- myframe$p[[1]]
        xs <- myframe$xs[[1]]
        pars <- myframe$pars[[1]]
        
        # browser()
        perturbation_prediction <- (xs*p*p_pert)(c(0,Inf), pars, deriv = F)
        
        try({r_0 <- R_fun(pars_opt = pars_opt,
                         perturbation_prediction = perturbation_prediction,
                         obs_fun = g,
                         p_fun = (p*p_pert),
                         pars = pars) %>% local_response_matrix_eq10()})
        if(inherits(r_0, "try-error")) return(NULL)
        
        algo <- function(which_pars_opt, prefix = NULL) {
          
          out <- tibble(r_kept = list(NULL), parframes = list(NULL), r_opt = list(NULL))

          r_kept <- r_kept_fun(pars_opt = pars_opt[which_pars_opt],
                              perturbation_prediction = perturbation_prediction,
                              obs_fun = g,
                              p_fun = (p * p_pert),
                              pars = pars,
                              alpha =
                                -log(.Machine$double.eps) + alpha)
          out$r_kept <- list(r_kept)
          
          myfits <- mstrust(obj_alpha,
                            center =  structure(rep(0, length(which_pars_opt)), names = which_pars_opt),
                            studyname = "Fits",
                            cores = 1,
                            fits = 6,
                            sd = 1,
                            mypars = pars,
                            perturbation_prediction = perturbation_prediction,
                            r_kept = r_kept,
                            p_fun = (p * p_pert),
                            obs_fun = g)
          myparframe <- try(myfits %>% as.parframe())
          out$parframes <- list(myparframe)
          if (inherits(myparframe, "try-error")) {
            out$parframes <- list(myfits)
            return(out)
            }
          
          r_opt <- r_alpha_fun(pars_opt = myfits %>% as.parframe() %>% as.parvec() %>% unclass() ,
                              pars = pars,
                              perturbation_prediction = perturbation_prediction,
                              p_fun = (p*p_pert))
          out$r_opt <- list(r_opt)
          return(out)
        }
        
        to_upstream_module <- algo(c("a_tox_Cxy", "a_toy_Cyz", "a_toz_Czx"), "upstream")
        to_both_modules <- algo(c("a_Cxy", "a_Cyz", "a_Czx"), "both")
        
        bind_cols(tibble(r_0 = list(r_0 %>% round(2)), pars_perturbed = list(pp), alpha = alpha, gainlf = gainlf), to_upstream_module, to_both_modules)
      }, mc.cores = cores) %>% 
        do.call(rbind,.) 
    }) %>% 
      do.call(rbind,.)
  }) %>% 
    do.call(rbind,.)
  
}, machine = c("fermi"), input = ls(pos=.GlobalEnv), filename = "2018_07_18_33_10_gainlf"
, recover = T
)
# myrunbg_job$check()
# myrunbg <- myrunbg_job$get()$fermi
# saveRDS(myrunbg, file = "2018_07_18_33_10_gainlfmyrunbg.rds")
# results <- readRDS("2018_07_18_33_10_gainlfmyrunbg.rds")
# myrunbg_job$purge()

# saveRDS(results, "test_different_perturbations.rds")
```

```{r}
results %>% 
  # head(1)
  str1
```

```{r}
results$gainlf %>% unique()
cat("\n")
results$alpha %>% unique()
```


```{r}

bla <- results %>% 
  filter(alpha == 0) %>% 
  filter(gainlf == 0) %>% 
  filter(map_lgl(pars_perturbed,. %>% names %>% str_detect_any("k1[12]") %>% `!`)) %>% 
  # filter(map_lgl(pars_perturbed, . %>% names %>% str_detect_any("[kv][45]"))) %>%
with({
  
  map(1:length(r_0), function(i) {
    
    cat(names(pars_perturbed[[i]]), "\t", "alpha = ", alpha[i],  "\t", "gainlf = ", gainlf[i], "\n")
    # print(upstreamr_kept[[i]])
    # cat("\n")
    # print(r_kept[[i]]) # r_kept = upstream, r_kept1 = both
    print(r_kept1[[i]]) # r_kept = upstream, r_kept1 = both
    # print(r_opt[[i]] %>% round(2)) # r_kept = upstream, r_kept1 = both
    print(r_opt1[[i]] %>% round(2)) # r_kept = upstream, r_kept1 = both
    cat("\n\n")
    
  })
  
})
```


```{r}
# comp to upstream
r_kept_0 <- matrix(c(0,0,0,1,0,0,0,1,0) %>% as.logical, 3)  # %>% `rownames<-`(LETTERS[24:26]) %>% `colnames<-`(LETTERS[24:26])
wup <- 
  results %>% 
  filter(alpha == 2) %>% 
  filter(gainlf == .5) %>% 
  filter(map_lgl(r_kept, . %>% identical(r_kept_0))) %>% # r_kept = upstream
  .[["pars_perturbed"]] %>% 
  map(names) %>% 
  do.call(rbind,.)
  
  wup
```


```{r}
# comp to both
r_kept_0 <- matrix(c(0,0,0,1,0,0,0,1,0) %>% as.logical, 3)  # %>% `rownames<-`(LETTERS[24:26]) %>% `colnames<-`(LETTERS[24:26])
wup <- 
  results %>% 
  filter(alpha == 2) %>% 
  filter(gainlf == 0.5) %>% 
  filter(map_lgl(r_kept1, . %>% identical(r_kept_0))) %>% # r_kept1 = both
  # filter(map_lgl(r_kept1, 8)) %>% #take those whose 8th element of rkept is true, ie where r23 is in the optimization
  .[["pars_perturbed"]] %>% 
  map(names) %>% 
  do.call(rbind,.)
  
  wup

```



# Results
gainlf = c(0.5, 2), alpha = 2:
    * comp to upstream führen die z-perturbations k7 k8 v8 noch zu r23 in der optimierung (nicht gut)
    * comp to both, alle perturbations haben r23 nicht in der optimization (auskommentierte zeile mit map_lgl(...,8)


gainlf = 0.5, alpha = 0
    * k7 k8 v8 führen in beiden communicating species zu r23 in der optimierung





























