---
title: "Enzymatic Cascade"
author: "Daniel Lill"
output:
  pdf_document:
    keep_tex: yes
---


```{r Initialization, results='hide', echo=FALSE, message=FALSE}
# Load libraries and define some useful paths
rm(list = ls())
library(MRAr)
library(scales)
library(xtable)
```


# Setup Model 

Careful, it might be broken due to the new comple Cx1z2 in the model...

## Prediction function
odemodel() writes the function f as a C-function, which is called by x and xs.
```{r Prediction function Cascade, results='asis'}
modelname   <- "Cascade"
f     <- as.eqnlist(read.csv(paste0(modelname, ".csv")))

# If the odemodel has been calculated already, it can be loaded with "load".
# If not, uncomment the next two lines.

# myodemodel <- odemodel(f, modelname = paste0(modelname), deriv = T)
# save(myodemodel, file = paste0("odemodel",modelname, ".RData"))
load(paste0("odemodel",modelname, ".RData"))

# Prediction function for continuous times
# x() is a function of time and parameters
x <- Xs(odemodel = myodemodel)
# Prediction function for steady states
xs <- Xs_steady(myodemodel) # When computing steady states, set deriv = F
loadDLL(x)

```



## Observation function
Observation function which returns our communicating species. 
The free parameters a_ij are introduced here to perform the linear combination of states which we wish to observe.
We chose to call them pars_opt, since they will be used to optimize the objective function later on.
```{r Most general observation function}
modules <- modules0 <- c("X","Y", "Z")

mypars_opt <- paste0("a_", outer(paste0("to", c("x","y","z"), "_C"), c("xy","yz","zx"), paste0)) %>% matrix(ncol=3)
mypars_opt[c(3,4,7,8,9)] <- "0"

mycomplexes <- matrix(c("Cx2y1", "Cy2z1", "Cx2z2"), ncol = 1)

samepars_opt <- paste0("a_", outer(rep("C",3), c("xy","yz","zx"), paste0)) %>% matrix(ncol=3)
samepars_opt[c(3,4,7,8,9)] <- "0"

obs <- obs0 <- paste(c("x2", "y2", "z2"), "+", prodSymb(mypars_opt, mycomplexes), "+", prodSymb(samepars_opt, mycomplexes))


g <- g0 <- Y(as.eqnvec(structure(obs0, names = modules0)), f, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```


## Inner parameters
Load all parameters and format them for further use.
```{r Parameters Cascade}

pars_raw <- read.csv(paste0("pars.csv"), header = FALSE)  # dynamical parameters
pars_raw <- structure(pars_raw$V2, names = as.character(pars_raw$V1))
ic_raw <- read.csv(paste0("IC.csv"), header = F)          # initial conditions
ic_raw <- structure(ic_raw$V2, names = as.character(ic_raw$V1))
# sort ics like they appear in x, so it can be used in stode()
vars_x <- attr(x, "parameters")[attr(x, "parameters") %in% names(ic_raw)]
ic_raw <- ic_raw[vars_x]

pars_inner_opt_0 <- setdiff(attr(g, "parameters"),names(c(pars_raw,ic_raw)))
pars_inner_opt_0 <- structure(rep(0, length(pars_inner_opt_0)), 
                              names = pars_inner_opt_0)

pars_inner_0 <- pars_inner <- c(ic_raw, pars_raw, pars_inner_opt_0)


# check if all parameters are assigned a value. Both lines should evaluate to the same number.
# x_pars <- unique(union((g) %>% attr("parameters"),(xs) %>% attr("parameters")))
# names(pars) %in% x_pars %>% sum()
# x_pars %in% names(pars) %>% sum()
```


## Parameter transformation function
```{r}
# Logtrafo of all pars
logtrafo <- structure(paste0("exp( log", names(pars_inner_0), ")"), 
                             names = names(pars_inner_0))

# Kill Synthesis and degradation. They can be used to study what happens, if proteins are synthesized and degraded.
syn_deg <- c(outer(c("ksyn", "kdeg"), c("x1","y1","z1"), FUN = paste0), 
              paste0("kdeg", c("x","y", "z"), "2"))
syn_deg <- structure(rep(0, length(syn_deg)), names = syn_deg)
logtrafo[names(syn_deg)] <- syn_deg

# Kill xz-feedback. This is part of a different study.
xz_fb <- c("Cy2z1", "k111", "k112", "k12")
xz_fb <- structure(rep(0, length(xz_fb)), names = xz_fb)
logtrafo[names(xz_fb)] <- xz_fb

# initial values of some complexes.
logtrafo[c("x2", "Cx2y1", "y2", "Cy2z1", "z2", "Cx2z2")] <- 0


p_log <- P(trafo = logtrafo,
           # compile = TRUE,
           modelname = "p_log")
           

# When we transform parameters that have the value 0 to log-pars we run into problems, because p() cannot cope with "-Inf". Therefore, we add a small number to them and trust that they are small enough to lie in the asymptotic range par -> -Inf.
pars_0 <- pars <- structure(log(pars_inner_0+.Machine$double.eps), names = paste0("log", names(pars_inner_0))) 

pars_opt_0 <- pars_opt <- structure(log(pars_inner_opt_0+.Machine$double.eps), names = paste0("log", names(pars_inner_opt_0)))

# pars_0
# p_log(pars_0, deriv = F)
```


# Simulate perturbation data

For the perturbation experiments we use another parameter transformation to scale some parameters.
p_pert_fun automatically generates a parameter transformation function, which returns different sets of parameters for different conditions and a control condition.
Eg. if we have pars_perturbed = c("logx1" = log(0.9)), p_pert_fun(pars_perturbed) will return a parameter transformation with the control condition with unchanged parameters and one condition in which logx' = logx + log(0.9)

"log(0.9)" corresponds to a ~10% reduction of the original parameter value.
```{r}
# Which pars shall be perturbed?
pars_perturbed_0 <- pars_perturbed <- c("logx1" = log(0.9), "logy1" = log(0.9), "logz1" = log(0.9))

# pars_perturbed
p_pert <- p_pert_fun(pars_perturbed = pars_perturbed, pars = pars)
compile(p_log, p_pert)
```

## Final parameter values for different Figures

Uncomment/comment the respective parameters that you want.

Parameters for Figures 3,4
```{r}
pars <- c(
x1       	=	100	, # 1
x2       	=	0	, # 2
Cx2y1    	=	0	, # 3
y1       	=	100	, # 4
y2       	=	0	, # 5
Cy2z1    	=	0	, # 6
z1       	=	100	, # 7
z2       	=	0	, # 8
Cx2z2    	=	0	, # 9
Cx1z2    	=	0	, # 10
k1       	=	100	, # 11
v1       	=	1.5	, # 12
k2       	=	100	, # 13
v2       	=	1	, # 14
k31      	=	0.01	, # 15
k32      	=	0.4	, # 16
k4       	=	0.05	, # 17
k5       	=	100	, # 18
v5       	=	5	, # 19
k61      	=	0.05	, # 20
k62      	=	0.4	, # 21
k7       	=	0.008	, # 22
k8       	=	100	, # 23
v8       	=	1	, # 24
k9       	=	100	, # 25
k10      	=	100	, # 26
gainuf   	=	1	, # 27
gainlf   	=	1	, # 28
ksynx1   	=	0	, # 29
kdegx1   	=	0	, # 30
kdegx2   	=	0	, # 31
ksyny1   	=	0	, # 32
kdegy1   	=	0	, # 33
kdegy2   	=	0	, # 34
ksynz1   	=	0	, # 35
kdegz1   	=	0	, # 36
kdegz2   	=	0	, # 37
k111     	=	0	, # 38
k112     	=	0	, # 39
k12      	=	0	, # 40
k131     	=	0	, # 41
k132     	=	0	, # 42
k14      	=	0	, # 43
a_tox_Cxy	=	0	, # 44
a_Cxy    	=	0	, # 45
a_toy_Cxy	=	0	, # 46
a_toy_Cyz	=	0	, # 47
a_Cyz    	=	0	, # 48
a_toz_Cyz	=	0	  # 49 
)

pars_0 <- pars <- log(pars+.Machine$double.eps) %>% set_names(names(pars_0))
```

Parameters for Figure 5
```{r}
# pars <- c(100,0,0,100,0,0,100,0,0,
#           100,1.5,100,1,     0.002,0.4,0.05,
#           100,5,            0.005,0.4,0.05,
#           100,1,
#           100,100,1,1,
#           0,0,0,0,0,0,00,0,0,0,0,0,
#           rep(0, length(pars_opt_0))) %>% set_names(names(pars_inner_0))
# pars_0 <- pars <- log(pars+.Machine$double.eps) %>% set_names(names(pars_0))
```



## Generate the perturbation data
```{r}
perturbation_prediction <-  (xs*p_log*p_pert)(times = c(0,Inf), pars = pars, deriv = F)

(x*p_log*p_pert)(times = seq(0,2000, length.out = 50), pars = pars) %>%
  plotPrediction()

perturbation_prediction %>% lapply(round,1)
```


# Apply the algorithm
This is an example of how the algorithm is applied.

Infer the network without complexes
```{r}
r_0 <-   R_fun(pars_opt = pars_opt_0,
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_0%>% round(2)
```

We need to consider complexes between X and Y and between Y and Z.
Scan through the parameters to choose which elements to keep in the optimization.

Often it suffices to compare two points of the parameter space, pars_opt = a_ij = 0 and pars_opt ~ 1.
```{r}
which_pars_opt <- c("loga_Cxy", "loga_Cyz")
r_kept_0 <- r_kept_fun(pars_opt = pars_opt[which_pars_opt],
                       perturbation_prediction = perturbation_prediction, 
                       obs_fun = g,
                       p_fun = (p_log * p_pert), 
                       pars = pars,
                       alpha = -log(0+.Machine$double.eps))
r_kept_0
```

Evaluate the objective function to demonstrate how it works.
```{r}
# pars_opt <- rep(0,2) %>% set_names(c("loga_Cxy", "loga_Cyz"))
# obj_alpha(pars_opt, fixed = NULL, perturbation_prediction = perturbation_prediction, r_kept = r_kept_0, obs_fun = g, p_fun = (p_log * p_pert), mypars = pars_0)
```

Optimize the objective function with trust region optimizer starting from multiple random starting points in our parameter space and take a look at the fits.

This method combined with the trust-region optimizer has proven to return find the optimum most reliably compared to other optimizers like optim with various different options.
```{r}
myfits <- mstrust(obj_alpha, center =  structure(rep(0, length(which_pars_opt)), names = which_pars_opt), studyname = "Fits", cores = 3, fits = 3, sd = 1, mypars = pars, perturbation_prediction = perturbation_prediction, r_kept = r_kept_0)

myfits %>% as.parframe()
```

Look at the resulting local response matrix, when the parameter values of the best fit are fed into our parametric local response matrix
```{r}
best_fit <- myfits %>% as.parframe() %>% as.parvec()
r_alpha_fun(pars_opt = unclass(best_fit), pars = pars, perturbation_prediction = perturbation_prediction) %>% round(2)
```

# Run analaysis for different parameters

```{r}
names(pars_inner)
remove_c_and_o()
```


```{r}
myframe <- dMod.frame("1", g, x, p_log, NULL, NULL, p_pert = list(p_pert), xs = list(xs), pars = list(pars))

pp_x <- c("x1", "k1", "v1", "k2", "v2", "k31", "k32", "k4") %>% paste0("log", .)
pp_y <- c("y1", "k5", "v5", "k31", "k32", "k4", "k61", "k62", "k7") %>% paste0("log", .)
pp_z <- c("z1", "k8", "v8", "k61", "k62", "k7") %>% paste0("log", .)

mygrid <- expand.grid(pp_x, pp_y, pp_z, stringsAsFactors = F) 
pp_list <- map(1:nrow(mygrid), function(i) {
  mynames <- mygrid[i,] %>% unlist() %>% unname() %>% unique()
  if(length(mynames)<3) return(NULL)
  mynames %>% are_names_of(log(0.9))
}) %>% 
  .[!map_lgl(., is.null)]
```

```{r eval=FALSE}
alpha <- 0

results <- map(pp_list, function(pp) {
  p_pert <- p_pert_fun(pp, pars = pars, modelname = NULL)
  
  myframe <- dMod.frame("1", g, x, p_log, NULL, NULL, p_pert = list(p_pert), xs = list(xs), pars = list(pars))
  
  g <- myframe$g[[1]]
  p_pert <- myframe$p_pert[[1]]
  p <- myframe$p[[1]]
  xs <- myframe$xs[[1]]
  pars <- myframe$pars[[1]]
  
  # browser()
      perturbation_prediction <- (xs*p*p_pert)(c(0,Inf), pars, deriv = F)
      
      try({r0 <- R_fun(pars_opt = pars_opt,
                perturbation_prediction = perturbation_prediction,
                obs_fun = g,
                p_fun = (p*p_pert),
                pars = pars) %>% local_response_matrix_eq10()})
      if(inherits(r0, "try-error")) return(NULL)
      
      algo <- function(which_pars_opt, prefix = NULL) {
      rkept <- r_kept_fun(pars_opt = pars_opt[which_pars_opt],
                                 perturbation_prediction = perturbation_prediction,
                                 obs_fun = g,
                                 p_fun = (p * p_pert),
                                 pars = pars,
                                 alpha =
                                   -log(.Machine$double.eps) + alpha)
      myfits <- mstrust(obj_alpha,
                        center =  structure(rep(0, length(which_pars_opt)), names = which_pars_opt),
                        studyname = "Fits",
                        cores = 3,
                        fits = 3,
                        sd = 1,
                        mypars = pars,
                        perturbation_prediction = perturbation_prediction,
                        r_kept = rkept,
                        p_fun = (p * p_pert),
                        obs_fun = g)

      try(myfits %>% as.parframe()) %>%
      {e <- .; if (inherits(e, "try-error")) print(myfits)}


      ropt <- r_alpha_fun(pars_opt = myfits %>% as.parframe() %>% as.parvec() %>% unclass() ,
                         pars = pars,
                         perturbation_prediction = perturbation_prediction,
                         p_fun = (p*p_pert))
      
      out <- list(rkept, myfits %>% as.parframe,ropt %>% round(2))
      names(out) <- paste0(prefix, "r_kept", "parframes", "ropt")
      return(out)
      }
      
      to_upstream_module <- algo(c("loga_tox_Cxy", "loga_toy_Cyz"), "upstream")
      to_both_modules <- algo(c("loga_Cxy", "loga_Cyz"), "both")
      
      out <- c(list(r0 = r0 %>% round(2)), to_upstream_module, to_both_modules)

})
# saveRDS(results, "test_different_perturbations.rds")
```

```{r}
results[[1]] %>% names

ropt_upstream <- results %>% map(4) 

mygrid_2 <-
  pp_list %>% map(names) %>% 
  do.call(rbind,.) %>% 
  as.tibble() %>% 
  cbind(., ind = 1:nrow(.)) %>% 
  filter(!str_detect(V1, "(k3)|(k4)")) %>% 
  filter(!str_detect(V2, "(k6)|(k7)"))

map(mygrid_2$ind, function(i) {
  pp_list[[i]] %>% names %>% print
  print(ropt_upstream[[i]])
})
```




# Plots for main text
Run the algorithm for different values of the parameters which control the feedback.
Do this for the two different options of defining communicating species (adding them to both modules or only the upstream module).
Also, scan the pars_opt in a finer grid than in the section above.
```{r, eval=FALSE}
par_opt_names <- pars_opt_0 %>% names %>% sort

par_opt_settings <- list(par_opt_names[c(1,2)]
                         , par_opt_names[c(3,5)]
                         )

```

```{r}
# computations for Figure 4
pars <- c(100,0,0,100,0,0,100,0,0,
          100,1.5,100,1,     0.01,0.4,0.05,
          100,5,            0.05,0.4,0.008,
          100,1,
          100,100,1,1,
          0,0,0,0,0,0,00,0,0,0,0,0,
          rep(0, length(pars_opt_0))) %>% set_names(names(pars_inner_0))


pars_0 <- pars <- log(pars+.Machine$double.eps) %>% set_names(names(pars_0))

dr_list <- lapply(seq_along(par_opt_settings), function(i) {
  # list(
  r_opt_dose_response(which_par = "loggainuf",
                      dosages = sort(unique(c(seq(-3,3,by = 0.05)))),
                      pars = pars,
                      alpha_scan_range = seq(-3,3,by = 0.2),
                      pars_opt = pars_opt_0[par_opt_settings[[i]]],
                      pars_opt_default_value = 0)
  # )
})
```

```{r, eval=FALSE}
# computations for Figure 5
# pars <- c(100,0,0,100,0,0,100,0,0,
#           100,1.5,100,1,     0.002,0.4,0.05,
#           100,5,            0.005,0.4,0.05,
#           100,1,
#           100,100,1,1,
#           0,0,0,0,0,0,00,0,0,0,0,0,
#           rep(0, length(pars_opt_0))) %>% set_names(names(pars_inner_0))
# pars_0 <- pars <- log(pars+.Machine$double.eps) %>% set_names(names(pars_0))
# 
# dr_list2 <- lapply(seq_along(par_opt_settings), function(i) {
#   r_opt_dose_response(which_par = "loggainlf",
#                       dosages = sort(unique(c(seq(-3,3,by = 0.05)))),
#                       pars = pars,
#                       alpha_scan_range = seq(-3,3,by = 0.2),
#                       pars_opt = pars_opt_0[par_opt_settings[[i]]],
#                       pars_opt_default_value = 0)
# })
```

```{r, eval=FALSE}
# save(dr_list,dr_list2, file = "Output/datasets_for_plotting.RData")
```



# Plot the results

```{r, eval=FALSE}
load("Output/datasets_for_plotting.RData")
```


## Figure 4
Plot as dose-response with alpha as facets
```{r Plot generation}
# load("Output/2017-07-24 dr_lists large complex concentrations.RData")
# load("Output/dr_lists Cascade nice pars.RData")
# load("Output/dr_list3 Cascade nice pars given r_kept.RData")

myplot <- c(dr_list,dr_list2) %>% 
  # dr_list3 %>% 
  combine_dr_list() %>%
  filter(! (r_element %in% paste0("r", c(11,22,33)))) %>% 
  filter(dose >= -2, dose <= 2.5) %>%
  filter(par_opt_setting == 1) %>% 
  filter(alpha %in% 0) %>%
  filter(matrix != "r_kept") %>%
  {
    .$matrix[.$matrix == "r_alpha"] <- "r(a_i = 1)"
    .$matrix[.$matrix == "r_opt"] <- "r(a = a_opt)"
    .$matrix <- .$matrix %>% factor(levels = c("r(a_i = 1)", "r(a = a_opt)", "r_kept"))
    .
  } %>% 
  mutate(Element = r_element) %>% 
  ggplot(mapping = aes(x = dose, y = value, color = Element, linetype = Element)) +
  geom_line() +
  theme_dMod() +
  scale_color_dMod() +
  geom_abline(slope = 0,intercept = 0, linetype = 3) +
  geom_vline(xintercept = 0, linetype = 3, size = 0.1) +
  facet_grid(matrix~., scales = "free")+
  xlab("log(gu)")

myplot

# look at them interactively with plotly:
# plotly::ggplotly(myplot)

# ggsave(plot = myplot,
#        filename = paste0("gu large sequestration",".png"),
#        width = 10,
#        height = 10,
#        units = "cm",
#        device = "png")

```


## Figure 5
Problem with Figure 5 is that if the r23 element is kicked out and we have random multistart-fits, the curve of r23 doesn't look smooth anymore (since it's basically random). What do we do?
```{r Plot generation}
myplot <- c(dr_list,dr_list2) %>% 
  combine_dr_list() %>%
  filter(! (r_element %in% paste0("r", c(11,22,33)))) %>% 
  filter(dose >= -2, dose <= 2.5) %>%
  filter(par_opt_setting == 3) %>% 
  filter(alpha %in% 0) %>%
  # filter(matrix != "r_kept") %>%
  { .$matrix[.$matrix == "r_alpha"] <- "r(a_i = 1)"
    .$matrix[.$matrix == "r_opt"] <- "r(a = a_opt)"
    .$matrix <- .$matrix %>% factor(levels = c("r(a_i = 1)", "r(a = a_opt)", "r_kept"))
    .
  } %>% 
  mutate(Element = r_element) %>% 
  ggplot(mapping = aes(x = dose, y = value, color = Element, linetype = Element)) +
  geom_line() +
  theme_dMod() +
  scale_color_dMod() +
  geom_abline(slope = 0,intercept = 0, linetype = 3) +
  geom_vline(xintercept = 0, linetype = 3, size = 0.1) +
  facet_grid(matrix~., scales = "free")+
  xlab("log(gl)")

myplot

# look at them interactively with plotly:
# plotly::ggplotly(myplot)

# ggsave(plot = myplot,
#        filename = paste0("gl small sequestration",".png"),
#        width = 10,
#        height = 10,
#        units = "cm",
#        device = "png")

```

```{r Plot generation}
myplot <- c(dr_list,dr_list2) %>% 
  combine_dr_list() %>%
  filter(! (r_element %in% paste0("r", c(11,22,33)))) %>% 
  # filter(dose >= -2, dose <= 2.5) %>%
  # filter(near(dose,0.3)) %>% 
  filter(par_opt_setting == 3) %>% 
  # filter(alpha %in% 0) %>%
  # filter(matrix != "r_kept") %>%
  { .$matrix[.$matrix == "r_alpha"] <- "r(a_i = 1)"
    .$matrix[.$matrix == "r_opt"] <- "r(a = a_opt)"
    .$matrix <- .$matrix %>% factor(levels = c("r(a_i = 1)", "r(a = a_opt)", "r_kept"))
    .
  } %>% 
  mutate(Element = r_element) %>% 
  ggplot(mapping = aes(x = alpha, y = value, color = Element, linetype = Element, frame = dose)) +
  geom_line() +
  theme_dMod() +
  scale_color_dMod() +
  geom_abline(slope = 0,intercept = 0, linetype = 3) +
  geom_vline(xintercept = 0, linetype = 3, size = 0.1) +
  facet_grid(matrix~., scales = "free")+
  xlab("log(gl)")

# myplot

# look at them interactively with plotly:
plotly::ggplotly(myplot)

# ggsave(plot = myplot,
#        filename = paste0("gl small sequestration",".png"),
#        width = 10,
#        height = 10,
#        units = "cm",
#        device = "png")

```