---
title: "Enzymatic Cascade"
author: "Daniel Lill"
output:
  pdf_document:
    keep_tex: yes
---

# Load libraries and define some useful paths
```{r setup, results='hide', echo=FALSE, message=FALSE}
library(parallel)
library(MRAr)
library(xtable)
```

# Model setup

## ODE model setup
```{r Prediction function Cascade, results='asis'}
reactions <- NULL %>% 
  addReaction(from = "x1", to = "x2", rate = "((v1 *  x1)/(k1 + x1))", description = "act_x") %>% 
  addReaction(from = "x2", to = "x1", rate = "((v2 * x2 )/(k2 + x2))", description = "deact_x") %>% 
  addReaction(from = "x2 + y1", to = "Cx2y1", rate = "(k31 * y1 *x2) - (k32 *Cx2y1)", description = "x2_y_to_comp") %>% 
  addReaction(from = "Cx2y1", to = "x2 + y2", rate = "(k4 * Cx2y1)", description = "y2_formation") %>% 
  addReaction(from = "y2", to = "y1", rate = "((v5 * y2)/(k5 + y2))", description = "deact_y") %>% 
  addReaction(from = "y2 + z1", to = "Cy2z1", rate = "((k61 *z1 *y2) - (k62 * Cy2z1))", description = "y2_z_to_comp") %>% 
  addReaction(from = "Cy2z1", to = "y2 + z2", rate = "(k7 * Cy2z1)", description = "z2_formation") %>% 
  addReaction(from = "z2", to = "z1", rate = "((v8 * z2)/(k8 + z2))", description = "z_deact") 

compiled_model <- odemodel(reactions, modelname = "Cascade", estimate = c("x1", "k1"))

# Prediction function
x <- Xs(odemodel = compiled_model) 
# Prediction function for steady states
xs <- Xs_steady(compiled_model) # When computing steady states, set deriv = F
```

### Modules, their observables and the observation function
Observation function which depends also on the optimization parameters. With their help, the complexes can smoothly be switched on in the observation.

```{r Most general observation function}
modules <-  c("X", "Y", "Z")

alpha_parameters <- paste0("a_", outer(paste0("to", c("x","y","z"), "_C"), c("xy","yz","zx"), paste0)) %>% matrix(ncol=3)
alpha_parameters[c(3,4,8)] <- "0"

complexes <- matrix(c("Cx2y1", "Cy2z1", "Cx2z2"), ncol = 1)

alpha_parameters_to_both_modules <- paste0("a_", outer(rep("C",3), c("xy","yz","zx"), paste0)) %>% matrix(ncol=3)
alpha_parameters_to_both_modules[c(3,4,8)] <- "0"

observables <- paste(c("x2", "y2", "z2"), "+", prodSymb(alpha_parameters, complexes), "+", prodSymb(alpha_parameters_to_both_modules, complexes))


g <- Y(g = as.eqnvec(structure(observables, names = modules)), 
       f = reactions, 
       compile = TRUE, 
       modelname = "obsfn0", 
       attach.input = FALSE)
```


### General log transformation and setting some pars to zero
It's better to do a logtransformation for all Fitting purposes and work with the log-Parameters.
```{r}
trafo <- getParameters(x,g) %>% 
  setNames(.,.) %>% 
  insert("x~0", x = str_subset(.currentSymbols, "^C*[xyz](2|1.2)")) %>% # set inits to zero
  insert("x~exp(x)", x = .currentSymbols) %>%
  # {.} %T>% print %>% 
  {.} 

p_log <- P(trafo = trafo,
           modelname = "p_log",
           compile = T)
```


```{r}
pars <- c(
  a_Cxy    	=	.Machine$double.eps	, # 1
  a_Cyz    	=	.Machine$double.eps	, # 2
  a_Czx    	=	.Machine$double.eps	, # 3
  a_tox_Cxy	=	.Machine$double.eps	, # 4
  a_tox_Czx	=	.Machine$double.eps	, # 5
  a_toy_Cxy	=	.Machine$double.eps	, # 6
  a_toy_Cyz	=	.Machine$double.eps	, # 7
  a_toz_Cyz	=	.Machine$double.eps	, # 8
  a_toz_Czx	=	.Machine$double.eps	, # 9
  
  x1 = 100,      # 30
  y1 = 100,      # 31
  z1 = 100,       # 32

  k1       	=	100	, # 11
  v1       	=	1.5	, # 12
  k2       	=	100	, # 13
  v2       	=	1	, # 14
  k31      	=	0.01	, # 15
  k32      	=	0.4	, # 16
  k4       	=	0.05	, # 17
  k5       	=	100	, # 18
  v5       	=	5	, # 19
  k61      	=	0.05	, # 20
  k62      	=	0.4	, # 21
  k7       	=	0.008	, # 22
  k8       	=	100	, # 23
  v8       	=	1	
) %>% 
  log

alpha_pars <- pars[str_detect(names(pars), "^a_")]
```

```{r}
# test the algorithm.
test <- algo(pars, c("x1", "y1", "z1"), c("a_tox_Cxy", "a_toy_Cyz"), alpha_pars, 0, g, xs, p_log)
unlist(test, recursive = F)
```

# Apply different perturbations and run the algorithm
Since all possible combinations of paramter perturbations are many, these computations are done in a parallelized manner on a remote server.

```{r Prepare computations}
perturbations_module_x <- c("x1", "k1", "v1", "k2", "v2")
perturbations_module_y <- c("y1", "k5", "v5", "k31", "k32", "k4")
perturbations_module_z <- c("z1", "k8", "v8", "k61", "k62", "k7")
perturbation_combinations <- expand.grid(perturbations_module_x, perturbations_module_y, perturbations_module_z, stringsAsFactors = F) 

perturbations <- map(1:nrow(perturbation_combinations), function(i) {
  mynames <- perturbation_combinations[i,] %>% unlist() %>% unname() %>% unique()
  if(length(mynames)<3) return(NULL)
  mynames %>% are_names_of(log(0.9))
}) %>% 
  .[!map_lgl(., is.null)]

dose_pars <- pars["x1"]
alphas <- c(-3, -1, 0, 1, 3)
alpha_par_settings <- list(upstream = c("a_tox_Cxy", "a_toy_Cyz"),
                           both = c("a_Cxy", "a_Cyz"))

```

If you quickly want to check, how the algorithm works, set run_subset to T
```{r}
run_subset <- F
if (run_subset) {
  perturbations <- perturbations[1:5]
  alphas <- alphas[c(3,5)]
  alpha_par_settings <- alpha_par_settings[1]
}
```

Send the computations via ssh to a cluster or a server. Requires the R-packages to be installed there as well and that ssh is set up to use aliases. Replace "localhost" by the respective ssh alias. See ?runbg for more details.
```{r}
ssh_machine <- "localhost"

remote_computing_job <- runbg({
  cores <- detectFreeCores()
  results <- run_different_perturbations(dose_pars = dose_pars, 
                                         pars0 = pars, 
                                         alpha_pars = alpha_pars,
                                         alphas = alphas, 
                                         perturbations = perturbations,
                                         xs = xs,
                                         p_log = p_log,
                                         p_pert = p_pert,
                                         g = g, 
                                         cores = cores,
                                         alpha_par_settings = alpha_par_settings)
  results 
}, machine = ssh_machine, input = ls(pos=.GlobalEnv), filename = "cascade"
# , recover = T
)
```

Check if results are ready.
```{r}
while(!remote_computing_job$check()) {
  Sys.sleep(60)
}
```

Fetch/save/restore results
```{r}
results <- remote_computing_job$get()[[1]]
saveRDS(results, file = "results_cascade.rds")
results <- readRDS("results_cascade.rds")
```

Purge files from the remote computing job.
```{r}
remote_computing_job$purge()
```


# Explore results
Cluster matrices by similarity across different perturbations.
This section assumes that all perturbations have been computed (run_subset == F).
```{r Prepare data for clustering}
results_unnested_matrices <- results %>%
  mutate(r_optupstream = map(r_optupstream, matrix_2_wide_df, nm = (paste0("upstream_ropt", outer(1:3,1:3, paste0)))),
         r_optboth = map(r_optboth, . %>% matrix_2_wide_df(paste0("both_ropt", outer(1:3,1:3, paste0)))),
         
         r_keptupstream = map(r_keptupstream, matrix_2_wide_df, nm = (paste0("upstream_rkept", outer(1:3,1:3, paste0)))),
         r_keptboth = map(r_keptboth, . %>% matrix_2_wide_df(paste0("both_rkept", outer(1:3,1:3, paste0)))),
         
         r_1upstream = map(r_1upstream, matrix_2_wide_df, nm = (paste0("upstream_r1", outer(1:3,1:3, paste0)))),
         r_1both = map(r_1both, . %>% matrix_2_wide_df(paste0("both_r1", outer(1:3,1:3, paste0)))),
         
         r_0 = map(r_0, . %>% matrix_2_wide_df(paste0("r_0", outer(1:3,1:3, paste0)))),
         
         pars_perturbed = map(pars_perturbed, . %>% names %>% `names<-`(LETTERS[24:26]) %>% t %>% as.data.frame(stringsAsFactors = F)),
         perturbed_set = seq_along(alpha),
         alpha = as.numeric(alpha)) %>% 
  unnest(r_optupstream, r_optboth,r_keptupstream, r_keptboth,r_1upstream, r_1both, r_0, pars_perturbed)
```

There are about 4 different clusters of local response matrices when no complexes are measured:
```{r}
cluster_by_r_0 <- cluster_matrices_and_print_perturbations(results_unnested_matrices, 0, "r_0", 4,NULL)
```
Also 4 clusters when total phosphorylated species are measured.
```{r}
cluster_by_r_1 <- cluster_matrices_and_print_perturbations(results_unnested_matrices, 0, "upstream_r1", 4,NULL)
```

When the complexes are measured independently, the matrices are similar after detecting the sequestration elements and optimizing the free parameters to minimize these elements.
```{r}
cluster_by_ropt_complex_to_upstream <- cluster_matrices_and_print_perturbations(results_unnested_matrices, 0, "upstream_ropt", 2,NULL)
```

When the complex is measured independently, but added to both modules, some combinations of parameter perturbations don't work out anymore, but most of them do.
```{r}
clustered <- cluster_by_ropt_complex_to_both <- cluster_matrices_and_print_perturbations(results_unnested_matrices, 0, "both_ropt", 5,NULL)
```



# Export of results

## Parameters/ODEs
```{r}
reactions %>% 
  as.eqnvec() %>% 
  print2word_eqnvec("odes_cascade")
```

```{r}
pars %>% 
  exp %>% 
  print2word_namedvec("pars_cascade", names_in_mathenv = T)
```

## Matrices
Export some tables for the publication
```{r}
to_export <-  results %>% 
  filter(alpha == 0) %>% 
  filter(map_lgl(pars_perturbed, ~identical(names(.x), c("x1", "y1", "z1")))) %>% 
  unlist(recursive = F)

to_export[c("r_0", "r_1upstream", "r_optupstream")] %>% 
  print2word_matrixlist("cascade_matrices")
```

```{r}
to_export <-  results %>% 
  filter(alpha == 0) %>% 
  filter(map_lgl(pars_perturbed, ~identical(names(.x), c("x1", "y1", "z1")))) %>% 
  unlist(recursive = F)

to_export[c("r_optupstream", "r_optboth")] %>% 
 print2word_matrixlist2("cascade_matrices_SI")
```



```{r}
cluster_by_ropt_complex_to_upstream <- cluster_matrices_and_print_perturbations(results_unnested_matrices, 0, "upstream_ropt", 2,NULL)
set.seed(2)
perturbations <- sort(c(1,sample(2:180, 9)))

to_export <- results %>% 
  filter(alpha == 0) %>% 
  .[perturbations,]

r_opt_export <- to_export$r_optupstream %>% setNames(paste0("Perturbed:", 
                                                            map_chr(to_export$pars_perturbed,~paste0(names(.x), collapse = ","))))
print2word_matrixlist2(r_opt_export, "cascade_ropt_many_different_perturbations")
r_0_export <- to_export$r_0 %>% setNames(paste0("Perturbed:", 
                                                            map_chr(to_export$pars_perturbed,~paste0(names(.x), collapse = ","))))
print2word_matrixlist2(r_0_export, "cascade_r0_many_different_perturbations")
x <- r_opt_export

# Call Script S7
```



# Clean up working directory
```{r}
unlink(paste0("*.", c("c", "o", "so", "tex")))
unlink("Fits/", T)
```




























