---
title: "Cascade - x2z2 mass action feedback"
author: "Daniel Lill"
output:
  pdf_document:
    keep_tex: yes
---

# Load libraries and define some useful paths
```{r setup, results='hide', echo=FALSE, message=FALSE}
library(xtable)
library(parallel)
library(MRAr)
```

# Model setup

## ODE model setup
```{r Prediction function Cascade, results='asis'}
reactions <- NULL %>% 
  addReaction(from = "x1", to = "x2", rate = "((v1 *  x1)/(k1 + x1))", description = "act_x") %>% 
  addReaction(from = "x2", to = "x1", rate = "((v2 * x2 )/(k2 + x2))", description = "deact_x") %>% 
  addReaction(from = "x2 + y1", to = "Cx2y1", rate = "(k31 * y1 *x2) * (1 + u2*z2/k10)/(1 + z2/k10) - (k32 *Cx2y1)", description = "x2_y_to_comp") %>% 
  addReaction(from = "Cx2y1", to = "x2 + y2", rate = "(k4 * Cx2y1)", description = "y2_formation") %>% 
  addReaction(from = "y2", to = "y1", rate = "((v5 * y2)/(k5 + y2))", description = "deact_y") %>% 
  addReaction(from = "y2 + z1", to = "Cy2z1", rate = "((k61 *z1 *y2) - (k62 * Cy2z1))", description = "y2_z_to_comp") %>% 
  addReaction(from = "Cy2z1", to = "y2 + z2", rate = "(k7 * Cy2z1)", description = "z2_formation") %>% 
  addReaction(from = "z2", to = "z1", rate = "((v8 * z2)/(k8 + z2))", description = "z_deact") %>% 
  addReaction(from = "x2 + z2", to = "Cx2z2", rate = "k111*x2*z2-k112*Cx2z2", description = "z2_x2_to_comp") %>% 
  addReaction(from = "Cx2z2", to = "x1 + z2", rate = "k12*Cx2z2", description = "deact_x_by_z2") 

compiled_model <- odemodel(reactions, modelname = "Cascade", estimate = c("x1", "k1"))

# Prediction function
x <- Xs(odemodel = compiled_model) 
# Prediction function for steady states
xs <- Xs_steady(compiled_model) # When computing steady states, set deriv = F
```

### Modules, their observables and the observation function
Observation function which depends also on the optimization parameters. With their help, the complexes can smoothly be switched on in the observation.

```{r Most general observation function}
modules <-  c("X", "Y", "Z")

alpha_parameters <- paste0("a_", outer(paste0("to", c("x","y","z"), "_C"), c("xy","yz","zx"), paste0)) %>% matrix(ncol=3)
alpha_parameters[c(3,4,8)] <- "0"

complexes <- matrix(c("Cx2y1", "Cy2z1", "Cx2z2"), ncol = 1)

alpha_parameters_to_both_modules <- paste0("a_", outer(rep("C",3), c("xy","yz","zx"), paste0)) %>% matrix(ncol=3)
alpha_parameters_to_both_modules[c(3,4,8)] <- "0"

observables <- paste(c("x2", "y2", "z2"), "+", prodSymb(alpha_parameters, complexes), "+", prodSymb(alpha_parameters_to_both_modules, complexes))


g <- Y(g = as.eqnvec(structure(observables, names = modules)), 
       f = reactions, 
       compile = TRUE, 
       modelname = "obsfn0", 
       attach.input = FALSE)
```


### General log transformation and setting some pars to zero
It's better to do a logtransformation for all Fitting purposes and work with the log-Parameters.
```{r}
trafo <- getParameters(x,g) %>% 
  setNames(.,.) %>% 
  insert("x~0", x = str_subset(.currentSymbols, "^C*[xyz](2|1.2)")) %>% # set inits to zero
  insert("x~exp(x)", x = .currentSymbols) %>%
  # {.} %T>% print %>% 
  {.} 

p_log <- P(trafo = trafo,
           modelname = "p_log",
           compile = T)
```


```{r}
pars <- c(
a_Cxy    	=	.Machine$double.eps	, # 1
a_Cyz    	=	.Machine$double.eps	, # 2
a_Czx    	=	.Machine$double.eps	, # 3
a_tox_Cxy	=	.Machine$double.eps	, # 4
a_tox_Czx	=	.Machine$double.eps	, # 5
a_toy_Cxy	=	.Machine$double.eps	, # 6
a_toy_Cyz	=	.Machine$double.eps	, # 7
a_toz_Cyz	=	.Machine$double.eps	, # 8
a_toz_Czx	=	.Machine$double.eps	, # 9
k1       	=	100	, # 12
v1       	=	1.5	, # 30
k2       	=	100	, # 20
v2       	=	1	, # 31
k31      	=	0.002	, # 21
k32      	=	0.4	, # 22
k4       	=	0.05	, # 23
k5       	=	100	, # 24
v5       	=	5	, # 32
k61      	=	0.005	, # 25
k62      	=	0.4	, # 26
k7       	=	0.05	, # 27
k8       	=	100	, # 28
v8       	=	1	, # 33
k111     	=	0.002	, # 14
k112     	=	0.4	, # 15
k12      	=	0.05	, # 16

x1       	=	100	, # 34
y1       	=	100	, # 35
z1       	=	100	  # 36 
) %>% 
  log

alpha_pars <- pars[str_detect(names(pars), "^a_")]
```

```{r}
# test the algorithm.
test <- algo(pars, c("x1", "y1", "z1"), c("a_tox_Cxy", "a_toy_Cyz", "a_toz_Czx"), alpha_pars, 0, g, xs, p_log)
unlist(test[-4], recursive = F)
```

# Apply different perturbations and run the algorithm
Since all possible combinations of paramter perturbations are many, these computations are done in a parallelized manner on a remote server.

```{r Prepare computations}
perturbations_module_x <- c("x1", "k1", "v1", "k2", "v2")
perturbations_module_y <- c("y1", "k5", "v5", "k31", "k32", "k4")
perturbations_module_z <- c("z1", "k8", "v8", "k61", "k62", "k7", "k111", "k112", "k12")
perturbation_combinations <- expand.grid(perturbations_module_x, perturbations_module_y, perturbations_module_z, stringsAsFactors = F) 

perturbations <- map(1:nrow(perturbation_combinations), function(i) {
  mynames <- perturbation_combinations[i,] %>% unlist() %>% unname() %>% unique()
  if(length(mynames)<3) return(NULL)
  mynames %>% are_names_of(log(0.9))
}) %>% 
  .[!map_lgl(., is.null)]

dose_pars <- pars["x1"]
alphas <- c(0, 1, 3)
alpha_par_settings <- list(upstream = c("a_tox_Cxy", "a_toy_Cyz", "a_toz_Czx"),
                           both = c("a_Cxy", "a_Cyz", "a_Czx"))

```

If you quickly want to check, how the algorithm works, set run_subset to F
```{r}
run_subset <- F
if (run_subset) {
  perturbations <- perturbations[1:5]
  dose_pars <- dose_pars[c(1,4,5)]
  alphas <- alphas[c(3,5)]
  alpha_par_settings <- alpha_par_settings[1]
}
```

Send the computations via ssh to a cluster or a server. Requires the R-packages to be installed there as well and that ssh is set up to use aliases. Replace "localhost" by the respective ssh alias. See ?runbg for more details.
```{r}
ssh_machine <- "fermi"

remote_computing_job <- runbg({
  cores <- detectFreeCores()
  results <- run_different_perturbations(dose_pars = dose_pars, 
                                         pars0 = pars, 
                                         alpha_pars = alpha_pars,
                                         alphas = alphas, 
                                         perturbations = perturbations,
                                         xs = xs,
                                         p_log = p_log,
                                         p_pert = p_pert,
                                         g = g, 
                                         cores = cores,
                                         alpha_par_settings = alpha_par_settings)
  results 
}, machine = ssh_machine, input = ls(pos=.GlobalEnv), filename = "cascade_zx_complex_feedback"
# , recover = T
)
```

Check if results are ready.
```{r}
while(!remote_computing_job$check()) {
  Sys.sleep(60)
}
```

Fetch/save/restore results
```{r}
results <- remote_computing_job$get()[[1]]
saveRDS(results, file = "results_cascade_zx_complex_feedback.rds")
results <- readRDS("results_cascade_zx_complex_feedback.rds")
```

Purge files from the remote computing job.
```{r}
remote_computing_job$purge()
```

# Explore results
Cluster matrices by similarity across different perturbations.
This section assumes that all perturbations have been computed (run_subset == F).
```{r Prepare data for clustering}
results_unnested_matrices <- results %>%
  filter(!map_lgl(r_optupstream, is.null), 
         !map_lgl(r_optboth, is.null), 
         !map_lgl(r_1upstream, is.null), 
         !map_lgl(r_1both, is.null), 
         !map_lgl(r_keptupstream, is.null), 
         !map_lgl(r_keptboth, is.null)) %>% 
  mutate(r_optupstream = map(r_optupstream, matrix_2_wide_df, nm = (paste0("upstream_ropt", outer(1:3,1:3, paste0)))),
         r_optboth = map(r_optboth, . %>% matrix_2_wide_df(paste0("both_ropt", outer(1:3,1:3, paste0)))),
         
         r_keptupstream = map(r_keptupstream, matrix_2_wide_df, nm = (paste0("upstream_rkept", outer(1:3,1:3, paste0)))),
         r_keptboth = map(r_keptboth, . %>% matrix_2_wide_df(paste0("both_rkept", outer(1:3,1:3, paste0)))),
         
         r_1upstream = map(r_1upstream, matrix_2_wide_df, nm = (paste0("upstream_r1", outer(1:3,1:3, paste0)))),
         r_1both = map(r_1both, . %>% matrix_2_wide_df(paste0("both_r1", outer(1:3,1:3, paste0)))),
         
         r_0 = map(r_0, . %>% matrix_2_wide_df(paste0("r_0", outer(1:3,1:3, paste0)))),
         
         pars_perturbed = map(pars_perturbed, . %>% names %>% `names<-`(LETTERS[24:26]) %>% t %>% as.data.frame(stringsAsFactors = F)),
         perturbed_set = seq_along(alpha),
         alpha = as.numeric(alpha)) %>% 
  unnest(r_optupstream, r_optboth,r_keptupstream, r_keptboth,r_1upstream, r_1both, r_0, pars_perturbed)
```

```{r}
clustered_by_ropt <- results_unnested_matrices %>% 
  cluster_matrices_and_print_perturbations(0, "upstream_ropt", 4,1)
```



# Export of results

## Parameters/ODEs
```{r}
reactions %>% 
  as.eqnvec() %>% 
  print2word_eqnvec("odes_cascade_zx_complex_fb")
```

```{r}
pars %>% 
  exp %>% 
  print2word_namedvec("pars_cascade_zx_complex_fb", names_in_mathenv = T)
```

## Export matrices
```{r}
matrices_cs19 <- algo(pars, paste0(letters[24:26], "1"), c("a_tox_Cxy", "a_toy_Cyz","a_toz_Czx"), alpha_pars, 0, g, xs, p_log) %>% unlist(F) %>% .[-4]

print2word_matrixlist2(matrices_cs19, "cascade_zx_complex_feedback")
```

```{r}
matrices_cs20 <- algo(pars, paste0(letters[24:26], "1"), c("a_Cxy", "a_Cyz","a_Czx"), alpha_pars, 0, g, xs, p_log) %>% unlist(F) %>% .[-4]

print2word_matrixlist2(matrices_cs20, "cascade_zx_complex_feedback_CS20")
```

```{r}
clustered_by_ropt %>% 
  filter(.cluster == 1) %>% 
  select(X,Y,Z) %>% 
  write_csv("permissible_perturbations.csv")
```



## Clean up Workdir
```{r}
unlink(paste0("*.", c("c", "o", "so", "tex", "lnk")))
unlink("Fits/", T)
unlink(".pars_controlled/", T)
```

