---
title: "Enzymatic Cascade"
author: "Daniel Lill"
output:
  pdf_document:
    keep_tex: yes
---

# Load libraries and define some useful paths
```{r setup, results='hide', echo=FALSE, message=FALSE}
library(parallel)
library(MRAr)
library(xtable)

postProcess <- function(test) {
  test <- unlist(test, recursive = F)[c(1,2,3,5)]
  test <- lapply(test, function(x) cbind(c("x","y","z"), x ))
  test <- do.call(cbind, test)
  test <- data.table(test)
  cfoutput_MdTable(test, FLAGsummaryRow = FALSE)
}
```




# Model setup

## ODE model setup
```{r Prediction function Cascade, results='asis'}
reactions <- NULL %>% 
  addReaction(from = "x1", to = "x2", rate = "((v1 *  x1)/(k1 + x1))*(1 + u1*z2/k9)/(1 + z2/k9)", description = "act_x") %>% 
  addReaction(from = "x2", to = "x1", rate = "((v2 * x2 )/(k2 + x2))", description = "deact_x") %>% 
  addReaction(from = "x2 + y1", to = "Cx2y1", rate = "(k31 * y1 *x2) * (1 + u2*z2/k10)/(1 + z2/k10) - (k32 *Cx2y1) ", description = "x2_y_to_comp") %>% 
  addReaction(from = "Cx2y1", to = "x2 + y2", rate = "(k4 * Cx2y1)", description = "y2_formation") %>% 
  addReaction(from = "y2", to = "y1", rate = "((v5 * y2)/(k5 + y2))", description = "deact_y") %>% 
  addReaction(from = "y2 + z1", to = "Cy2z1", rate = "((k61 *z1 *y2) - (k62 * Cy2z1))", description = "y2_z_to_comp") %>% 
  addReaction(from = "Cy2z1", to = "y2 + z2", rate = "(k7 * Cy2z1)", description = "z2_formation") %>% 
  addReaction(from = "z2", to = "z1", rate = "((v8 * z2)/(k8 + z2))", description = "z_deact") 

compiled_model <- odemodel(reactions, modelname = "Cascade", estimate = c("x1", "k1"))

# Prediction function
x <- Xs(odemodel = compiled_model) 
# Prediction function for steady states
xs <- Xs_steady(compiled_model) # When computing steady states, set deriv = F
```

### Modules, their observables and the observation function
Observation function which depends also on the optimization parameters. With their help, the complexes can smoothly be switched on in the observation.

```{r Most general observation function}
modules <-  c("X", "Y", "Z")

alpha_parameters <- paste0("a_", outer(paste0("to", c("x","y","z"), "_C"), c("xy","yz","zx"), paste0)) %>% matrix(ncol=3)
alpha_parameters[c(3,4,8)] <- "0"

complexes <- matrix(c("Cx2y1", "Cy2z1", "Cx2z2"), ncol = 1)

alpha_parameters_to_both_modules <- paste0("a_", outer(rep("C",3), c("xy","yz","zx"), paste0)) %>% matrix(ncol=3)
alpha_parameters_to_both_modules[c(3,4,8)] <- "0"

observables <- paste(c("x2", "y2", "z2"), "+", prodSymb(alpha_parameters, complexes), "+", prodSymb(alpha_parameters_to_both_modules, complexes))


g <- Y(g = as.eqnvec(structure(observables, names = modules)), 
       f = reactions, 
       compile = TRUE, 
       modelname = "obsfn0", 
       attach.input = FALSE)
```


### General log transformation and setting some pars to zero
It's better to do a logtransformation for all Fitting purposes and work with the log-Parameters.
```{r}
trafo <- getParameters(x,g) %>% 
  setNames(.,.) %>% 
  insert("x~0", x = str_subset(.currentSymbols, "^C*[xyz](2|1.2)")) %>% # set inits to zero
  insert("x~exp(x)", x = .currentSymbols) %>%
  # {.} %T>% print %>% 
  {.} 

p_log <- P(trafo = trafo,
           modelname = "p_log",
           compile = T)
```


```{r}
pars <- c(
  a_Cxy    	=	.Machine$double.eps	, # 1
  a_Cyz    	=	.Machine$double.eps	, # 2
  a_Czx    	=	.Machine$double.eps	, # 3
  a_tox_Cxy	=	.Machine$double.eps	, # 4
  a_tox_Czx	=	.Machine$double.eps	, # 5
  a_toy_Cxy	=	.Machine$double.eps	, # 6
  a_toy_Cyz	=	.Machine$double.eps	, # 7
  a_toz_Cyz	=	.Machine$double.eps	, # 8
  a_toz_Czx	=	.Machine$double.eps	, # 9
  
  x1 = 100,      # 30
  y1 = 100,      # 31
  z1 = 100,       # 32
  
  u1   	    =	1	, # 10
  u2   	    =	1	, # 11
  k9       	=	50	, # 29
  k10      	=	50	, # 13

  k1       	=	100	, # 11
  v1       	=	1.5	, # 12
  k2       	=	100	, # 13
  v2       	=	1	, # 14
  
  # k31      	=	0.01	, # 15
  # k32      	=	0.4	, # 16
  # k4       	=	0.05	, # 17
  # k5       	=	100	, # 18
  # v5       	=	5	, # 19
  # k61      	=	0.05	, # 20
  # k62      	=	0.4	, # 21
  # k7       	=	0.008	, # 22
  
  # 3-1 and 3-2 feedback
  k31      	=	0.002	, # 21     # Different than above
  k32      	=	0.4	, # 22
  k4       	=	0.05	, # 23
  k5       	=	100	, # 24
  v5       	=	5	, # 32
  k61      	=	0.005	, # 25    # Different than above
  k62      	=	0.4	, # 26
  k7       	=	0.05	, # 27    # Different than above
  
  k8       	=	100	, # 23
  v8       	=	1	
  
  

  
) %>% 
  log

alpha_pars <- pars[str_detect(names(pars), "^a_")]
```



# Matrices for different feedbacks

Base parameterization
```{r}
test <- algo(pars, c("x1", "y1", "z1"), c("a_tox_Cxy", "a_toy_Cyz"), alpha_pars, 0, g, xs, p_log)
postProcess(test)
```

3-1 feedback
```{r}
mypars <- pars
mypars["u1"] <- log(0.5)
test <- algo(mypars, paste0(letters[24:26], "1"), c("a_tox_Cxy", "a_toy_Cyz"), alpha_pars, 0, g, xs, p_log)
postProcess(test)
```

Weak 3-2 feedback
```{r}
mypars <- pars
mypars["u2"] <- log(1.3)
test <- algo(mypars, paste0(letters[24:26], "1"), c("a_tox_Cxy", "a_toy_Cyz"), alpha_pars, 0, g, xs, p_log)
postProcess(test)
```
Strong 3-2 feedback
```{r}
mypars <- pars
mypars["u2"] <- log(3)
test <- algo(mypars, paste0(letters[24:26], "1"), c("a_tox_Cxy", "a_toy_Cyz"), alpha_pars, 0, g, xs, p_log)
postProcess(test)
```



# Clean up working directory
```{r}
unlink(paste0("*.", c("c", "o", "so", "tex")))
unlink("Fits/", T)
```




























