---
title: "MEK-ERK-Model"
author: "Daniel Lill"
output: 
  pdf_document:
    keep_tex: true
---

# Load libraries and define some useful paths

```{r setup, results='hide', echo=FALSE, message=FALSE}
library(MRAr)
library(scales)
library(xtable)
```

# Model Setup

## ODE model setup
```{r Prediction function}
reactions <- NULL %>% 
  addReaction(from = "m1", to = "m2", rate = "k1*m1", description = "act_m") %>% 
  addReaction(from = "m2", to = "m1", rate = "k2*m2", description = "deact_m") %>% 
  addReaction(from = "m2 + e1", to = "Cm2e1", rate = "(k31 * e1 *m2)", description = "m2_e1_to_comp") %>% 
  addReaction(from = "Cm2e1", to = "m2 + e1", rate = "(k32 *Cm2e1)", description = "comp_to_m2_e1") %>% 
  addReaction(from = "Cm2e1", to = "m2 + e2", rate = "(k4 * Cm2e1)", description = "e2_formation") %>% 
  addReaction(from = "e2 + p1", to = "Ce2p1", rate = "((k61 *p1 *e2))", description = "e2_p1_to_comp") %>% 
  addReaction(from = "Ce2p1", to = "e2 + p1", rate = "(k62 * Ce2p1)", description = "compt_to_e2_p1") %>% 
  addReaction(from = "Ce2p1", to = "e1 + p1", rate = "k5*Ce2p1", description = "e2_deact") %>% {.}

compiled_model <- odemodel(reactions, modelname = "Prabakaran", estimate = c("m1", "k1"))

# Prediction function
x <- Xs(odemodel = compiled_model)
# Prediction function for steady states
xs <- Xs_steady(compiled_model)
```


## Observation function
Observation function which returns our communicating species. 
The free parameters a_ij are introduced here to perform the linear combination of states which we wish to observe.
We chose to call them alpha_oars, since they will be used to optimize the objective function later on.
```{r Naive observation function}
modules <- c("X","Y")
observables <-  c("m2+a_xy*Cm2e1+a_tox_xy*Cm2e1",
                 "e2+a_xy*Cm2e1+a_toy_xy*Cm2e1")
g <- Y(as.eqnvec(structure(observables, names = modules)), x, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```


## Parameter transformation function
Define logtrafo, define initial conditions
```{r}
trafo <- getParameters(x,g) %>% 
  setNames(.,.) %>% 
  insert("x~0", x = str_subset(.currentSymbols, "^C*[me](2|1.2)")) %>% # set inits to zero
  insert("x~exp(x)", x = .currentSymbols)

p_log <- P(trafo = trafo,
           compile = T,
           modelname = "p_log")
```

```{r}
pars <- c(
  m1      	=	100	, # 1
  e1      	=	1000	, # 2
  p1      	=	100	, # 3
  k1      	=	0.1	, # 4
  k2      	=	0.1	, # 5
  k31     	=	0.01	, # 6
  k32     	=	1	, # 7
  k4      	=	1	, # 8
  k5      	=	1	, # 9
  k61     	=	0.01	, # 10
  k62     	=	1	, # 11
  a_xy    	=	.Machine$double.eps	, # 12
  a_tox_xy	=	.Machine$double.eps	, # 13
  a_toy_xy	=	.Machine$double.eps	  # 14 
) %>% 
  log

alpha_pars <- pars[str_detect(names(pars), "^a_")]
```

# Simulate perturbation data
For the perturbation experiments we use another parameter transformation to scale some parameters.
p_pert_fun automatically generates a parameter transformation function, which returns different sets of parameters for different conditions and a control condition.
Eg. if we have pars_perturbed = c("x1" = log(0.9)), p_pert_fun(pars_perturbed) will return a parameter transformation with the control condition with unchanged parameters and one condition in which x' = x + log(0.9)

"log(0.9)" corresponds to a ~10% reduction of the original parameter value.
```{r which pars are perturbed}
# Which pars shall be perturbed?
pars_perturbed <- c(m1 = log(0.9), p1 = log(0.9))

p_pert <- p_pert_fun(pars_perturbed = pars_perturbed,
                     pars = pars)
```

## Generate the perturbation data and run the algorithm

```{r}
perturbation_prediction <-  (xs*p_log*p_pert)(times = c(0,Inf), pars = pars, deriv = F)

r_0 <-   R_fun(pars_opt = alpha_pars[c("loga_tox_xy")],
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_0 %>% round(2)

r_kept <- r_kept_fun(pars_opt = alpha_pars, 
                     perturbation_prediction = perturbation_prediction,
                     obs_fun = g, 
                     p_fun = (p_log*p_pert),
                     pars = pars)


myfits <- mstrust(obj_alpha,
                  center = alpha_pars["a_tox_xy"] * 0 ,
                  studyname = "Fits",
                  cores = 3,
                  fits = 3,
                  sd = 3, 
                  r_kept = r_kept,
                  mypars = pars,
                  obs_fun = g,
                  p_fun = (p_log*p_pert),
                  perturbation_prediction = perturbation_prediction)

best_fit <- myfits %>% as.parframe %>% as.parvec
r_opt <- r_alpha_fun(best_fit, pars = pars, p_fun = (p_log*p_pert), obs_fun = g, perturbation_prediction = perturbation_prediction) %>% round(3)
r_opt %>% round(2)
```



# Run through all ERK perturbations, a = 1

```{r which pars are perturbed}
# Which pars shall be perturbed?

pars_perturbed_list <- list(
  e1 = c(m1 = log(0.9), e1 = log(0.9)),
  k31 = c(m1 = log(0.9), k31 = log(0.9)),
  k32 = c(m1 = log(0.9), k32 = log(0.9)),
  k4 = c(m1 = log(0.9), k4 = log(0.9)),
  p1 = c(m1 = log(0.9), p1 = log(0.9)),
  k5 = c(m1 = log(0.9), k5 = log(0.9)),
  k61 = c(m1 = log(0.9), k61 = log(0.9)),
  k62 = c(m1 = log(0.9), k62 = log(0.9))
  )

r_list <- lapply(pars_perturbed_list, function(pars_perturbed) {
p_pert <- p_pert_fun(pars_perturbed = pars_perturbed,
                     pars = pars)
## Generate the perturbation data and run the algorithm
perturbation_prediction <-  (xs*p_log*p_pert)(times = c(0,Inf), pars = pars, deriv = F)
r_1 <-   R_fun(pars_opt = alpha_pars[c("a_tox_xy")] * 0,
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)
})

r_list
```
# Pmatrix function
```{r}
pmatrix = function(x, digits=NULL, ...) {
  library(xtable)
  default_args = list(include.colnames=FALSE, only.contents=TRUE,
                      include.rownames=FALSE, hline.after=NULL, comment=FALSE,
                      print.results=FALSE)
  passed_args = list(...)
  calling_args = c(list(x=xtable(x, digits=digits)),
                   c(passed_args,
                     default_args[setdiff(names(default_args), names(passed_args))]))
  cat("\\begin{pmatrix}\n",
      do.call(print.xtable, calling_args),
      "\\end{pmatrix}\n")
}


pmatrix(r_list[["p1"]])

```

# Run through all ERK perturbations, a = aopt

```{r which pars are perturbed}
# Which pars shall be perturbed?
pars_perturbed_list <- list(
  e1 = c(m1 = log(0.9), e1 = log(0.9)),
  k31 = c(m1 = log(0.9), k31 = log(0.9)),
  k32 = c(m1 = log(0.9), k32 = log(0.9)),
  k4 = c(m1 = log(0.9), k4 = log(0.9)),
  p1 = c(m1 = log(0.9), p1 = log(0.9)),
  k5 = c(m1 = log(0.9), k5 = log(0.9)),
  k61 = c(m1 = log(0.9), k61 = log(0.9)),
  k62 = c(m1 = log(0.9), k62 = log(0.9))
  )

r_list <- lapply(pars_perturbed_list, function(pars_perturbed) {
p_pert <- p_pert_fun(pars_perturbed = pars_perturbed,
                     pars = pars)
## Generate the perturbation data and run the algorithm
perturbation_prediction <-  (xs*p_log*p_pert)(times = c(0,Inf), pars = pars, deriv = F)
r_1 <-   R_fun(pars_opt = alpha_pars[c("a_tox_xy")] * 0 + log(0.5),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)
})

r_list
```



# Compute the local response matrix in dependence of a
Use par_opt_setting to define, to which module(s) the complex should be added.
```{r, warning=FALSE}
par_opt_settings <- list("a_xy", "a_tox_xy") %>% `names<-`(1:2)
# debugonce(r_opt_dose_response)
dr_list <- lapply(par_opt_settings, function(i) {
  r_opt_dose_response(which_par = "k1",
                      dosages = pars["k1"],
                      pars_opt = alpha_pars[i],
                      alpha_scan_range = sort(unique(seq(-10,10,by = .1))),
                      pars = pars,
                      obs_fun = g,
                      p_fun = (p_log * p_pert),
                      cores = 4
  )
})

```


# Run through all MEK perturbations a=aopt

```{r which pars are perturbed}
# Which pars shall be perturbed?
pars_perturbed_list <- list(
  m1 = c(m1 = log(0.9), e1 = log(0.9)),
  k1 = c(k1 = log(0.9), e1 = log(0.9)),
  k2 = c(k2 = log(0.9), e1 = log(0.9))
  )

r_list <- lapply(pars_perturbed_list, function(pars_perturbed) {
p_pert <- p_pert_fun(pars_perturbed = pars_perturbed,
                     pars = pars)
## Generate the perturbation data and run the algorithm
perturbation_prediction <-  (xs*p_log*p_pert)(times = c(0,Inf), pars = pars, deriv = F)
r_1 <-   R_fun(pars_opt = alpha_pars[c("a_tox_xy")] * 0 + log(0.5),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)
})

r_list
```


# Outputs

```{r plotting helpers}
fancify_labels <- . %>% 
  {.$matrix[.$matrix == "r_alpha"] <- "r"
    .$matrix[.$matrix == "r_opt"] <- "r(a = a_opt)"
    .$matrix <- .$matrix %>% factor(levels = c("r", "r(a = a_opt)", "r_kept"))
    .} %>% 
  mutate(Element = r_element) %>% 
  {
    .$par_opt_setting[.$par_opt_setting == 1] <- "X* and Y*"
    .$par_opt_setting[.$par_opt_setting == 2] <- "X*"
    .$par_opt_setting <- .$par_opt_setting %>% factor(levels = c("X* and Y*","X*"))
    .}

assign_point_locations <- . %>% {.} %T>% {assign("point_locations", 
                                                 c(min(.$alpha), 
                                                   .$alpha[near(.$value,0,tol = 0.01)], 
                                                   max(.$alpha)), 
                                                 envir = .GlobalEnv)} %>% #assign values where points should be to .GlobalEnv
  mutate(points = replace(value, map_lgl(alpha, . %>% near(point_locations, tol = 1e-6) %>% any %>% `!`), NA))


remove_diags <- . %>% filter(! (r_element %in% paste0("r", c(11,22,33))))


print_matrices_at_point_locations <- function(point_locations, plot_data, path) {
  map(point_locations, function(i) {
    plot_data  %>%
      filter(near(alpha, i, tol = 1e-6)) %>%
      .$value %>% c(-1, ., -1) %>% matrix(ncol = 2) %>% round(2) %>%
      xtable() %>% print(tabular.enivronment = "pmatrix", floating = FALSE, hline.after=NULL, include.rownames=FALSE, include.colnames=FALSE) %>%
      str_replace_all(c("tabular"= "pmatrix", "\\{rr\\}" = "", "\\\\begin" = "$\\\\begin")) %>% paste0(., "$") 
  }) %>% 
    paste0(collapse = "\n\n") %>% 
    write_lines(paste0(path, ".tex"))
  system2("pandoc", c(paste0("-o ", path, ".docx"), paste0(path, ".tex")))
}
```

## Plot adding the complex to upper module
```{r}
plot_data <-
  dr_list %>%   
  combine_dr_list() %>% 
  filter(!is.na(value)) %>%
  remove_diags %>% 

  filter(matrix == "r_alpha") %>%

  mutate(alpha = exp(alpha)) %>%
  filter(alpha<5) %>% #.$alpha
  
  fancify_labels %>%
  filter(par_opt_setting %in% "X*") %>%

  assign_point_locations %>% 
  {.}

geom_rect_data <- plot_data %>% 
  filter(alpha %in% point_locations) %>% 
  group_by(alpha) %>% 
  summarise(maxval = max(value), minval = min(value))

text_data <- geom_rect_data %>% 
  mutate(label=rank(alpha)) %>% 
  group_by(alpha) %>% 
  mutate(value = c(minval-0.12, maxval+0.12)[which.min(abs(c(minval, maxval)))]) %>% 
  ungroup

plot_data <- plot_data %>% left_join(geom_rect_data)


myplot <- plot_data %>%   
  ggplot(mapping = aes(x = alpha, y = value, color = Element
                       # uncomment for one par_opt_setting only
                       # , alpha = par_opt_setting
                       )) +


  geom_rect(aes(xmin = alpha - 0.17, xmax = alpha + 0.17, ymin = minval - 0.07, ymax = maxval + 0.07), alpha = 1, fill = "#E9E9E9", color = "black", size = 0.3) + 
  geom_text(aes(label = label), data = text_data, color = "black", size = 5, fontface = "bold") +

  geom_line() +
  theme_dMod() +
  scale_colour_manual(values = c("blue", "red", "green")) +
  geom_abline(slope = 0,intercept = 0, linetype = 3) +
  geom_point(aes(y = points), size = 3)+
  scale_alpha_discrete(name = "Complex \nadded to", range = c(0.3,1)) +
  xlab("\tFree parameter a") +
  ylab("Connection coefficient")

myplot
ggsave("plot2upper.png", myplot)
print_matrices_at_point_locations(point_locations, plot_data, path = "matrices_of_plot2upper_Prabakaran")
```


## Plot adding the complex to both modules
```{r}
plot_data <-
  dr_list %>%   
  combine_dr_list() %>% 
  filter(!is.na(value)) %>%
  remove_diags %>% 

  filter(matrix == "r_alpha") %>%

  mutate(alpha = exp(alpha)) %>%
  filter(alpha<5) %>% #.$alpha
  
  fancify_labels %>%
  filter(par_opt_setting %in% "X* and Y*") %>%

  assign_point_locations %>% 
  {.}

geom_rect_data <- plot_data %>% 
  filter(alpha %in% point_locations) %>% 
  group_by(alpha) %>% 
  summarise(maxval = max(value), minval = min(value))

text_data <- geom_rect_data %>% 
  mutate(label=rank(alpha)) %>% 
  group_by(alpha) %>% 
  mutate(value = c(minval-0.12, maxval+0.12)[which.min(abs(c(minval, maxval)))]) %>% 
  ungroup

plot_data <- plot_data %>% left_join(geom_rect_data)

# print_matrices_at_point_locations(point_locations, plot_data)

myplot <- plot_data %>%   
  ggplot(mapping = aes(x = alpha, y = value, color = Element
                       # uncomment for one par_opt_setting only
                       # , alpha = par_opt_setting
                       )) +

  geom_rect(aes(xmin = alpha - 0.17, xmax = alpha + 0.17, ymin = minval - 0.07, ymax = maxval + 0.07), alpha = 1, fill = "#E9E9E9", color = "black", size = 0.3) + 
  geom_text(aes(label = label), data = text_data, color = "black", size = 5, fontface = "bold") +

  geom_line() +
  theme_dMod() +
  scale_colour_manual(values = c("blue", "red", "green")) +
  geom_abline(slope = 0,intercept = 0, linetype = 3) +
  geom_point(aes(y = points), size = 3)+
  scale_alpha_discrete(name = "Complex \nadded to", range = c(0.3,1)) +
  xlab("\tFree parameter a") +
  ylab("Connection coefficient")

myplot
ggsave("plot2both.png", myplot)
print_matrices_at_point_locations(point_locations, plot_data, path = "matrices_of_plot2both_Prabakaran")
```

## Tables
Print parameters
```{r}
pars %>% 
  .[!str_detect(names(.), "^a")] %>% 
  exp %>% 
  print2word_namedvec("pars_Prabakaran", names_in_mathenv = T)
```

Print ODEs
```{r}
reactions %>% 
  as.eqnvec() %>% 
  print2word_eqnvec("odes_Prabakaran")
```

## Clean up Workdir
```{r}
unlink(paste0("*.", c("c", "o", "so", "tex", "lnk")))
unlink("Fits/", T)
unlink(".pars_controlled/", T)
```















