---
title: "MEK-ERK-Model"
author: "Daniel Lill"
output: 
  pdf_document:
    keep_tex: true
---

# Load libraries and define some useful paths

```{r setup, results='hide', echo=FALSE, message=FALSE}
rm(list = ls())
library(MRAr)
library(scales)
library(xtable)
```

# Setup Model 
This is the setup of the ODE model.

Our (internal) states x are described by the odemodel which is turned into a prediction via the prediction function x():

dx/dt = f(x,p_dyn)
x = x(t, p_dyn, x_ini)

p_dyn are dynamical parameters like rate constants etc, x_ini are initial values for all states.
The prediction function xs() runs the stimulation until it reaches a steady state.


Then, these states are linked to observables y via the observation function g(). In our case the observables will be the communicating species:

y = g(x, p_obs)

p_obs are observation parameters like scaling factors.


Since all parameters are positive, it is beneficial to use log-parameters, therefore all parameters are transformed to the log-scale. These we call "outer parameters", p_outer. To feed them into our original model, we transform them back to the linear scale ("p_inner") by a parameter transformation function p(). This way, all our calculations can be made in log-space:

p_inner = p(p_outer) = p(log(p_obs), log(p_dyn), log(x_ini))


These three types of functions can be concatenated by the "*"-operator.
g(x(t, p(p_outer))) = g*x*p(t, p_outer)


## Prediction function
odemodel() writes the function f as a C-function, which is called by x and xs.
```{r Prediction function}
modelname   <- "Prabakaran"
mymodel     <- as.eqnlist(read.csv(paste0(modelname, ".csv")))
# mymodel

# myodemodel <- odemodel(mymodel, modelname = modelname, deriv = T)
# saveRDS(myodemodel, file = paste0("odemodel",modelname, ".rds"))
myodemodel <- readRDS(paste0("odemodel",modelname, ".rds"))

# Prediction function
x <- Xs(odemodel = myodemodel)
# Prediction function for steady states
xs <- Xs_steady(myodemodel)
# loadDLL(xs)
print(mymodel)
```


## Observation function
Observation function which returns our communicating species. 
The free parameters a_ij are introduced here to perform the linear combination of states which we wish to observe.
We chose to call them pars_opt, since they will be used to optimize the objective function later on.
```{r Naive observation function}
modules <- modules0 <- c("X","Y")

obs <- obs0 <- c("m2+a_xy*Cm2e1+a_tox_xy*Cm2e1",
                 "e2+a_xy*Cm2e1+a_toy_xy*Cm2e1")
g <- g0 <- Y(as.eqnvec(structure(obs0, names = modules0)), mymodel, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```




## Inner parameters
Load all parameters and format them for further use.
```{r Parameters}

pars_raw <- read.csv(paste0("pars.csv"), header = FALSE)  # dynamical parameters
pars_raw <- structure(pars_raw$V2, names = as.character(pars_raw$V1))
ic_raw <- read.csv(paste0("IC.csv"), header = F)          # initial conditions
ic_raw <- structure(ic_raw$V2, names = as.character(ic_raw$V1))
# sort ics like they appear in x, so it can be used in stode()
vars_x <- attr(x, "parameters")[attr(x, "parameters") %in% names(ic_raw)]
ic_raw <- ic_raw[vars_x]

pars_inner_opt_0 <- setdiff(attr(g, "parameters"),names(c(pars_raw,ic_raw)))
pars_inner_opt_0 <- structure(rep(0, length(pars_inner_opt_0)), 
                              names = pars_inner_opt_0)

pars_inner_0 <- pars_inner <- c(ic_raw, pars_raw, pars_inner_opt_0)


# # check if all parameters are assigned a value. Both lines should evaluate to the same number.
# x_pars <- unique(union((g) %>% attr("parameters"),(xs) %>% attr("parameters")))
# names(pars) %in% x_pars %>% sum()
# x_pars %in% names(pars) %>% sum()

```

## Parameter transformation function
```{r}
# Logtrafo of all pars
logtrafo <- structure(paste0("exp( log", names(pars_inner_0), ")"), 
                             names = names(pars_inner_0))

# initial values
logtrafo[c("m2", "Cm2e1", "e2", "Ce2p1")] <- 0



p_log <- P(trafo = logtrafo,
           compile = F,
           modelname = "p_log")
           

pars_opt_0 <- pars_opt <- structure(rep(log(0+.Machine$double.eps), length(pars_inner_opt_0)), names = paste0("log", names(pars_inner_opt_0)))

pars_0 <- pars <- structure(log(pars_inner_0+.Machine$double.eps), names = paste0("log", names(pars_inner_0))) 
pars_0[names(pars_opt_0)] <- log(0+.Machine$double.eps)


# pars_0
# p_log(pars_0, deriv = F)
```


# Simulate perturbation data

For the perturbation experiments we use another parameter transformation to scale some parameters.
p_pert_fun automatically generates a parameter transformation function, which returns different sets of parameters for different conditions and a control condition.
Eg. if we have pars_perturbed = c("logx1" = log(0.9)), p_pert_fun(pars_perturbed) will return a parameter transformation with the control condition with unchanged parameters and one condition in which logx' = logx + log(0.9)

"log(0.9)" corresponds to a ~10% reduction of the original parameter value.
```{r which pars are perturbed}
# Which pars shall be perturbed?
pars_perturbed_0 <- pars_perturbed <- c(logk1 = log(0.9), loge1 = log(0.9))

p_pert <- p_pert_fun(pars_perturbed = pars_perturbed_0,
                     pars = pars_0)
compile(p_pert, p_log, output = "p")
remove_c_and_o()
```



## Generate the perturbation data and run the algorithm
```{r}
loadDLL(xs, p_log, p_pert, g)
perturbation_prediction <- perturbation_prediction_0 <-  (xs*p_log*p_pert)(times = c(0,Inf), pars = pars, deriv = F)
# perturbation_prediction
r_kept_0 <- r_kept_fun(pars_opt = pars_opt_0, obs_fun = g, p_fun = (p_log*p_pert))
# r_kept_0 <- c(F,F,T,F)

# perturbation_prediction_0 <-  (xs*p_pert)(times = c(0,Inf), pars = pars_0, deriv = F)

r_0 <-   R_fun(pars_opt = pars_opt[c("loga_tox_xy")],
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_0%>% round(3)

myfits <- mstrust(obj_alpha,
                  center =  (pars_opt_0-pars_opt_0)["loga_tox_xy"],
                  studyname = "Fits",
                  cores = 3,
                  fits = 3,
                  sd = 3, mypars = pars)
# myfits %>% as.parframe() %>% plotValues()

(best_fit <- myfits %>% as.parframe() %>% as.parvec())
r_alpha_fun(best_fit, pars = pars) %>% round(3)

```

# Compute the local response matrix in dependence of a
Use par_opt_setting to define, to which module(s) the complex should be added.
```{r, warning=FALSE}
# par_opt_settings <- list("loga_xy", "loga_tox_xy", "loga_toy_xy", c("loga_tox_xy", "loga_toy_xy")) %>% set_names(1:4)
# par_opt_settings <- par_opt_settings[1:2]
# dr_list <- lapply(par_opt_settings, function(i) {
#   r_opt_dose_response(which_par = "logk1", 
#                       dosages = pars_0["logk1"],#+seq(-2,2,by = 2),
#                       pars_opt = pars_opt_0[i],
#                       alpha_scan_range = sort(unique(seq(-10,10,by = .1))),
#                       pars = pars
#   )
# })

# saveRDS(dr_list, file = "Prabakaran_dr_list.rds")

```


# Plot the local response matrix in dependence of a
 - Different opacities correspond to different communicating species

```{r plot}
dr_list <- readRDS("Prabakaran_dr_list.rds")

plot_data <- dr_list %>%   
  # extract2(1) %>%
  combine_dr_list() %>% 
  filter(par_opt_setting %in% 1:2) %>%
  filter(near(dose,log(0.1))) %>%
  filter(! (r_element %in% paste0("r", c(11,22,33)))) %>% 
  # filter(alpha >-5, alpha <3) %>% 
  filter(matrix == "r_alpha") %>%
  {
    .$matrix[.$matrix == "r_alpha"] <- "r"
    .$matrix[.$matrix == "r_opt"] <- "r(a = a_opt)"
    .$matrix <- .$matrix %>% factor(levels = c("r", "r(a = a_opt)", "r_kept"))
    .
  } %>% 
  mutate(Element = r_element) %>% 
  # mutate(points = value) %>% 
  # { logpoints <- c(-8.5,-0.7,2.1,9)
  #   nears <- seq(-10,10,by=0.1)[apply(sapply(logpoints, function(i) near(seq(-10,10,by=0.1),i)),1,any)]
  #   .[!(.$alpha %in% nears),"points"] <- NA
  #   .[(.$alpha < (-6))|(.$alpha > (7)),"value"] <- NA
  #   filter(., alpha >= (-9))
  # } %>% 
  {
    .$par_opt_setting[.$par_opt_setting == 1] <- "X* and Y*"
    .$par_opt_setting[.$par_opt_setting == 2] <- "X*"
    .$par_opt_setting <- .$par_opt_setting %>% factor(levels = c("X* and Y*","X*"))
    .
  } %>%
  
  
  # linear x axis
  mutate(alpha = exp(alpha)) %>%
  filter(alpha<5) %>%
  filter(!is.na(value)) %>%
  {.} %T>% {assign("nears", c(min(.$alpha), 0.4965853, 4.953032), envir = .GlobalEnv)} %>% #assign values where points should be to .GlobalEnv
  mutate(points = replace(value, map_lgl(alpha, . %>% near(nears, tol = 1e-6) %>% any %>% `!`), NA)) %>% #filter(!is.na(points)) %>% print
  # choose par_opt_settings to plot, uncomment alpha in ggplot call, if there is only one par_opt_setting
  filter(par_opt_setting == "X*")  

# Print the patrices at the points
map(nears, function(i) {
  plot_data  %>% 
  filter( near(alpha, i, 1e-6), par_opt == "loga_tox_xy") %>% 
  .$value %>% c(-1, ., -1) %>% matrix(ncol = 2) %>% round(2) %>% 
  xtable() %>% print(tabular.enivronment = "pmatrix", floating = FALSE, hline.after=NULL, include.rownames=FALSE, include.colnames=FALSE) %>% 
  str_replace_all(c("tabular"= "pmatrix", "\\{rr\\}" = "", "\\\\begin" = "$\\\\begin")) %>% paste0(., "$") %>% 
  cat 
})

myplot <- plot_data %>%   
  ggplot(mapping = aes(x = alpha, y = value, color = Element
                       # uncomment for one par_opt_setting only
                       # , alpha = par_opt_setting
                       )) +
  geom_line() +
  theme_dMod() +
  scale_color_dMod() +
  geom_abline(slope = 0,intercept = 0, linetype = 3) +
  # geom_rect(xmin = -0.7, xmax = 3.2, ymin = -10, ymax = 10, alpha = 0.01, size = 0)+
  geom_point(aes(y = points), size = 3)+
  # facet_grid(.~par_opt_setting, scales = "free")+
  # facet_grid(matrix~par_opt_setting, scales = "free")+
  scale_alpha_discrete(name = "Complex \nadded to", range = c(0.3,1)) +
  # scale_x_continuous(breaks = c(-5,0,5))+
  xlab("\tFree parameter a") +
  ylab("Connection coefficient")
# plotly::ggplotly(myplot)
myplot




ggshave(plot = myplot,
       filename = paste0("~/Promotion/Writing/Conferences/2018 07 SBMC/poster/pics/algorithm linear x-axis",".png"),
       width = 10,
       height = 8,
       units = "cm",
       device = "png",
       check_overwrite = F
       , commit_script = "Moved xlab a bit to the right"
       )



```

