---
title: "Hub model"
author: "Daniel Lill"
output: html
---

# Load libraries and define some useful paths
```{r setup}
library(parallel)
library(MRAr)
library(xtable)
```

# Model setup

## ODE model setup
```{r Prediction function Cascade, results='asis'}
reactions <- NULL %>% 
  addReaction(from = "x1", to = "x2", rate = "((v1 *  x1)/(k1 + x1))", description = "act_x") %>% 
  addReaction(from = "x2", to = "x1", rate = "((v2 * x2 )/(k2 + x2))", description = "deact_x") %>% 
  addReaction(from = "x2 + y1", to = "Cx2y1", rate = "(k31 * y1 *x2)  - (k32 *Cx2y1)", description = "x2_y_to_comp") %>% 
  addReaction(from = "Cx2y1", to = "x2 + y2", rate = "(k4 * Cx2y1)", description = "y2_formation") %>% 
  addReaction(from = "y2", to = "y1", rate = "((v5 * y2)/(k5 + y2))", description = "deact_y") %>% 
  addReaction(from = "x2 + z1", to = "Cx2z1", rate = "(k61 * z1 *x2) -  (k62 *Cx2z1)", description = "x2_z_to_comp") %>% 
  addReaction(from = "Cx2z1", to = "x2 + z2", rate = "(k7 * Cx2z1)", description = "z2_formation") %>% 
  addReaction(from = "z2", to = "z1", rate = "((v8 * z2)/(k8 + z2))", description = "deact_z") 

compiled_model <- odemodel(reactions, modelname = "Hub", estimate = c("x1", "k1"))

# Prediction function
x <- Xs(odemodel = compiled_model) 
# Prediction function for steady states
xs <- Xs_steady(compiled_model) # When computing steady states, set deriv = F
loadDLL(x)
```

## Observation function
Observation function which returns our communicating species. 
The free parameters a_ij are introduced here to perform the linear combination of states which we wish to observe.
We chose to call them pars_opt, since they will be used to optimize the objective function later on.
```{r Observation function with Cx2z2 in x-module}
modules <- c("X", "Y", "Z")

obs <-  c("x2 + a_1 * Cx2y1 + a_2 * Cx2z1 + a_tox_Cxy * Cx2y1 + a_tox_Cxz * Cx2z1",
          "y2 + a_1 * Cx2y1 + a_3 * Cx2z1",
          "z2 + a_2 * Cx2z1 + a_4 * Cx2y1 ")
g <-   Y(as.eqnvec(structure(obs, names = modules)), x, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```

### General log transformation and setting some pars to zero
It's better to do a logtransformation for all Fitting purposes and work with the log-Parameters.
The way to implement is this, is
log_pars %>% p_log %>% p_other %>% p_other_2 %>% ...
```{r}
trafo <- getParameters(x,g) %>% 
  setNames(.,.) %>% 
  insert("x~0", x = str_subset(.currentSymbols, "^C*[xyz](2|1.2)")) %>% # set inits to zero
  insert("x~exp(x)", x = .currentSymbols) %>%
  # {.} %T>% print %>% 
  {.} 

p_log <- P(trafo = trafo,
           modelname = "p_log",
           compile = T)
```

# These pars are in SI 22
```{r Parameters which imbalance the effects of y and z}
pars <- c(
x1       	=	100	, # 1
y1       	=	100	, # 4
z1       	=	100	, # 7

k1       	=	100	, # 9
v1       	=	1.5	, # 10
k2       	=	100	, # 11
v2       	=	1	, # 12
k31      	=	0.01	, # 13
k32      	=	0.4	, # 14
k4       	=	0.05	, # 15
k5       	=	100	, # 16
v5       	=	5	, # 17
k61      	=	0.05	, # 18
k62      	=	0.4	, # 19
k7       	=	0.05	, # 20
k8       	=	100	, # 21
v8       	=	5	, # 22

a_1      	=	.Machine$double.eps	, # 23
a_2      	=	.Machine$double.eps	, # 24
a_tox_Cxy	=	.Machine$double.eps	, # 25
a_tox_Cxz	=	.Machine$double.eps	, # 26
a_3      	=	.Machine$double.eps	, # 27
a_4      	=	.Machine$double.eps	  # 28
) %>% log 
alpha_pars <- pars[str_detect(names(pars), "^a_")]
```


# Apply different perturbations and run the algorithm
Since all possible combinations of paramter perturbations are many, these computations are done in a parallelized manner on a remote server.
If you want to quickly check the implementation with only one perturbation per module, uncomment the lower three lines assigning "perturbations_module_*", if you want to check the full set of perturbations, uncomment the upper three lines
```{r}
perturbations_module_x <- c("x1", "k1", "v1", "k2", "v2")
perturbations_module_y <- c("y1", "k5", "v5", "k31", "k32", "k4")
perturbations_module_z <- c("z1", "k8", "v8", "k61", "k62", "k7")

# perturbations_module_x <- c("x1")
# perturbations_module_y <- c("y1")
# perturbations_module_z <- c("z1")

perturbation_combinations <- expand.grid(perturbations_module_x, perturbations_module_y, perturbations_module_z, stringsAsFactors = F) 

perturbations <- map(1:nrow(perturbation_combinations), function(i) {
  mynames <- perturbation_combinations[i,] %>% unlist() %>% unname() %>% unique()
  if(length(mynames)<3) return(NULL)
  mynames %>% are_names_of(log(0.9))
}) %>% 
  .[!map_lgl(., is.null)]

myrunbg_job <- runbg({
  cores <- detectFreeCores()
  alphas <- c(-3, -2, -1, 0, 1, 2, 3)
    map(alphas, function(alpha) {
      
      mclapply(perturbations, function(perturbations_module_) {
        p_pert <- p_pert_fun(perturbations_module_, pars = pars, modelname = NULL)
        
        perturbation_prediction <- (xs*p_log*p_pert)(c(0,Inf), pars, deriv = F)
        
        try({r_0 <- R_fun(pars_opt = alpha_pars,
                          perturbation_prediction = perturbation_prediction,
                          obs_fun = g,
                          p_fun = (p_log*p_pert),
                          pars = pars) %>% local_response_matrix_eq10()})
        if(inherits(r_0, "try-error")) return(NULL)
        
        algo <- function(which_alpha_pars, prefix = NULL) {
          
          out <- tibble(r_kept = list(NULL), parframes = list(NULL), r_opt = list(NULL))
          
          r_kept <- r_kept_fun(pars_opt = alpha_pars[which_alpha_pars],
                               perturbation_prediction = perturbation_prediction,
                               obs_fun = g,
                               p_fun = (p_log * p_pert),
                               pars = pars,
                               alpha = -log(.Machine$double.eps) + alpha)
          out$r_kept <- list(r_kept)
          
          myfits <- mstrust(obj_alpha,
                            center =  structure(rep(0, length(which_alpha_pars)), names = which_alpha_pars),
                            studyname = "Fits",
                            cores = 1,
                            fits = 3,
                            sd = 1,
                            mypars = pars,
                            perturbation_prediction = perturbation_prediction,
                            r_kept = r_kept,
                            p_fun = (p_log * p_pert),
                            obs_fun = g)
          myparframe <- try(myfits %>% as.parframe())
          out$parframes <- list(myparframe)
          if (inherits(myparframe, "try-error")) {
            out$parframes <- list(myfits)
            return(out)
          }
          
          r_opt <- r_alpha_fun(pars_opt = myfits %>% as.parframe() %>% as.parvec() %>% unclass() ,
                               pars = pars,
                               perturbation_prediction = perturbation_prediction,
                               p_fun = (p_log*p_pert),
                               obs_fun = g)
          out$r_opt <- list(r_opt)
          return(out)
        }
        
        to_upstream_module <- algo(c("a_tox_Cxy", "a_toy_Cxz"), "upstream")
        to_both_modules <- algo(c("a_1", "a_2"), "both")
        
        bind_cols(tibble(r_0 = list(r_0 %>% round(2)), pars_perturbed = list(perturbations_module_), alpha = alpha), to_upstream_module, to_both_modules)
      }, mc.cores = cores) %>% 
        do.call(rbind,.) 
  }) %>% 
    do.call(rbind,.)
  
}, machine = c("knecht5"), input = ls(pos=.GlobalEnv), filename = "2018_09_24_hub"
, recover = T
)
```

Check if results are ready. Uncomment if needed
```{r}
# myrunbg_job$check()
```

Fetch/save/restore results. Comment out the first two lines after saving the fetched fit results.
```{r}
# myrunbg <- myrunbg_job$get()[[1]]
# saveRDS(myrunbg, file = "2018_09_24_hubresult.rds")
results <- readRDS("2018_09_24_hubresult.rds")

```

Purge files from the remote computing job.  Uncomment if needed
```{r}
# myrunbg_job$purge()
```


# Explore results

```{r unnest runbg job}
# helper functions
matrix_2_wide_df <- function(mat, nm) {mat %>% 
    `dim<-`(NULL) %>% 
    t %>% 
    as.data.frame %>% 
    `names<-`(nm)}

# prepare data for clustering
results_unnested_matrices <-
  results %>%
  rename(upstreamropt = r_opt, 
         bothropt = r_opt1,
         rzero = r_0) %>% 
  filter(map_lgl(upstreamropt, is.double), 
         map_lgl(bothropt, is.double),
         map_lgl(rzero, is.double)) %>% 
  mutate(upstreamropt = map(upstreamropt, matrix_2_wide_df, nm = (paste0("up_r", outer(1:3,1:3, paste0)))),
         bothropt = map(bothropt, . %>% matrix_2_wide_df(paste0("both_r", outer(1:3,1:3, paste0)))),
         rzero = map(rzero, . %>% matrix_2_wide_df(paste0("rzero_", outer(1:3,1:3, paste0)))),
         pars_perturbed = map(pars_perturbed, . %>% names %>% `names<-`(LETTERS[24:26]) %>% t %>% as.data.frame(stringsAsFactors = F)),
         perturbed_set = seq_along(alpha),
         alpha = as.numeric(alpha)) %>% 
  unnest(upstreamropt, bothropt, rzero, pars_perturbed)

#' clustering and print function
#'
#' @param results_unnested_matrices result from above
#' @param a alpha
#' @param col_start regex like "up_" or "both_"
#' @param nclust nclusters
#' @param which_clust print perturbations of which cluster?
cluster_matrices_and_print_perturbations <- function(results_unnested_matrices, a, col_start, nclust, which_clust) {
  kclusts <- results_unnested_matrices %>% 
    filter(near(alpha, a)) %>% 
    select(starts_with(col_start)) %>% 
    {.} %T>% {set.seed(1)} %>%
    kmeans(nclust) %>% 
    {.} %T>% print(digits = 2) %>%
    {.}
  
  kclusts$centers %>% 
    round(2) %>% 
    apply(1, . %>% matrix( nrow = 3) %>% 
    {.} %T>% assign(".mat", ., pos = .GlobalEnv) %>%
      print %>% {NULL})
  
  augmented <- broom::augment(kclusts, results_unnested_matrices %>% filter(near(alpha, a)))
  
  augmented %>% 
    filter(.cluster %in% which_clust) %>% 
    select(X,Y,Z) %>% as.matrix %>% print
  
  out <- augmented
}
```

```{r}
cluster_rzero <- cluster_matrices_and_print_perturbations(results_unnested_matrices, 0, "rzero_", 5, 4)
representative_r0_cl1 <- c("x1", "y1", "k8")
representative_r0_cl2 <- c("k1", "k5", "k8")
representative_r0_cl3 <- c("x1", "y1", "z1")
representative_r0_cl4 <- c("x1", "k5", "k8")
```

```{r}
cluster_up <- cluster_matrices_and_print_perturbations(results_unnested_matrices, 0, "up_", 2, 2)
representative_rup_cl1 <- c("x1", "y1", "z1")
representative_rup_cl2 <- c("x1", "y1", "k8")
```

```{r}
cluster_matrices_and_print_perturbations(results_unnested_matrices, 0, "both_", 5, 6)%>% 
  print(n = 100)
```


```{r}
#' Get the relevant matrices for a perturbation and alpha-setting
#'
#' @param results as is from runbg
#' @param a alpha value
#' @param pp names of perturbed parameters
#' @param pars,xs,obs_fun,p_log objects that were used to generate the result
#'
#' @return list of matrices
get_r0_r1_ropt_rkept <- function(results, a, pp, pars, xs, obs_fun, p_log) {
  myresults <- filter(results, alpha == a, map_lgl(pars_perturbed, . %>% names %>% identical(pp)))
  
  p_pert <- p_pert_fun(myresults$pars_perturbed[[1]], pars, NULL)
  perturbation_prediction <- (xs*p_log*p_pert)(c(0,Inf), pars, deriv = F)
  r_1 <-  r_alpha_fun(pars_opt = myresults$parframes[[1]] %>% as.parvec %>% unclass %>%  `*`(0),
                               pars = pars,
                               perturbation_prediction = perturbation_prediction,
                               p_fun = (p_log*p_pert),
                               obs_fun = g) %>% round(2)
  
  list(perturbed_parameters = t(pp), r_0 = myresults$r_0[[1]], r_1 = r_1, r_opt = myresults$r_opt[[1]] %>% round(2), r_kept = myresults$r_kept[[1]])
}

get_r0_r1_ropt_rkept(results, 0, representative_r0_cl1, pars, xs, obs_fun, p_log) %>% 
  imap(~.x %>% {cbind(c(.y, rep("", nrow(.x)-1)), .)}) %>% 
  print2word_matrixlist("matrices_cluster1")
get_r0_r1_ropt_rkept(results, 0, representative_r0_cl2, pars, xs, obs_fun, p_log)%>%   
  imap(~.x %>% {cbind(c(.y, rep("", nrow(.x)-1)), .)}) %>% 
  print2word_matrixlist("matrices_cluster2")
get_r0_r1_ropt_rkept(results, 0, representative_r0_cl3, pars, xs, obs_fun, p_log)%>%   
  imap(~.x %>% {cbind(c(.y, rep("", nrow(.x)-1)), .)}) %>% 
  print2word_matrixlist("matrices_cluster3")
get_r0_r1_ropt_rkept(results, 0, representative_r0_cl4, pars, xs, obs_fun, p_log)%>%   
  imap(~.x %>% {cbind(c(.y, rep("", nrow(.x)-1)), .)}) %>% 
  print2word_matrixlist("matrices_cluster4")
get_r0_r1_ropt_rkept(results, 0, representative_rup_cl1, pars, xs, obs_fun, p_log)%>%   
  imap(~.x %>% {cbind(c(.y, rep("", nrow(.x)-1)), .)}) %>% 
  print2word_matrixlist("matrices_cluster_opt1")
get_r0_r1_ropt_rkept(results, 0, representative_rup_cl2, pars, xs, obs_fun, p_log)%>%   
  imap(~.x %>% {cbind(c(.y, rep("", nrow(.x)-1)), .)}) %>% 
  print2word_matrixlist("matrices_cluster_opt2")
```





# Description of results
```{r}
# helper functions/objects for exporting to ms word
communicating_species <- paste0(LETTERS[24:26],"*")
```

Export some tables for the publication
```{r}
# results %>% 
#   filter(alpha == 0) %>% 
#   filter(map_lgl(pars_perturbed, . %>% {str_detect(names(.), c("x1", "y1", "z1"))} %>% all)) %>% 
#   .[c("r_opt","r_opt1")] %>%
#   unlist(F) %>% #.[t(matrix(1:6, nrow = 3))] %>% 
#   {mymatrix <- .; `names<-`(mymatrix, names(mymatrix) %>% str_replace_all(c("1$" = "gl = -10", "2$" = "gl = 0", "3$" = "gl = 2",
#                                                "r_0" = "a = 0, ", "r_1" = "a = 1, ", "r_opt" = "a = a_opt, ")))} %>%
#   imap(~ .x  %>% round(2) %>% as.tibble() %>% `names<-`(communicating_species) %>% bind_cols(aopt = rep(.y ,3), module = communicating_species,.) ) %>% 
#   print_msword_table("matrices")
