---
title: "Hub model"
author: "Daniel Lill"
output:
  pdf_document:
    keep_tex: yes
---

# Load libraries and define some useful paths
```{r setup, results='hide', echo=FALSE, message=FALSE}
library(parallel)
library(MRAr)
library(xtable)
```

# Model setup

## ODE model setup
```{r Prediction function Cascade, results='asis'}
reactions <- NULL %>% 
  addReaction(from = "x1", to = "x2", rate = "((v1 *  x1)/(k1 + x1))", description = "act_x") %>% 
  addReaction(from = "x2", to = "x1", rate = "((v2 * x2 )/(k2 + x2))", description = "deact_x") %>% 
  addReaction(from = "x2 + y1", to = "Cx2y1", rate = "(k31 * y1 *x2)  - (k32 *Cx2y1)", description = "x2_y_to_comp") %>% 
  addReaction(from = "Cx2y1", to = "x2 + y2", rate = "(k4 * Cx2y1)", description = "y2_formation") %>% 
  addReaction(from = "y2", to = "y1", rate = "((v5 * y2)/(k5 + y2))", description = "deact_y") %>% 
  addReaction(from = "x2 + z1", to = "Cx2z1", rate = "(k61 * z1 *x2) -  (k62 *Cx2z1)", description = "x2_z_to_comp") %>% 
  addReaction(from = "Cx2z1", to = "x2 + z2", rate = "(k7 * Cx2z1)", description = "z2_formation") %>% 
  addReaction(from = "z2", to = "z1", rate = "((v8 * z2)/(k8 + z2))", description = "deact_z") 

compiled_model <- odemodel(reactions, modelname = "Hub", estimate = c("x1", "k1"))

# Prediction function
x <- Xs(odemodel = compiled_model) 
# Prediction function for steady states
xs <- Xs_steady(compiled_model) # When computing steady states, set deriv = F
loadDLL(x)
```

## Observation function
Observation function which returns our communicating species. 
The free parameters a_ij are introduced here to perform the linear combination of states which we wish to observe.
We chose to call them pars_opt, since they will be used to optimize the objective function later on.
```{r Observation function with Cx2z2 in x-module}
modules <- c("X", "Y", "Z")

obs <-  c("x2 + a_1 * Cx2y1 + a_2 * Cx2z1 + a_tox_Cxy * Cx2y1 + a_tox_Cxz * Cx2z1",
          "y2 + a_1 * Cx2y1 + a_3 * Cx2z1",
          "z2 + a_2 * Cx2z1 + a_4 * Cx2y1 ")
g <-   Y(as.eqnvec(structure(obs, names = modules)), reactions, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```

### General log transformation and setting some pars to zero
It's better to do a logtransformation for all Fitting purposes and work with the log-Parameters.
The way to implement is this, is
log_pars %>% p_log %>% p_other %>% p_other_2 %>% ...
```{r}
trafo <- getParameters(x,g) %>% 
  setNames(.,.) %>% 
  insert("x~0", x = str_subset(.currentSymbols, "^C*[xyz](2|1.2)")) %>% # set inits to zero
  insert("x~exp(x)", x = .currentSymbols) %>%
  # {.} %T>% print %>% 
  {.} 

p_log <- P(trafo = trafo,
           modelname = "p_log",
           compile = T)
```

# These pars are in SI 22
```{r Parameters which imbalance the effects of y and z}
pars <- c(
x1       	=	100	, # 1
y1       	=	100	, # 4
z1       	=	100	, # 7

k1       	=	100	, # 9
v1       	=	1.5	, # 10
k2       	=	100	, # 11
v2       	=	1	, # 12
k31      	=	0.01	, # 13
k32      	=	0.4	, # 14
k4       	=	0.05	, # 15
k5       	=	100	, # 16
v5       	=	5	, # 17
k61      	=	0.05	, # 18
k62      	=	0.4	, # 19
k7       	=	0.05	, # 20
k8       	=	100	, # 21
v8       	=	5	, # 22

a_1      	=	.Machine$double.eps	, # 23
a_2      	=	.Machine$double.eps	, # 24
a_tox_Cxy	=	.Machine$double.eps	, # 25
a_tox_Cxz	=	.Machine$double.eps	, # 26
a_3      	=	.Machine$double.eps	, # 27
a_4      	=	.Machine$double.eps	  # 28
) %>% log 
alpha_pars <- pars[str_detect(names(pars), "^a_")]
```


# Apply different perturbations and run the algorithm
Since all possible combinations of paramter perturbations are many, these computations are done in a parallelized manner on a remote server.
If you want to quickly check the implementation with only one perturbation per module, uncomment the lower three lines assigning "perturbations_module_*", if you want to check the full set of perturbations, uncomment the upper three lines
```{r}
perturbations_module_x <- c("x1", "k1", "v1", "k2", "v2")
perturbations_module_y <- c("y1", "k5", "v5", "k31", "k32", "k4")
perturbations_module_z <- c("z1", "k8", "v8", "k61", "k62", "k7")

# perturbations_module_x <- c("x1")
# perturbations_module_y <- c("y1")
# perturbations_module_z <- c("z1")

perturbation_combinations <- expand.grid(perturbations_module_x, perturbations_module_y, perturbations_module_z, stringsAsFactors = F) 

perturbations <- map(1:nrow(perturbation_combinations), function(i) {
  mynames <- perturbation_combinations[i,] %>% unlist() %>% unname() %>% unique()
  if(length(mynames)<3) return(NULL)
  mynames %>% are_names_of(log(0.9))
}) %>% 
  .[!map_lgl(., is.null)]

myrunbg_job <- runbg({
  cores <- detectFreeCores()
  alphas <- c(-3, -2, -1, 0, 1, 2, 3)
    map(alphas, function(alpha) {
      
      mclapply(perturbations, function(perturbations_module_) {
        p_pert <- p_pert_fun(perturbations_module_, pars = pars, modelname = NULL)
        
        perturbation_prediction <- (xs*p_log*p_pert)(c(0,Inf), pars, deriv = F)
        
        try({r_0 <- R_fun(pars_opt = alpha_pars,
                          perturbation_prediction = perturbation_prediction,
                          obs_fun = g,
                          p_fun = (p_log*p_pert),
                          pars = pars) %>% local_response_matrix_eq10()})
        if(inherits(r_0, "try-error")) return(NULL)
        
        algo <- function(which_alpha_pars, prefix = NULL) {
          
          out <- tibble(r_kept = list(NULL), parframes = list(NULL), r_opt = list(NULL))
          
          r_kept <- r_kept_fun(pars_opt = alpha_pars[which_alpha_pars],
                               perturbation_prediction = perturbation_prediction,
                               obs_fun = g,
                               p_fun = (p_log * p_pert),
                               pars = pars,
                               alpha = -log(.Machine$double.eps) + alpha)
          out$r_kept <- list(r_kept)
          
          myfits <- mstrust(obj_alpha,
                            center =  structure(rep(0, length(which_alpha_pars)), names = which_alpha_pars),
                            studyname = "Fits",
                            cores = 1,
                            fits = 3,
                            sd = 1,
                            mypars = pars,
                            perturbation_prediction = perturbation_prediction,
                            r_kept = r_kept,
                            p_fun = (p_log * p_pert),
                            obs_fun = g)
          myparframe <- try(myfits %>% as.parframe())
          out$parframes <- list(myparframe)
          if (inherits(myparframe, "try-error")) {
            out$parframes <- list(myfits)
            return(out)
          }
          
          r_opt <- r_alpha_fun(pars_opt = myfits %>% as.parframe() %>% as.parvec() %>% unclass() ,
                               pars = pars,
                               perturbation_prediction = perturbation_prediction,
                               p_fun = (p_log*p_pert),
                               obs_fun = g)
          out$r_opt <- list(r_opt)
          return(out)
        }
        
        to_upstream_module <- algo(c("a_tox_Cxy", "a_toy_Cxz"), "upstream")
        to_both_modules <- algo(c("a_1", "a_2"), "both")
        
        bind_cols(tibble(r_0 = list(r_0 %>% round(2)), pars_perturbed = list(perturbations_module_), alpha = alpha), to_upstream_module, to_both_modules)
      }, mc.cores = cores) %>% 
        do.call(rbind,.) 
  }) %>% 
    do.call(rbind,.)
  
}, machine = c("knecht5"), input = ls(pos=.GlobalEnv), filename = "2018_09_24_hub"
# , recover = T
)
```

Check if results are ready.
```{r}
myrunbg_job$check()
```

Fetch/save/restore results
```{r}
myrunbg <- myrunbg_job$get()[[1]]
saveRDS(myrunbg, file = "2018_09_24_hubresult.rds")
results <- readRDS("2018_09_24_hubresult.rds")

```

Purge files from the remote computing job.
```{r}
myrunbg_job$purge()
```





# Description of results
```{r}
# helper functions/objects for exporting to ms word
communicating_species <- paste0(LETTERS[24:26],"*")
print_msword_table <- function(x, path, digits = 2) {
  tables <- map(x, . %>% xtable::xtable(digits = digits) %>% {xtable:::print.xtable(invisible(.), include.rownames = F)}) %>% 
    paste0(collapse = "\n\n")
  
  tables %>%
    paste("\\documentclass[10pt,a4paper]{article}
\\usepackage[latin1]{inputenc}
\\usepackage{amsmath}
\\usepackage{amsfonts}
\\usepackage{amssymb}
\\usepackage{graphicx}
\\begin{document}", ., "\\end{document}") %>%
    write_lines(paste0(path,".tex"))
  system2("pandoc", c(paste0("-o ",path,".docx"), paste0(path,".tex")))
  }

```

Export some tables for the publication
```{r}
results %>% 
  filter(alpha == 0) %>% 
  filter(map_lgl(pars_perturbed, . %>% {str_detect(names(.), c("x1", "y1", "z1"))} %>% all)) %>% 
  .[c("r_opt","r_opt1")] %>%
  unlist(F) %>% #.[t(matrix(1:6, nrow = 3))] %>% 
  {mymatrix <- .; `names<-`(mymatrix, names(mymatrix) %>% str_replace_all(c("1$" = "gl = -10", "2$" = "gl = 0", "3$" = "gl = 2",
                                               "r_0" = "a = 0, ", "r_1" = "a = 1, ", "r_opt" = "a = a_opt, ")))} %>%
  imap(~ .x  %>% round(2) %>% as.tibble() %>% `names<-`(communicating_species) %>% bind_cols(aopt = rep(.y ,3), module = communicating_species,.) ) %>% print_msword_table("matrices")
