---
title: "Hub model"
author: "Daniel Lill"
output:
  pdf_document:
    keep_tex: yes
---

Load libraries and define some useful paths

```{r Libraries, results='hide', echo=FALSE, message=FALSE}
rm(list = ls())
library(plotly)
library(tidyverse)
library(MRAr)
library(scales)
library(xtable)
```

# Setup Model 
This is the setup of the ODE model.

Our (internal) states x are described by the odemodel which is turned into a prediction via the prediction function x():

dx/dt = f(x,p_dyn)
x = x(t, p_dyn, x_ini)

p_dyn are dynamical parameters like rate constants etc, x_ini are initial values for all states.
The prediction function xs() runs the stimulation until it reaches a steady state.


Then, these states are linked to observables y via the observation function g(). In our case the observables will be the communicating species:

y = g(x, p_obs)

p_obs are observation parameters like scaling factors.


Since all parameters are positive, it is beneficial to use log-parameters, therefore all parameters are transformed to the log-scale. These we call "outer parameters", p_outer. To feed them into our original model, we transform them back to the linear scale ("p_inner") by a parameter transformation function p(). This way, all our calculations can be made in log-space:

p_inner = p(p_outer) = p(log(p_obs), log(p_dyn), log(x_ini))


These three types of functions can be concatenated by the "*"-operator.
g(x(t, p(p_outer))) = g*x*p(t, p_outer)


## Prediction function
odemodel() writes the function f as a C-function, which is called by x and xs.
```{r Prediction function Cascade, results='asis'}
modelname   <- "Hub"
mymodel     <- as.eqnlist(read.csv(paste0(modelname, ".csv")))

myodemodel <- odemodel(mymodel, modelname = paste0(modelname), deriv = T)
save(myodemodel, file = paste0("odemodel",modelname, ".RData"))
load(paste0("odemodel",modelname, ".RData"))

# Prediction function
x <- Xs(odemodel = myodemodel)
# Prediction function for steady states
xs <- Xs_steady(myodemodel) # When computing steady states, set deriv = F
print(mymodel)
# mymodel %>% as.eqnvec() %>% print(pander = T)
# myodemodel[[1]] %>% attr("eq")
```

## Observation function
Observation function which returns our communicating species. 
The free parameters a_ij are introduced here to perform the linear combination of states which we wish to observe.
We chose to call them pars_opt, since they will be used to optimize the objective function later on.
```{r Observation function with Cx2z2 in x-module}
modules <- modules0 <- c("x2","y2", "z2")

obs <- obs0 <- c("x2 + a_1 * Cx2y1 + a_2 * Cx2z1 + a_tox_Cxy * Cx2y1 + a_tox_Cxz * Cx2z1",
          "y2 + a_1 * Cx2y1 + a_3 * Cx2z1",
          "z2 + a_2 * Cx2z1 + a_4 * Cx2y1 ")
g <-  g0 <- Y(as.eqnvec(structure(obs0, names = modules0)), mymodel, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```

## Inner parameters
Load all parameters and format them for further use.
```{r Parameters Cascade}

pars_raw <- read.csv(paste0("pars.csv"), header = FALSE)  # dynamical parameters
pars_raw <- structure(pars_raw$V2, names = as.character(pars_raw$V1))
ic_raw <- read.csv(paste0("IC.csv"), header = F)          # initial conditions
ic_raw <- structure(ic_raw$V2, names = as.character(ic_raw$V1))
# sort ics like they appear in x, so it can be used in stode()
vars_x <- attr(x, "parameters")[attr(x, "parameters") %in% names(ic_raw)]
ic_raw <- ic_raw[vars_x]

pars_inner_opt_0 <- setdiff(attr(g, "parameters"),names(c(pars_raw,ic_raw)))
pars_inner_opt_0 <- structure(rep(0, length(pars_inner_opt_0)), 
                              names = pars_inner_opt_0)

pars_inner_0 <- pars_inner <- c(ic_raw, pars_raw, pars_inner_opt_0)


# check if all parameters are assigned a value. Both lines should evaluate to the same number.
# x_pars <- unique(union((g) %>% attr("parameters"),(xs) %>% attr("parameters")))
# names(pars) %in% x_pars %>% sum()
# x_pars %in% names(pars) %>% sum()
```


## Parameter transformation function
```{r}
# Logtrafo of all pars
logtrafo <- structure(paste0("exp( log", names(pars_inner_0), ")"), 
                             names = names(pars_inner_0))


# initial values
logtrafo[c("x2", "Cx2y1", "y2", "Cx2z1", "z2")] <- 0


p_log <- P(trafo = logtrafo,
           compile = TRUE,
           modelname = "p_log")
           

pars_0 <- pars <- structure(log(pars_inner_0+.Machine$double.eps), names = paste0("log", names(pars_inner_0))) 

pars_opt_0 <- pars_opt <- structure(log(pars_inner_opt_0+.Machine$double.eps), names = paste0("log", names(pars_inner_opt_0)))

# pars_opt_0 <- pars_opt <- pars_opt_0[c("loga_1", "loga_2")]
# pars_0
# p_log(pars_0, deriv = F)
```

# Simulate perturbation data

For the perturbation experiments we use another parameter transformation to scale some parameters.
p_pert_fun automatically generates a parameter transformation function, which returns different sets of parameters for different conditions and a control condition.
Eg. if we have pars_perturbed = c("logx1" = log(0.9)), p_pert_fun(pars_perturbed) will return a parameter transformation with the control condition with unchanged parameters and one condition in which logx' = logx + log(0.9)

"log(0.9)" corresponds to a ~10% reduction of the original parameter value.
```{r}
# Which pars shall be perturbed?
# pars_perturbed_0 <- pars_perturbed <- c("logx1" = log(0.9), "logy1" = log(0.9), "logz1" = log(0.9))
pars_perturbed_0 <- pars_perturbed <- c("logx1" = log(0.8), "logy1" = log(0.8), "logz1" = log(0.8))



# pars_perturbed
p_pert <- p_pert_fun(pars_perturbed = pars_perturbed, pars = pars, modelname = NULL)

```

## Final parameter values for different Figures

Parameters for Figure 6
```{r}
# pars <- c(100,0,0,100,0,0,100,0, 
#           100,1.5,100,1,     0.01,0.4,0.05,
#           100,5,            0.05,0.4,0.008,
#           100,1,
#           rep(0, length(pars_opt_0))) %>%
#   set_names(names(pars_inner_0)) # as in cascade
pars <- c(
x1       	=	100	, # 1
x2       	=	0	, # 2
Cx2y1    	=	0	, # 3
y1       	=	100	, # 4
y2       	=	0	, # 5
Cx2z1    	=	0	, # 6
z1       	=	100	, # 7
z2       	=	0	, # 8
k1       	=	100	, # 9
v1       	=	1.5	, # 10
k2       	=	100	, # 11
v2       	=	1	, # 12
k31      	=	0.01	, # 13
k32      	=	0.4	, # 14
k4       	=	0.05	, # 15
k5       	=	100	, # 16
v5       	=	5	, # 17
k61      	=	0.05	, # 18
k62      	=	0.4	, # 19
k7       	=	0.008	, # 20
k8       	=	100	, # 21
v8       	=	1	, # 22
a_1      	=	0	, # 23
a_2      	=	0	, # 24
a_tox_Cxy	=	0	, # 25
a_tox_Cxz	=	0	, # 26
a_3      	=	0	, # 27
a_4      	=	0	  # 28 
)

pars_0 <- pars <- log(pars+.Machine$double.eps) %>% set_names(names(pars_0))
```

# These pars are in SI 22
```{r Parameters which imbalance the effects of y and z}
# # pars <- c(100,0,0,100,0,0,100,0,
# #           100,1.5,100,1,     0.01,0.4,0.05,
# #           100,5,            0.05,0.4,0.008,
# #           100,1,
# #           rep(0, length(pars_opt_0))) %>%
# #   set_names(names(pars_inner_0)) # as in cascade
pars <- c(
x1       	=	100	, # 1
x2       	=	0	, # 2
Cx2y1    	=	0	, # 3
y1       	=	100	, # 4
y2       	=	0	, # 5
Cx2z1    	=	0	, # 6
z1       	=	100	, # 7
z2       	=	0	, # 8
k1       	=	100	, # 9
v1       	=	1.5	, # 10
k2       	=	100	, # 11
v2       	=	1	, # 12
k31      	=	0.01	, # 13
k32      	=	0.4	, # 14
k4       	=	0.05	, # 15
k5       	=	100	, # 16
v5       	=	5	, # 17
k61      	=	0.05	, # 18
k62      	=	0.4	, # 19
k7       	=	0.05	, # 20
k8       	=	100	, # 21
v8       	=	5	, # 22
a_1      	=	0	, # 23
a_2      	=	0	, # 24
a_tox_Cxy	=	0	, # 25
a_tox_Cxz	=	0	, # 26
a_3      	=	0	, # 27
a_4      	=	0	  # 28
)
# 
pars_0 <- pars <- log(pars+.Machine$double.eps) %>% set_names(names(pars_0))
```



```{r try to run the gadget}
# myframe <- dMod.frame("1", g,x,p_log, NULL, NULL, xs = list(xs), p_pert = list(p_pert), pars = list(pars))
# which_pars_opt <- c("loga_tox_Cxy", "loga_tox_Cxz", "loga_1","loga_2", "loga_3", "loga_4")
# ropt_gadget(which_pars_opt = which_pars_opt, myframe)
```



## Generate the perturbation data
```{r}
perturbation_prediction_0 <- perturbation_prediction <-  (xs*p_log*p_pert)(times = c(0,Inf), pars = pars, deriv = F)

# (x*p_log*p_pert)(times = seq(0,2000, length.out = 50), pars = pars) %>%
#   plotPrediction()

# perturbation_prediction %>% lapply(round,1)
```

# Apply the algorithm as an example
This is an example of how the algorithm is applied.
```{r}
which_pars_opt <- c("loga_tox_Cxy", "loga_tox_Cxz", "loga_1","loga_2", "loga_3", "loga_4")[1:2]
# which_pars_opt <- names(pars_opt_0)

r_kept_0 <- r_kept_fun(pars_opt = pars_opt[which_pars_opt],
                       perturbation_prediction = perturbation_prediction, 
                       obs_fun = g,
                       p_fun = (p_log * p_pert), 
                       pars = pars,
                       alpha = -log(0+.Machine$double.eps))
# r_kept_0[7]<- F
r_kept_0

# r_kept_0 <- c(F,F,T,T,F,F,F,T,F) %>% matrix(ncol = 3)

r_0 <-   R_fun(pars_opt = pars_opt[which_pars_opt],
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_0%>% round(2)

r_1 <-   R_fun(pars_opt = structure(rep(0, length(which_pars_opt)), names = which_pars_opt),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)
r_kept_0 <- c(F,F,F, T,F,T, T,T,F)
# r_kept_0 <- c(F,F,F, T,F,F, T,F,F)
# matrix(c(F,F,F,T,F,F,F,T,F), ncol = 3)
myfits <- mstrust(obj_alpha, center =  structure(rep(0, length(which_pars_opt)), names = which_pars_opt), studyname = "Fits", cores = 3, fits = 3, sd = 1, mypars = pars, perturbation_prediction = perturbation_prediction_0, r_kept = r_kept_0)
# myfits %>% as.parframe()
best_fit <- myfits %>% as.parframe() %>% as.parvec()
exp(best_fit)
r_alpha_fun(pars_opt = unclass(best_fit), pars = pars, perturbation_prediction = perturbation_prediction) %>% round(2)
# pars
```




```{r}
# Which pars shall be perturbed?
# pars_perturbed_0 <- pars_perturbed <- c("logx1" = log(0.9), "logy1" = log(0.9), "logz1" = log(0.9))
pars_perturbed_0 <- pars_perturbed <- c("logv1" = log(0.8), "logk31" = log(0.8), "logz1" = log(0.8))
# pars_perturbed
p_pert <- p_pert_fun(pars_perturbed = pars_perturbed, pars = pars, modelname = NULL)
perturbation_prediction_0 <- perturbation_prediction <-  (xs*p_log*p_pert)(times = c(0,Inf), pars = pars, deriv = F)


which_pars_opt <- c("loga_tox_Cxy", "loga_tox_Cxz", "loga_1","loga_2", "loga_3", "loga_4")[1:2]
# which_pars_opt <- names(pars_opt_0)

r_kept_0 <- r_kept_fun(pars_opt = pars_opt[which_pars_opt],
                       perturbation_prediction = perturbation_prediction, 
                       obs_fun = g,
                       p_fun = (p_log * p_pert), 
                       pars = pars,
                       alpha = -log(0+.Machine$double.eps))
# r_kept_0[7]<- F
r_kept_0

# r_kept_0 <- c(F,F,T,T,F,F,F,T,F) %>% matrix(ncol = 3)

r_0 <-   R_fun(pars_opt = pars_opt[which_pars_opt],
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_0%>% round(2)

r_1 <-   R_fun(pars_opt = structure(rep(0, length(which_pars_opt)), names = which_pars_opt),
               perturbation_prediction = perturbation_prediction,
               obs_fun = g,
               p_fun = (p_log*p_pert),
               pars = pars) %>% local_response_matrix_eq10()
r_1 %>% round(2)
r_kept_0 <- c(F,F,F, T,F,T, T,T,F)
# r_kept_0 <- c(F,F,F, T,F,F, T,F,F)
# matrix(c(F,F,F,T,F,F,F,T,F), ncol = 3)
myfits <- mstrust(obj_alpha, center =  structure(rep(0, length(which_pars_opt)), names = which_pars_opt), studyname = "Fits", cores = 3, fits = 3, sd = 1, mypars = pars, perturbation_prediction = perturbation_prediction_0, r_kept = r_kept_0)
# myfits %>% as.parframe()
best_fit <- myfits %>% as.parframe() %>% as.parvec()
exp(best_fit)
r_opt <- r_alpha_fun(pars_opt = unclass(best_fit), pars = pars, perturbation_prediction = perturbation_prediction) %>% round(2)
# pars
r_opt
```

```{r}
mypublish4 <- function(x, path, digits = 2) {
  
  tables <- 
    map(x, . %>%
    # replaceSymbols(what = mynames, by = paste0("[",mynames, "]"), x = .) %>%
    # {`names<-`(paste0(" ", ., " "), paste0("d/dt[ ", names(.), "]"))} %>%
    # tibble(names(.),.) %>%
    xtable(digits = digits) %>% {xtable:::print.xtable(invisible(.), include.rownames = F)}) %>% 
    paste0(collapse = "\n\n")
  
  
  tables %>%
    paste("\\documentclass[10pt,a4paper]{article}
\\usepackage[latin1]{inputenc}
\\usepackage{amsmath}
\\usepackage{amsfonts}
\\usepackage{amssymb}
\\usepackage{graphicx}
\\begin{document}", ., "\\end{document}") %>%
    write_lines(paste0(path,".tex"))
  system2("pandoc", c(paste0("-o ",path,".docx"), paste0(path,".tex")))
  }

list(r_0 = r_0, r_1 = r_1, r_opt = r_opt) %>% 
  imap(~ .x  %>% round(2) %>% as.tibble() %>% `names<-`(mynames) %>% bind_cols(aopt = rep(.y ,3), module = mynames,.) ) %>% mypublish4("matrices_gainuf")
```


# Plots for main text
Run the algorithm for different values of the parameters which control the feedback.
Do this for the two different options of defining communicating species (adding them to both modules or only the upstream module).
Also, scan the pars_opt in a finer grid than in the section above.
```{r, eval=FALSE}
par_opt_names <- pars_opt_0 %>% names %>% sort

par_opt_settings <- list(par_opt_names[c(1,3,2)]
                         , par_opt_names[c(1,2)]
                         )
dr_list <- lapply(seq_along(par_opt_settings), function(i) {
  # list(
  r_opt_dose_response(which_par = "logk31",
                      dosages = sort(unique(c(seq(-6,-3,by = 0.5)))),
                      pars = pars,
                      alpha_scan_range = seq(-5,5,by = 0.1),
                      pars_opt = pars_opt_0[par_opt_settings[[i]]])
  # )
})

```


# General plotting functions
```{r Plot generation}
myplot <- c(dr_list) %>% 
  combine_dr_list() %>%
  filter(! (r_element %in% paste0("r", c(11,22,33)))) %>% 
  # filter(dose >= -2, dose <= 2.5) %>%
  filter(par_opt_setting == 2) %>% 
  filter(alpha %in% 0) %>%
  # filter(matrix != "r_kept") %>%
  apply_expression({
    x$matrix[x$matrix == "r_alpha"] <- "r(a_i = 1)"
    x$matrix[x$matrix == "r_opt"] <- "r(a = a_opt)"
    x$matrix <- x$matrix %>% factor(levels = c("r(a_i = 1)", "r(a = a_opt)", "r_kept"))
    x
  }) %>% 
  mutate(Element = r_element) %>% 
  ggplot(mapping = aes(x = dose, y = value, color = Element, linetype = Element)) +
  geom_line() +
  theme_dMod() +
  scale_color_dMod() +
  geom_abline(slope = 0,intercept = 0, linetype = 3) +
  geom_vline(xintercept = 0, linetype = 3, size = 0.1) +
  facet_grid(matrix~., scales = "free")+
  xlab("alpha")

myplot
# plotly::ggplotly(myplot)

# ggsave(plot = myplot,
#        filename = paste0("hub",".png"),
#        width = 10,
#        height = 10,
#        units = "cm",
#        device = "png")

```












