---
title: "Prabakaran-Model"
author: "Daniel Lill"
output: 
  pdf_document:
    keep_tex: true
---

# Load libraries and define some useful paths

```{r Libraries, results='hide', echo=FALSE, message=FALSE}
library(MRAr)
library(scales)
library(xtable)
path <- "~/Promotion/Projects/MRA/Cascade/"
pathout <- paste0(path, "Output/")
```

```{r Names for Fit storage of different initial Conditions}
# fit_name <- "Multiplefits.RData"
# fit_name <- "Gain_Multiplefits.RData"
# fit_name <- paste0(c("Pert1","Pert2", "Pert3", "Pert4"), ".RData")
fit_name <- "Summary"
```

```{r Call render without overwriting any previous documents, echo=FALSE, eval=FALSE}
rmarkdown::render("Prabakaran.Rmd", output_file = paste0(pathout, format(Sys.time(), "%Y%m%d-%H%M")
, " - " , fit_name, ".pdf"))
```
# Cascade

## Setup Model 
This is the setup of the ODE model.

### Prediction function
odemodel writes a C-function, which is executed in x and xs.
x is a regular prediction function while xs only finds and returns the steady states.
If the odemodel has been calculated already, it can be loaded with "load".
```{r Prediction function Cascade}
modelpath <- paste0(path, "Models/Cascade/")
modelname   <- "Cascade"
mymodel     <- as.eqnlist(read.csv(paste0(modelpath, modelname, ".csv")))
# mymodel

# myodemodel <- odemodel(mymodel, modelname = modelname, deriv = T)
# save(myodemodel, file = paste0(modelpath, "odemodel",modelname, ".RData"))
load(paste0(modelpath, "odemodel",modelname, ".RData"))


# Prediction function
x <- Xs(odemodel = myodemodel)
# Prediction function for steady states
xs <- Xs_steady(myodemodel)
print(mymodel)
```

### Modules, their observables and the observation function
Observation function which depends also on the optimization parameters. With their help, the complexes can smoothly be switched on in the observation.

This is the naive modularization used in the Prabakaran Paper, desregarding, which species are actually the communicating ones. This is fair enough, because usually the "active" forms are consedered to be the communicating species in a non-parametric network reconstruction.
```{r Naive observation function Cascade}
modules0 <- c("x2","y2", "z2")

obs0 <- c(  "x2+a_1*Cx2y1",
            "y2+a_2*Cy2z1",
            "z2")
g0 <- Y(as.eqnvec(structure(obs0, names = modules0)), mymodel, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```

Better modularization (also, set some defaults.)
```{r good observation function Cascade}
modules1 <- c("Cx2y1","Cy2z1", "z2")

obs1 <- c("a_1*x2+Cx2y1",
         "a_2*y2+Cy2z1",
         "z2")
g <- g1 <- Y(as.eqnvec(structure(obs1, names = modules1)), mymodel, compile = TRUE, modelname = "obsfn1", attach.input = FALSE)
```


### Parameters
Load all parameters and format them for further use.
```{r Parameters Cascade}

pars_raw <- read.csv(paste0(modelpath, "pars.csv"), header = FALSE)  # dynamical parameters
pars_raw <- structure(pars_raw$V2, names = as.character(pars_raw$V1))
ic_raw <- read.csv(paste0(modelpath, "IC.csv"), header = F)          # initial conditions
ic_raw <- structure(ic_raw$V2, names = as.character(ic_raw$V1))
# sort ics like they appear in x, so it can be used in stode()
vars_x <- attr(x, "parameters")[attr(x, "parameters") %in% names(ic_raw)]
ic_raw <- ic_raw[vars_x]

pars_opt_0 <- setdiff((g) %>% attr("parameters"),names(c(pars_raw,ic_raw)))
pars_opt_0 <- structure(rep(0, length(pars_opt_0)), names = pars_opt_0)

pars_0 <- pars <- c(ic_raw, pars_raw, pars_opt_0)


# check if all parameters are assigned a value. Both lines should evaluate to the same number.
x_pars <- unique(union((g) %>% attr("parameters"),(xs) %>% attr("parameters")))
names(pars) %in% x_pars %>% sum()
x_pars %in% names(pars) %>% sum()

```


## Show the code
### Perturb modules
Now the module can be simulated to generate perturbation data.

### Parameter trafos for convenient perturbation data
Just to show the code...
Set up the perturbations
```{r  which pars are perturbed  Cascade, eval=FALSE}
# Which pars shall be perturbed?
# which_pars_perturbed_0 <- which_pars_perturbed <- c("m1", "e1")
```


```{r define perturbation conditions Cascade, eval=FALSE}
p_pert <- p_pert_fun(which_pars_perturbed = which_pars_perturbed_0,
                     perturbation_strength = 1.1,
                     pars = pars_0)
```

### Perturbation data generation and elements to keep
Just to show the code...
This function will actually be evaluated each time a new condition is tested.
```{r Generate perturbation data Cascade, eval=FALSE}
perturbation_prediction_0 <-  (xs*p_pert)(times = c(0,Inf), pars = pars_0, deriv = F)
r_kept_0 <- r_kept_fun()
```

### Fitting
Just to show the code...
```{r Fit Cascade, eval=FALSE}
myfits <- mstrust(obj_alpha, center =  pars_opt_0+0.5, studyname = "Fits", cores = 4, fits = 10, sd = 0.3)
myfits %>% as.parframe() %>% plotValues()
best_fit <- myfits %>% as.parframe()
best_fit <- best_fit[abs(best_fit$a_1) < 1 & abs(best_fit$a_2) < 1, ] %>% as.parvec()
```



## Find good Parameter values to compare behaviour for different conditions
Run this chunk of code with the "plot"-line uncommented to see predictions.
Once happy, comment out the plotting and proceed.
```{r Find good pars Cascade, eval=FALSE}
ic_list <- list(x1 = c(1000,1100),
                y1 = c(100,110)
                )

pars <- pars_0
pars[names(reduce_ic_list(ic_list)[[2]])] <- reduce_ic_list(ic_list)[[2]]
ic_list <- reduce_ic_list(ic_list)[[1]]
p_ic <- p_ic_fun(ic_list)
plot((g*x*p_ic)(times = seq(0,1000, length.out = 50), pars))
# print(pars)
```

## Fit and analyze different conditions and perturbations

### Create lists for different settings
```{r Summary of pars and ic_list Cascade}
which_pars_perturbed_list <- list()
ic_list_list <- list()
pars_list <- list()
obs_fun_list <- list()
setting_description <- c("No feedback + moiety conservation",
                         "Feedback + moiety conservation",
                         "Feedback + synthesis and degradation of z",
                         "Feedback + synthesis and degradation of all species",
                         "Feedback + moiety conservation in different parameters than at setting 2")
```

## Setting 1: No feedback + moiety conservation
### Various settings with complex concentrations comparable to/much smaller than free species
```{r Setting1  Cascade}
obs_fun_list[[1]] <- g0
which_pars_perturbed_list[[1]] <- c("x1", "y1","z1")

ic_list <- list( k4=0.1,
                 k7 = c(0.01,0.1),
                 v5 =c(5,20)) 

pars <- pars_0

pars[names(reduce_ic_list(ic_list)[[2]])] <- reduce_ic_list(ic_list)[[2]]
ic_list <- reduce_ic_list(ic_list)[[1]]
# p_ic <- p_ic_fun(ic_list)
# plot((x*p_ic)(times = seq(0,1000, length.out = 50), pars))
# 
pars_list[[1]] <- pars
ic_list_list[[1]] <- ic_list
```

## Setting 2: Feedback + moiety conservation
```{r pars_2 Cascade, echo=FALSE}
obs_fun_list[[2]] <- g0
which_pars_perturbed_list[[2]] <- c("x1", "y1","z1")

ic_list <- list( gainuf =c(0,1,20),
                 gainlf = c(0,1,20),
                 k4=0.1,
                 k7 = c(0.01,0.1)) 
pars <- pars_0

pars[names(reduce_ic_list(ic_list)[[2]])] <- reduce_ic_list(ic_list)[[2]]
ic_list <- reduce_ic_list(ic_list)[[1]]
# p_ic <- p_ic_fun(ic_list)
# plot((x*p_ic)(times = seq(0,1000, length.out = 50), pars))
# 
pars_list[[2]] <- pars
ic_list_list[[2]] <- ic_list
```

## Setting 3: Feedback + synthesis and degradation of z
Synthesis and degradation of z
```{r pars_3 Cascade, echo=FALSE}
obs_fun_list[[3]] <- g0
which_pars_perturbed_list[[3]] <- c("x1", "y1","kdegz2")

ic_list <- list( gainuf =c(0,1,20),
                 gainlf = c(0,1,20),
                 k4=0.1,
                 k7 = c(0.01,0.1),
                 ksynz1 = 5,
                 kdegz1 = 0.1,
                 kdegz2 = 0.1) 

pars <- pars_0

# pars[names(pars) %>% str_detect("syn|deg")] <- pars[names(pars) %>% str_detect("syn|deg")]*10^(-2)
pars[names(reduce_ic_list(ic_list)[[2]])] <- reduce_ic_list(ic_list)[[2]]
ic_list <- reduce_ic_list(ic_list)[[1]]
# p_ic <- p_ic_fun(ic_list)
# plot((x*p_ic)(times = seq(0,1000, length.out = 50), pars))

pars_list[[3]] <- pars
ic_list_list[[3]] <- ic_list
```

## Setting 4: Feedback + synthesis and degradation
Synthesis and degradation of everything
```{r pars_4 Cascade, echo=FALSE}
obs_fun_list[[4]] <- g0
which_pars_perturbed_list[[4]] <- c("kdegx2", "kdegy2","kdegz2")

ic_list <- list( gainuf =c(0,1,20),
                 gainlf = c(0,1,20),
                 k4=0.1,
                 k7 = c(0.01,0.1),
                 ksynz1 = 5,
                 kdegz1 = 0.1,
                 kdegz2 = 0.1,
                 ksynx1 = 5,
                 kdegx1 = 0.1,
                 kdegx2 = 0.1,
                 ksyny1 = 5,
                 kdegy1 = 0.1,
                 kdegy2 = 0.1) 

pars <- pars_0

# pars[names(pars) %>% str_detect("syn|deg")] <- pars[names(pars) %>% str_detect("syn|deg")]*10^(-2)
pars[names(reduce_ic_list(ic_list)[[2]])] <- reduce_ic_list(ic_list)[[2]]
ic_list <- reduce_ic_list(ic_list)[[1]]
p_ic <- p_ic_fun(ic_list)
plot((x*p_ic)(times = seq(0,1000, length.out = 50), pars))

pars_list[[4]] <- pars
ic_list_list[[4]] <- ic_list
```

## Setting 5: Feedback and moiety conservation, closer to the original parameters
Synthesis and degradation of everything
```{r pars_5 Cascde, echo=FALSE}
obs_fun_list[[5]] <- g0
which_pars_perturbed_list[[5]] <- c("x1", "y1","z1")
ic_list <- list(k7 = c(0.01,0.15,1), gainlf = c(0,1,10))

pars <- pars_0

# pars[names(pars) %>% str_detect("syn|deg")] <- pars[names(pars) %>% str_detect("syn|deg")]*10^(-2)
pars[names(reduce_ic_list(ic_list)[[2]])] <- reduce_ic_list(ic_list)[[2]]
ic_list <- reduce_ic_list(ic_list)[[1]]
# p_ic <- p_ic_fun(ic_list)
# plot((x*p_ic)(times = seq(0,1000, length.out = 50), pars))

pars_list[[5]] <- pars
ic_list_list[[5]] <- ic_list

```


## Fit the multiple conditions and save the fits to a file.
Run this only once before knitting and knit the document with the loaded dataset.
```{r Fit multiple conditions Cascade, eval=FALSE}
machine <- "tesla"
fit_job1<- runbg({
  myfits <- mclapply(seq_along(pars_list), function(i) {
    # cat("Setting ", i, " of ", length(pars_list), " \n")
    run_simulations(ic_list = ic_list_list[[i]],
                    nfits = 6,
                    cores = 6,
                    pars = pars_list[[i]],
                    which_pars_perturbed = which_pars_perturbed_list[[i]],
                    obs_fun = obs_fun_list[[i]])
    

  }, mc.cores = 4)
  names(myfits) <- paste0("Setting", seq_along(pars_list))
  myfits
}, machine = machine, filename = "fit_job1")

```

```{r Get the fit Cascade, eval=FALSE}
mytime <- 0

while(!fit_job1$check()) {
  cat("Running since ", mytime, "s. \n")
  Sys.sleep(5)
  mytime <- mytime + 5
}
fit_job1$check()
multiple_fits <- fit_job1$get()
multiple_fits <- multiple_fits[[machine]]
# multiple_fits
# fit_job1$purge()
save(multiple_fits, file = paste0(pathout,fit_name, "Cascade",".RData"))
```



## Analyze the fits
```{r Analyze Fits to knitr-Document Cascade, results='asis', out.width='45%'}
# Load the multiple fits
load(paste0(pathout,fit_name, "Cascade",".RData"))

lapply(seq_along(pars_list), function(setting) {
  
  cat("\\newpage
# Fit setting: ", setting, ": ", setting_description[setting], "
\n
### Standard pars of the following fits \n")
  print(xtable(as.matrix(pars_list[[setting]], nrow = 1), digits = 3), comment = F, floating = F)
    
  
  
  # multiple_fits <- lapply(seq_along(multiple_fits), function(j) {
  #   if(is.null(multiple_fits[[j]]$condition)) multiple_fits[[j]]$condition <- names(multiple_fits)[j]
  #   return(multiple_fits[[j]])
  # })
  output <- lapply(multiple_fits[[setting]], function(cond) {
    
    cat("\\newpage
      # Condition: ",(cond$condition %>% str_split("_") %>% extract2(1) %>% paste(collapse = ", ") %>% str_replace_all("=", " = ")), "\n
      # Perturbed Parameters: ", cond$which_pars_perturbed, "\n")
    
    # Plot the result of the mstrust()
    print(plotValues(cond$all_fits) )
    print(plot(cond$prediction))
    
    cat(" \n Local response matrices before and after optimization.\n")
    print(xtable(cond$r_0, digits = 5), comment = F, floating = F)
    cat(" ")
    print(xtable(cond$r_kept, digits = 5), comment = F, floating = F)
    cat(" ")
    print(xtable(cond$r_best_fit, digits = 5), comment = F, floating = F)
    
    cat("\n Best fit with reasonable parameter: \n")
    print(cond$best_fit %>% unclass %>%  matrix(.,nrow = 1, dimnames = list(NULL, names(.))) %>% xtable(digits = 3), comment = F, floating = F)
    
    cat("\n Steady state: \n")
    print(xtable(as.matrix(cond$steady_states[[1]])), comment = F, floating = F)
    
    cat("\n All fits: \n")
    print(xtable(cond$all_fits, digits = 4), comment = F, floating = F)
    return()
  })
})
```











<!-- ------------------------------------------------------------------------------------------------- -->












<!-- ------------------------------------------------------------------------------------------------- -->







# Prabakaran-Model

### Prediction function.
odemodel writes a C-function, which is executed in x and xs.
x is a regular prediction function while xs only finds and returns the steady states.
If the odemodel has been calculated already, it can be loaded with "load".
```{r Prediction function}
modelpath <- paste0(path, "Models/Prabakaran/")
modelname   <- "Prabakaran"
mymodel     <- as.eqnlist(read.csv(paste0(modelpath, modelname, ".csv")))
# mymodel

# myodemodel <- odemodel(mymodel, modelname = modelname, deriv = T)
# save(myodemodel, file = paste0(modelpath, "odemodel",modelname, ".RData"))
load(paste0(modelpath, "odemodel",modelname, ".RData"))


# Prediction function
x <- Xs(odemodel = myodemodel)
# Prediction function for steady states
xs <- Xs_steady(myodemodel)
print(mymodel)
```

### Modules, their observables and the observation function.
Observation function which depends also on the optimization parameters. With their help, the complexes can smoothly be switched on in the observation.

This is the naive modularization used in the Prabakaran Paper, desregarding, which species are actually the communicating ones. This is fair enough, because usually the "active" forms are consedered to be the communicating species in a non-parametric network reconstruction.
```{r Naive observation function}
modules0 <- c("m2","e2")

obs0 <- c(  "m2+a_1*Cm2e1",
            "e2")
g0 <- Y(as.eqnvec(structure(obs0, names = modules0)), mymodel, compile = TRUE, modelname = "obsfn0", attach.input = FALSE)
```

Better modularization (also, set some defaults.)
```{r good observation function}
modules <- modules1 <- c("Cm2e1","e1")

obs <- obs1 <- c(  "a_1*m2+Cm2e1",
                   "e1")
g <- g1 <- Y(as.eqnvec(structure(obs1, names = modules1)), mymodel, compile = TRUE, modelname = "obsfn1", attach.input = FALSE)
```


```{r Observation function with better modularization in m-module, but not in e-module.}
modules2 <- c("Cm2e1","e2")

obs2 <- c(  "a_1*m2+Cm2e1",
                   "e2")
g2 <- Y(as.eqnvec(structure(obs2, names = modules2)), mymodel, compile = TRUE, modelname = "obsfn2", attach.input = FALSE)
```


### Parameters.
Load all parameters and format them for further use.
```{r Parameters}

pars_raw <- read.csv(paste0(modelpath, "pars.csv"), header = FALSE)  # dynamical parameters
pars_raw <- structure(pars_raw$V2, names = as.character(pars_raw$V1))
ic_raw <- read.csv(paste0(modelpath, "IC.csv"), header = F)          # initial conditions
ic_raw <- structure(ic_raw$V2, names = as.character(ic_raw$V1))
# sort ics like they appear in x, so it can be used in stode()
vars_x <- attr(x, "parameters")[attr(x, "parameters") %in% names(ic_raw)]
ic_raw <- ic_raw[vars_x]

pars_opt_0 <- setdiff((g) %>% attr("parameters"),names(c(pars_raw,ic_raw)))
pars_opt_0 <- structure(rep(0, length(pars_opt_0)), names = pars_opt_0)

pars_0 <- pars <- c(ic_raw, pars_raw, pars_opt_0)


# check if all parameters are assigned a value. Both lines should evaluate to the same number.
x_pars <- unique(union((g) %>% attr("parameters"),(xs) %>% attr("parameters")))
names(pars) %in% x_pars %>% sum()
x_pars %in% names(pars) %>% sum()

```


## Perturb modules.
Now the module can be simulated to generate perturbation data.

### Parameter trafos for convenient perturbation data.
Set up the perturbations
```{r which pars are perturbed, eval=FALSE}
# Which pars shall be perturbed?
which_pars_perturbed_0 <- which_pars_perturbed <- c("m1", "e1")

p_pert <- p_pert_fun(which_pars_perturbed = which_pars_perturbed_0,
                     perturbation_strength = 1.1,
                     pars = pars_0)
```


### Perturbation data generation and elements to keep.
Just to show the code...
This function will actually be evaluated each time a new condition is tested.


```{r Generate perturbation data, eval=FALSE}
perturbation_prediction_0 <-  (xs*p_pert)(times = c(0,Inf), pars = pars_0, deriv = F)
r_kept_0 <- r_kept_fun(obs_fun = g0, obs_fun_2 = g2)

```

### Fitting.
Just to show the code...
```{r Fit, eval=FALSE}
myfits <- mstrust(obj_alpha, center =  pars_opt_0+0.5, studyname = "Fits", cores = 4, fits = 10, sd = 0.3, obs_fun = g1)
myfits %>% as.parframe() %>% plotValues()
myfits %>% as.parframe()
# best_fit <- best_fit[abs(best_fit$a_1) < 1, ] %>% as.parvec()
```


## Find good Parameter values to compare behaviour for different conditions.

Run this chunk of code with the "plot"-line uncommented to see predictions.
Once happy, comment out the plotting and proceed.
```{r Find good pars, eval=FALSE}
ic_list <- list(e1 = c(1000,1100),
                p1 = c(100,110)
                )

pars <- pars_0
pars[names(reduce_ic_list(ic_list)[[2]])] <- reduce_ic_list(ic_list)[[2]]
ic_list <- reduce_ic_list(ic_list)[[1]]
p_ic <- p_ic_fun(ic_list)
plot((g0*x*p_ic)(times = seq(0,100, length.out = 50), pars))
# print(pars)
```

## Fit and analyze different conditions and perturbations.

### Create lists for different settings.
```{r Summary of pars and ic_list}
which_pars_perturbed_list <- list()
ic_list_list <- list()
pars_list <- list()
obs_fun_list <- list()
setting_description <- c("Wrong modularization, eT perturbed",
                         "Wrong modularization, pT perturbed",
                         "Correct modularization, eT perturbed",
                         "Correct modularization, pT perturbed"
                         )
```

### Setting 1: Wrong modularization, eT perturbed.
```{r}
obs_fun_list[[1]] <- g0
which_pars_perturbed_list[[1]] <- c("m1", "e1")

ic_list <- list(e1 = c(100,100))

pars <- pars_0
# pars[names(pars) %>% str_detect("syn|deg")] <- pars[names(pars) %>% str_detect("syn|deg")]*10^(-2)
# pars[names(reduce_ic_list(ic_list)[[2]])] <- reduce_ic_list(ic_list)[[2]]
# ic_list <- reduce_ic_list(ic_list)[[1]]
# # p_ic <- p_ic_fun(ic_list)
# # plot((x*p_ic)(times = seq(0,1000, length.out = 50), pars))
# 
pars_list[[1]] <- pars
ic_list_list[[1]] <- ic_list
```

### Setting 2: Wrong modularization, pT perturbed.
```{r pars_2, echo=FALSE}
obs_fun_list[[2]] <- g0
which_pars_perturbed_list[[2]] <- c("m1", "p1")

ic_list <- list(e1 = c(100,100))

pars <- pars_0
# pars[names(pars) %>% str_detect("syn|deg")] <- pars[names(pars) %>% str_detect("syn|deg")]*10^(-2)
# pars[names(reduce_ic_list(ic_list)[[2]])] <- reduce_ic_list(ic_list)[[2]]
# ic_list <- reduce_ic_list(ic_list)[[1]]
# # p_ic <- p_ic_fun(ic_list)
# # plot((x*p_ic)(times = seq(0,1000, length.out = 50), pars))
# 
pars_list[[2]] <- pars
ic_list_list[[2]] <- ic_list
```

### Setting 3: Correct modularization, eT perturbed.
```{r pars_3}
obs_fun_list[[3]] <- g1
which_pars_perturbed_list[[3]] <- c("m1", "e1")

ic_list <- list(e1 = c(100,100))

pars <- pars_0
# pars[names(pars) %>% str_detect("syn|deg")] <- pars[names(pars) %>% str_detect("syn|deg")]*10^(-2)
# pars[names(reduce_ic_list(ic_list)[[2]])] <- reduce_ic_list(ic_list)[[2]]
# ic_list <- reduce_ic_list(ic_list)[[1]]
# # p_ic <- p_ic_fun(ic_list)
# # plot((x*p_ic)(times = seq(0,1000, length.out = 50), pars))
# 
pars_list[[3]] <- pars
ic_list_list[[3]] <- ic_list
```

### Setting 4: Correct modularization, pT perturbed.
```{r pars_4}
obs_fun_list[[4]] <- g1
which_pars_perturbed_list[[4]] <- c("m1", "p1")

ic_list <- list(e1 = c(100,100))

pars <- pars_0
# pars[names(pars) %>% str_detect("syn|deg")] <- pars[names(pars) %>% str_detect("syn|deg")]*10^(-2)
# pars[names(reduce_ic_list(ic_list)[[2]])] <- reduce_ic_list(ic_list)[[2]]
# ic_list <- reduce_ic_list(ic_list)[[1]]
# # p_ic <- p_ic_fun(ic_list)
# # plot((x*p_ic)(times = seq(0,1000, length.out = 50), pars))
# 
pars_list[[4]] <- pars
ic_list_list[[4]] <- ic_list
```


## Fit the multiple conditions and save the fits to a file..
Run this only once before knitting and knit the document with the loaded dataset.
```{r Fit multiple conditions, eval=FALSE}
machine <- "localhost"
fit_job1<- runbg({
  myfits <- lapply(seq_along(pars_list), function(i) {
    # cat("Setting ", i, " of ", length(pars_list), " \n")
    run_simulations(ic_list = ic_list_list[[i]],
                    nfits = 10,
                    cores = 3,
                    pars = pars_list[[i]],
                    which_pars_perturbed = which_pars_perturbed_list[[i]],
                    obs_fun = obs_fun_list[[i]])
    

  })
  names(myfits) <- paste0("Setting", seq_along(pars_list))
  myfits
}, machine = machine, filename = "Prabakaran_job_intermediates")
mytime <- 0

```

```{r, eval=FALSE}
while(!fit_job1$check()) {
  cat("Running since ", mytime, "s. \n")
  Sys.sleep(5)
  mytime <- mytime + 5
}
fit_job1$check()
multiple_fits <- fit_job1$get()
multiple_fits <- multiple_fits[[machine]]
fit_job1$purge()
save(multiple_fits, file = paste0(pathout,fit_name, "Prabakaran", ".RData"))
```



## Analyze the fits.
```{r Analyze Fits to knitr-Document, results='asis', out.width='45%'}
# Load the multiple fits
load(paste0(pathout,fit_name, "Prabakaran", ".RData"))

lapply(seq_along(pars_list), function(setting) {
  
  cat("\\newpage
# Fit setting: ", setting, ": ", setting_description[setting], "
\n
### Standard pars of the following fits \n")
  print(xtable(as.matrix(pars_list[[setting]], nrow = 1), digits = 3), comment = F, floating = F)
    
  
  
  # multiple_fits <- lapply(seq_along(multiple_fits), function(j) {
  #   if(is.null(multiple_fits[[j]]$condition)) multiple_fits[[j]]$condition <- names(multiple_fits)[j]
  #   return(multiple_fits[[j]])
  # })
  output <- lapply(multiple_fits[[setting]], function(cond) {
    
    cat("\\newpage
      # Condition: ",(cond$condition %>% str_split("_") %>% extract2(1) %>% paste(collapse = ", ") %>% str_replace_all("=", " = ")), "\n
      # Perturbed Parameters: ", cond$which_pars_perturbed, "\n")
    
    # Plot the result of the mstrust()
    print(plotValues(cond$all_fits) )
    print(plot(cond$prediction))
    
    cat(" \n Local response matrices before and after optimization.\n")
    print(xtable(cond$r_0, digits = 5), comment = F, floating = F)
    cat(" ")
    print(xtable(cond$r_kept, digits = 5), comment = F, floating = F)
    cat(" ")
    print(xtable(cond$r_best_fit, digits = 5), comment = F, floating = F)
    
    cat("\n Best fit with reasonable parameter: \n")
    print(cond$best_fit %>% unclass %>%  matrix(.,nrow = 1, dimnames = list(NULL, names(.))) %>% xtable(digits = 3), comment = F, floating = F)
    
    cat("\n Steady state: \n")
    print(xtable(as.matrix(cond$steady_states[[1]])), comment = F, floating = F)
    
    cat("\n All fits: \n")
    print(xtable(cond$all_fits, digits = 4), comment = F, floating = F)
    return()
  })
})
```